### 2024-04-04 Text/Identifying Heterogeneity in SAR Data with New Test Statistics.Rmd
updated paper version

---
title: "Identifying Heterogeneity in SAR Data with New Test Statistics"
author:
  - name: Alejandro C. Frery
    affil: 1,*
    orcid: 0000-0002-8002-5341
  - name: Janeth Alpala
    affil: 2 
    orcid: 0000-0002-0265-6236
  - name: Abraão D. C. Nascimento
    affil: 2
    orcid: 0000-0003-2673-219X
affiliation:
  - num: 1
    address: |
      School of Mathematics and Statistics, Victoria University of Wellington, Wellington 6140, New Zealand
#email: alejandro.frery@vuw.ac.nz
  - num: 2
    address: |
      Departamento de Estatística, Universidade Federal de Pernambuco, Recife 50670-901, PE, Brazil
#email: janeth.alpala@ufpe.br, abraao@de.ufpe.br

# author citation list in chicago format
authorcitation: |
  Frery, A.C.; Alpala, J.; Nascimento,A.D.C.
# firstnote to eighthnote
#firstnote: |
 # Current address: Updated affiliation
#secondnote: |
  #These authors contributed equally to this work.
correspondence: |
  alejandro.frery@vuw.ac.nz
# document options
journal: remotesensing
type: article
status: submit
# front matter
#simplesummary: |
  #A Simple summary goes here.
abstract: |
  This paper presents a statistical approach to identify the underlying roughness characteristics in synthetic aperture radar (SAR) intensity data. The physical modeling of this kind of data allows to use the Gamma distribution in the presence of fully-developed speckle, i.e., when there are infinitely many independent backscatterers per resolution cell. This situation is often called "homogeneous" or 
  "textureless" areas. The $\mathcal{G}_I^0$ distribution is also a widely accepted law for heterogeneous and extremely heterogeneous areas. We propose three test statistics for discriminating between homogeneous and non-homogeneous areas, i.e., between Gamma and GI0 distributed data, both with a known number of looks. The first test statistic utilizes a bootstrapped non-parametric estimator of the Shannon entropy, offering a robust evaluation in the face of uncertain distributional assumptions. The second test uses the classic coefficient of variation. The third test employs an alternative form of estimating the CV based on the ratio of the mean absolute deviation from the median to the median. We apply our test statistic to produce maps of $p$-values for the hypothesis of homogeneity. These maps identify different types of targets. We show that our proposal outperforms its competitors using simulated and SAR data.
# back matter
keywords: |
  SAR; heterogeneity; entropy;  coefficient of variation; hypothesis tests
acknowledgement: |
  This study was financed in part by the Coordenação de Aperfeiçoamento de Pessoal de Nível Superior - Brasil (CAPES) - Finance Code 001.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
funding: |
  Please add: "This research received no external funding" or "This research 
  was funded by NAME OF FUNDER grant number XXX."  and ``The APC was funded 
  by XXX''. Check carefully that the details given are accurate and use the 
  standard spelling of funding agency names at 
  \url{https://search.crossref.org/funding}, any errors may affect your future 
  funding.
institutionalreview: |
 Not applicable.
#informedconsent: |
  # Any research article describing a study involving humans should contain this 
  # statement. Please add ``Informed consent was obtained from all subjects 
  # involved in the study.'' OR ``Patient consent was waived due to REASON 
  # (please provide a detailed justification).'' OR ``Not applicable'' for 
  # studies not involving humans. You might also choose to exclude this statement 
  # if the study did not involve humans.

  # Written informed consent for publication must be obtained from participating 
  # patients who can be identified (including by the patients themselves). Please 
  # state ``Written informed consent has been obtained from the patient(s) to 
  # publish this paper'' if applicable.
dataavailability: |
  Not applicable.
conflictsofinterest: |
  The authors declare no conflict of interest.
supplementary: |
 This article is fully reproducible using RMarkdown. All code and data are accessible at
 \url{https://github.com/rjaneth/identifying-heterogeneity-in-sar-data-with-new-test-statistics} (accessed on 5 April 2024).
# abbreviations:
#   - short: MDPI
#     long: Multidisciplinary Digital Publishing Institute
#   - short: DOAJ
#     long: Directory of open access journals
#   - short: TLA
#     long: Three letter acronym
#   - short: LD 
#     long: linear dichroism

header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   - \usepackage[T1]{fontenc}
   - \usepackage{polski}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{amsmath}
   - \usepackage{multirow}
   - \usepackage{subcaption}
   - \captionsetup[sub]{labelfont=bf,textfont=bf}

bibliography: references.bib
#appendix: appendix.tex
endnotes: false
output: 
  rticles::mdpi_article:
    extra_dependencies: longtable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE)
#knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(rmutil)
library(invgamma)
library(tidyverse)
library(RColorBrewer)
library(ggsci)
library(ggpubr)
library(patchwork)
library(dplyr)
#options(kableExtra.latex.load_packages = FALSE)
#library(devtools)
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)
library(latex2exp)
library(e1071)# kurtosis
library(viridis)
library(nortest)# AD
library(MASS)#
library(fitdistrplus)# model selection AIC and BIC

theme_set(theme_minimal() +
            theme(text=element_text(family="serif"),
                  legend.position = "bottom")#  top , right , bottom , or left#, panel.grid = element_blank()
)

# if(!require("rstudioapi")) install("rstudioapi")
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# opts_chunk$set(comment="",
#                message=FALSE,
#                tidy.opts=list(keep.blank.line=TRUE, width.cutoff=120),
#                options(width=100), 
#                # cache=TRUE,
#                fig.align='center',
#                fig.height=6, fig.width=10,
#                fig.show='hold',
#                cache=TRUE)

source("../Code/R/MainFunctions/gamma_sar_sample.R")
source("../Code/R/MainFunctions/entropy_gamma_sar.R")
source("../Code/R/MainFunctions/entropy_gI0.R")
source("../Code/R/MainFunctions/gi0_sample.R")

source("../Code/R/MainFunctions/van_es_estimator.R")
source("../Code/R/MainFunctions/correa_estimator.R")
source("../Code/R/MainFunctions/bootstrap_correa_estimator_log_mean.R")
source("../Code/R/MainFunctions/ebrahimi_estimator.R")
source("../Code/R/MainFunctions/noughabi_arghami_estimator.R")
source("../Code/R/MainFunctions/vasicek_estimator.R")
source("../Code/R/MainFunctions/al_omari_1_estimator.R")
source("../Code/R/MainFunctions/al_omari_2_estimator.R")

source("../Code/R/MainFunctions/bootstrap_van_es_estimator.R")
source("../Code/R/MainFunctions/bootstrap_correa_estimator.R")
source("../Code/R/MainFunctions/bootstrap_ebrahimi_estimator.R")
source("../Code/R/MainFunctions/bootstrap_noughabi_arghami_estimator.R")
source("../Code/R/MainFunctions/bootstrap_vasicek_estimator.R")
source("../Code/R/MainFunctions/bootstrap_al_omari_1_estimator.R")
source("../Code/R/MainFunctions/bootstrap_al_omari_2_estimator.R")
#The next function contains the functions: generate_samples, calculate_bias_mse, generate_plot
source("../Code/R/Programs/functions_sample_bias_mse.R")# read_ENVI_images
source("../Code/R/Programs/read_ENVI_images.R")

```

```{r DefinitionSimulationRoutines, echo=FALSE}
#set.seed(1234567890, kind="Mersenne-Twister")
```
\setlength{\tabcolsep}{6pt} 
\newcommand{\bias}{\operatorname{Bias}}
\newcommand{\widebar}[1]{\overline{#1}}

# Introduction {#sec:Introduction}

Synthetic Aperture Radar (SAR) technology has become an important tool for environmental monitoring and disaster management, providing valuable imagery in different conditions, including day or night and diverse weather situations&nbsp;\cite{Moreira2013,Mu2019}.
However, the effective utilization of SAR data depends on a thorough understanding of its statistical properties, especially the impact of speckle effects, which are inherent in SAR data due to the coherent nature of the imaging process&nbsp;\cite{Argenti2013}.

In the presence of non-Gaussian speckle interference, SAR data requires a reliable statistical model for accurate processing. The $\mathcal{G}^0$ distribution, suitable for SAR data, includes the Gamma law as a limit case for fully developed speckle&nbsp;\cite{Ferreira2020}, providing flexibility with fewer parameters for analysis.

Our work is motivated to improve the identification of potential roughness features in SAR intensity data. 
Physical modeling of SAR data allows to take advantage of the gamma distribution in the presence of fully developed speckle, creating a scenario with an infinite number of independent backscatterers per unit of resolution, often referred to as homogeneous areas.

In this context, we present a set of three novel test statistics designed to discriminate between homogeneous and non-homogeneous areas, specifically aimed at differentiating between Gamma and $\mathcal{G}^0$ distributed data, each with a known number of appearances. 
Hence, to understand the core ideas behind our approach, we explore into the use of properties such as entropy and coefficient of variation.

The entropy is a fundamental concept in information theory with broad applications to pattern recognition, statistical physics, stochastic dynamics, and statistics. 
Shannon introduced it for a random variable in 1948&nbsp;\cite{Shannon1948} as a measure of information and uncertainty.  
In statistics, Shannon entropy is a crucial descriptive parameter, particularly for assessing data dispersion and conducting tests for normality, exponentiality, and uniformity&nbsp;\cite{Wieczorkowski1999,Zamanzade2012}. 
Entropy estimation involves several practical difficulties, especially when the model is unknown; in these cases, non-parametric methods are used.

Spacing methods have been discussed as a non-parametric approach in&nbsp;\cite{AlizadehNoughabi2010,Subhash2021}. 
This strategy offers flexibility to address a wide range of models without imposing specific parametric restrictions.

In our study, entropy serves as the fundamental parameter, providing a robust assessment by means of a bootstrapped non-parametric estimator of Shannon entropy.

Simultaneously, we use the classic coefficient of variation (CV) and an alternative form based on the ratio between the median absolute deviation and the median. 
The CV is known as an efficient and robust index of textural information, measures the homogeneity of the image&nbsp;\cite{Lopes1990}.

The CV, introduced by Pearson&nbsp;\cite{Pearson1896} in 1896, is a relative measure of dispersion widely used in various areas of applied statistics including sampling, biostatistics, medical and biological research, climatology, and other fields \cite{hendricks1936sampling, Tian2005,SubrahmanyaNairy2003,Chankham2024}.
It facilitates the comparison of variability between different populations and is especially valuable for comparing variables with different units. 
This is because when the main purpose is to compare variations of several variables, the standard deviation cannot serve as an adequate measure of variation unless all variables are expressed in the same unit of measurement and have identical mean values. 
When these requirements are not hold for the variables under study, CV is the relative measure typically used in real applications. 
The variable with the highest CV value is the one with the highest relative dispersion around the mean&nbsp;\cite{Banik2011}.

Applying these test statistics to generate $p$-value maps facilitates testing of the homogeneity hypothesis, revealing diverse types of targets within the SAR data. 
Through extensive experimentation involving real and simulated SAR data, we demonstrate the superiority of our proposed methodology over existing approaches.

The article is structured as follows: 
Section&nbsp;\ref{sec:Background} covers statistical modeling and entropy estimation for Intensity SAR data. Section&nbsp;\ref{sec:test} outlines hypothesis testing based on non-parametric entropy estimators and and coefficient of variation. 
In Section&nbsp;\ref{sec:Results}, we present experimental results. 
Finally, in Section&nbsp;\ref{sec:conclusion} conclusions are exhibited.

# Background {#sec:Background}

## Statistical Modeling of Intensity SAR data

The primary models used for intensity SAR data include the Gamma and $\mathcal{G}_I^0$  distributions&nbsp;[@Frery1997]. 
The first is suitable for fully developed speckle and is a limiting case of the second, which is appealing due to its versatility in accurately representing regions with various roughness characteristics&nbsp;[@Cassetti2022].
We denote $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ and $Z \sim G_I^0(\alpha, \gamma, L)$ to indicate that $Z$ follows the distributions characterized by the respective probability density functions:
\begin{align}
	f_Z(z;L, \mu)&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gamma1}\\
	f_Z(z; \alpha, \gamma, L)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\cdot\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gi01}
\end{align}
where, in&nbsp;\eqref{E:gamma1} $\mu > 0$ is the mean; in&nbsp;\eqref{E:gi01}  $\gamma > 0$ is the scale, $\alpha < -1$ measures the roughness,  $L \geq 1$ is the number of looks, $\Gamma(\cdot)$ is the gamma function, and $\mathbbm 1_{A}(z)$ is the indicator function of the set $A$.

From \eqref{E:gi01}, the $r$th moment of $Z$ is expressed as:
\begin{align}
	E_{G_I^0}\left(Z^r\right)=\left(\frac{\gamma}{L}\right)^r\frac{\Gamma(-\alpha-r)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+r)}{\gamma(L)}, \quad \alpha <-r. 
	\label{E:rmom}
\end{align}

Even though the $\mathcal{G}_I^0$  distribution is defined by the parameters $\alpha$ and $\gamma$, SAR literature commonly utilizes the texture $\alpha$ and the mean $\mu$&nbsp;\cite{Nascimento2010}.
In this way, we compute the expected value $\mu$ using the expression in&nbsp;\eqref{E:rmom}, and we reparametrize&nbsp;\eqref{E:gi01} using $\mu$, $\alpha$, and $L$. Then
\begin{align*}
	\mu=\left(\frac{\gamma}{L}\right)\frac{\Gamma(-\alpha-1)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+1)}{\gamma(L)}=-\frac{\gamma}{\alpha+1}.
\end{align*}
Thus, the probability density function that characterize the $G_I^0(\mu, \alpha, L)$ law is
\begin{equation}
		f_Z(z; \mu, \alpha, L)=\frac{L^L\Gamma(L-\alpha)}{\big(-\mu(\alpha+1)\big)^{\alpha}\Gamma(-\alpha)\Gamma(L)} \frac{z^{L-1}}{\big(-\mu(\alpha+1)+Lz\big)^{L-\alpha}}.\label{E:gi02}
\end{equation}

## The Shannon Entropy
The parametric representation of Shannon entropy for a system described by a continuous random variable is:
\begin{equation}
  \label{E:entropy2}
  H(Z)=-\int_{-\infty }^\infty \ f(z)\ln f(z)\, \mathrm{d}z,
\end{equation}
here, $f(\cdot)$ is the probability density function that characterizes the distribution of the real-valued random variable $Z$.

Using&nbsp;\eqref{E:entropy2}, we can express the Shannon entropy of $\Gamma_{\text{SAR}}$in&nbsp;\eqref{E:gamma1} and $G_I^0$ in&nbsp;\eqref{E:gi02} based on&nbsp;\cite{Cassetti2022} and&nbsp;\cite{Ferreira2020}:
\begin{equation}
\label{E:E-gamma}
H_{\Gamma_{\text{SAR}}}(L, \mu) =   L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) + \ln \mu, 
\end{equation}
\begin{multline}
\label{E:E-GIO}
H_{G_I^0}(\mu, \alpha, L) =L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) +\ln \mu -\ln\Gamma(L-\alpha)\\
+ (L-\alpha) \psi^{(0)}(L-\alpha)-(1-\alpha)\psi^{(0)}(-\alpha)+\ln (-1-\alpha)+\ln\Gamma(-\alpha)-L,
\end{multline}
where $\psi^{(0)}(\cdot)$ is the digamma function. Figure&nbsp;\ref{fig:Plot_GI0_to_gamma1}, shows the behaviour of the entropy of \(G_I^0\) when when \(\alpha \in \left\{-\infty, -20, -8, -3\right\}\), observing its convergence towards the entropy of \(\Gamma_{\text{SAR}}\) as \(\alpha\) takes large negative values.

```{r Plot_GI0_to_gamma1, echo=FALSE, message=FALSE, warning=FALSE, out.width="70%", fig.show="hold", fig.pos="hbt", fig.cap="$H_{ G_I^0}$ converges to the $H_{\\Gamma_{\\text{SAR}}}$ with $L=8$."}

L <- c(8)
alphas <- c(  -3, -8, -20, -1000)
alpha_labels <- c( expression(italic(alpha) == -3), expression(italic(alpha) == -8), expression(italic(alpha) == -20), expression(italic(alpha) == -1000))

mu <- seq(0.1, 10, length.out = 500)

# Entropy GI0

muEntropy <- data.frame()

for (alpha in alphas) {
  entropies_GI0 <- sapply(L, function(L) entropy_gI0(mu, alpha, L))
  muEntropy <- rbind(muEntropy, data.frame(mu = mu, Entropy = entropies_GI0, alpha = as.factor(alpha)))
}

muEntropy.molten <- melt(muEntropy, id.vars = c("mu", "alpha"),  value.name = "Entropy")

# Entropy Gamma SAR

entropies_gamma <- sapply(L, function(L) entropy_gamma_sar(L, mu))

Entropy_gamma <- data.frame(mu, entropies_gamma)

Entropy_gamma.molten <- melt(Entropy_gamma, id.vars = "mu", value.name = "Entropy Gamma")

#plot

ggplot(muEntropy.molten, aes(x = mu, y = Entropy, col = alpha)) +
  geom_line(data = Entropy_gamma.molten, aes(x = mu, y = `Entropy Gamma`), color = "blue", linetype = "solid",linewidth = 1.5) + 
  geom_line(linetype = "longdash",  linewidth = 2, alpha=.7) +
  annotate("text", x = max(mu)+0.2, y = max(Entropy_gamma.molten$`Entropy Gamma`), 
           label = TeX("$\\Gamma_{\\tiny{SAR}}$"), vjust = 0.9, hjust = 0.1, color = "blue")+
  theme_minimal() +
  scale_color_manual(values = brewer.pal(7, "Dark2")[1:5], labels = alpha_labels) +
  #scale_color_manual(values = pal_jama()(7)[2:5], labels = alpha_labels) +
  #scale_color_manual(values = 	pal_cosmic()(7)[1:5], labels = alpha_labels) +
  labs(col = "Roughness", linetype = NULL) +
  xlab(expression(paste(mu))) +
  ylab("Entropy") +  
  theme(text = element_text(family = "serif"),
        legend.position = "bottom")

```

## Estimation of the Shannon Entropy

The problem of non-parametric estimating of $H(Z)$ has been considered by many authors including&nbsp;\cite{vasicek1976test, Wieczorkowski1999,correa1995new,AlOmari2019}, who proposed estimators based on spacings.

One of the earliest non-parametric estimators relying on spacings was introduced by Vasicek&nbsp;\cite{vasicek1976test}. 
Assuming that $\bm{Z}=(Z_1, Z_2,\ldots,Z_n)$ is a random sample from the distribution $F(z)$, the estimator is defined as:
\begin{equation*}
\label{E:Vas}
	\widehat{H}_{\text{V}}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\ln\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, $Z_{(i+m)}-Z_{(i-m)}$ is the $m$-spacing and $Z_{(1)}\leq Z_{(2)}\leq\ldots\leq Z_{(n)}$ are the order statistics and $Z_{(i)}= Z_{(1)}$ if $i<1$, $Z_{(i)}= Z_{(n)}$ if $i>n$.

Several authors have explored adaptations to Vasicek's estimator. 
We consider three estimators known for their superior performance&nbsp;\cite{Cassetti2022}:

* Correa&nbsp;\cite{correa1995new}: $\widehat{H}_{\text{C}}$.

* Ebrahimi et al&nbsp;\cite{Ebrahimi1994}: $\widehat{H}_{\mathrm{E}}$.

* Al Omary&nbsp;\cite{IbrahimAlOmari2014}: $\widehat{H}_{\mathrm{AO}}$.

## Enhanced Bootstrap Technique

We employ the bootstrap technique to refine the precision of existing non-parametric entropy estimators. 
This approach involves generating new datasets through resampling with repetition from an existing one&nbsp;\cite{Michelucci2021}.

<!-- This approach involves generating new datasets through resampling with repetition from an existing one. -->

Let's assume that non-parametric entropy estimators \(\widehat{H}=\widehat{\theta}(\bm{Z})\) are inherently biased, that is:
\begin{equation}
\label{Eq:bias1}
\bias\big(\widehat{\theta}(\bm{Z})\big) = E\big[\widehat{\theta}(\bm{Z})\big] - \theta \neq 0.
\end{equation}
Our bootstrap-improved estimators are of the form:
<!-- introducing an "ideal unbiased estimator" \(\check{\theta}(\bm{Z})\): -->
<!-- \begin{equation} -->
<!-- \label{Eq:bias2} -->
<!-- \widecheck{\theta}(\bm{Z}) = \widehat{\theta}(\bm{Z}) - \bias\big(\widehat{\theta}(\bm{Z})\big). -->
<!-- \end{equation} -->
<!-- However, \(\check{\theta}(\bm{Z})\) is not an estimator, because it depends on the true parameter \(\theta\), prompting the formulation of a new estimator \(\widetilde{H}\).  -->
<!-- From&nbsp;\eqref{Eq:bias1} and&nbsp;\eqref{Eq:bias2} we have: -->
\begin{align*}
\widetilde{H} &= 2\widehat{\theta}(\bm{Z}) - \frac{1}{B}\sum_{b=1}^B \widehat{\theta}_b(\bm{Z}^{(b)}),
\end{align*}
where $B$ is the number of replications in the bootstrap technique.
Applying this methodology, the original estimators by Correa, Ebrahimi, and Al-Omari are now denoted as the proposed bootstrap-enhanced versions: $\widetilde{H}_{\text{C}}$, $\widetilde{H}_{\text{E}}$, and $\widetilde{H}_{\text{AO}}$, respectively.

We analyzed the performance of these estimators with a Monte Carlo study:
$500$ samples of $G_I^0$ of size \(n\in\left\{9, 25, 49, 81, 121\right\}\) and several values of the mean, number of looks and roughness.
We present the results with \(mu\in\left\{1, 10\right\}\), \(\alpha=-10\), and \(L=5\). 
The results are consistent for the other situations. 
In the case of the bootstrap technique, each sample is replicated $100$ times with replacement.
We choose to use the following heuristic formula for spacing, $m=\left[\sqrt{n}+0.5\right]$.

In Figure&nbsp;\ref{fig:Plot_bias_mse_gi0} we depict comparisons of bias and mean squared error (MSE) between the original non-parametric entropy estimators and their respective bootstrap-enhanced versions. 
The use of the bootstrap technique exhibits more precision, reduced bias and MSE, and improved convergence.
The results of simulation are exhibited in Table&nbsp;\ref{tab:table2}.

The precision of estimators, as evidenced by bias and MSE comparisons, benefits significantly from the bootstrap technique, particularly for sample sizes below 81.

```{r Simulated_data_gi0, echo=FALSE, message=FALSE}

set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c(9, 25, 49, 81, 121 )

# Number of replications
R <-200

# Number of bootstrap replications
B1 <- 5
mu_values <- c(1, 10)
alpha <- -20
L <- 5

estimators <- list(
  "Correa" = correa_estimator,
  "Ebrahimi" = ebrahimi_estimator,
  "Al Omari" = al_omari_1_estimator,
  "Correa Bootstrap" = bootstrap_correa_estimator,
  "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
  "Al Omari Bootstrap" = bootstrap_al_omari_1_estimator
)

calculate_results_gi0 <- function(sample_sizes, R, B1, mu_values, alpha, L, estimators) {
  results_list <- list()

  for (mu_val in mu_values) {

    results <- calculate_bias_mse_gi0(sample_sizes, R, B1, mu_val, alpha, L, estimators)
    df <- as.data.frame(results)

    results_list[[as.character(mu_val)]] <- df
  }

  return(results_list)
}

results_gi0 <- calculate_results_gi0(sample_sizes, R, B1, mu_values, alpha, L, estimators)

save(results_gi0, file = "./Data/resultsgi0_1.Rdata")

```

```{r Plot_bias_mse_gi0, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.pos="hbt",  fig.cap="Bias and MSE, $L=5$ and  $\\alpha=-10$.", fig.width=8, fig.height=5.5}

load("./Data/resultsgi0_1.Rdata")

estimators_to_plot <- c("Correa", "Ebrahimi", "Al Omari",  "Correa Bootstrap", "Ebrahimi Bootstrap", "Al Omari Bootstrap" )
  latex_estimator_names <- c("Correa" = expression("$\\widehat{italic(H)}_{C}$"),
                             "Correa Bootstrap" = expression("$\\widetilde{italic(H)}_{C}$"),
                            "Ebrahimi" = expression("$\\widehat{italic(H)}_{E}$"),
                            "Al Omari" = expression("$\\widehat{italic(H)}_{AO}$"),
                            "Ebrahimi Bootstrap" = expression("$\\widetilde{italic(H)}_{E}$"),
                            "Al Omari Bootstrap" = expression("$\\widetilde{italic(H)}_{AO}$"))
selected_estimators_latex <- latex_estimator_names[estimators_to_plot]

combined_plot_gi0 <- generate_plot_gi0_esp(results_gi0, mu_values, selected_estimators_latex, ncol = 1, nrow = 2)

print(combined_plot_gi0)

```

```{r Table_gi0, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#cat("\\setlength{\\tabcolsep}{0.5pt}\n")
load("./Data/resultsgi0_1.Rdata")

# Define estimators
estimators_to_table <- c( "Al Omari","Correa", "Ebrahimi", "Al Omari Bootstrap",  "Correa Bootstrap","Ebrahimi Bootstrap")

# Filter and reshape data
filtered_results <- purrr::map_dfr(results_gi0, ~ .x %>% filter(Estimator %in% estimators_to_table), .id = "mu")

reshaped_results <- filtered_results %>%
  pivot_wider(names_from = Estimator, values_from = c(Bias, MSE))

colnames(reshaped_results) <- c("$\\mu$", "$n$", "$\\widehat{H}_{\\text{AO}}$", "$\\widehat{H}_{\\text{C}}$", "$\\widehat{H}_{\\text{E}}$", "$\\widetilde{H}_{\\text{AO}}$", "$\\widetilde{H}_{\\text{C}}$", "$\\widetilde{H}_{\\text{E}}$",  "$\\widehat{H}_{\\text{AO}}$", "$\\widehat{H}_{\\text{C}}$", "$\\widehat{H}_{\\text{E}}$","$\\widetilde{H}_{\\text{AO}}$", "$\\widetilde{H}_{\\text{C}}$", "$\\widetilde{H}_{\\text{E}}$")

reshaped_results[] <- lapply(reshaped_results, function(x, col_name) {
  if (is.numeric(x)) {
    if (col_name == "\\mu") {
      formatted_numbers <- sprintf("$\\scriptstyle %d$", x)
    } else {
      if (all(x %% 1 == 0)) {
        formatted_numbers <- sprintf("$%d$", x)
      } else {
        formatted_numbers <- ifelse(x < 0, sprintf("$%.3f$", x), sprintf("$\\phantom{-}%.3f$", x))
      }
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
}, col_name = names(reshaped_results)[1])

#print(
 # kbl(
knitr::kable(
    reshaped_results,
    caption = "Bias and MSE of non-parametric entropy estimators and bootstrap versions.",
    format = "latex",
    booktabs = TRUE,
    align = "lrrrrrrrrrrrrrrr",
    escape = FALSE, 
    digits = 2,  
    label = "table2",
    centering = FALSE,
    table.envir = "table", position = "H") %>%
    add_header_above(c(" ", " ", "Bias" = 6,  "MSE" = 6)) %>%
    collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle") %>%
    row_spec(0,  align = "c")%>%
    kable_styling(latex_options = "scale_down")
  #)
  #kable_styling(latex_options = c("scale_down"))

```

## Coefficient of variation and a robust alternative

The population CV is defined as a ratio of the population standard deviation $(\sigma)$ to the population mean $(\mu)$:

\begin{align}
	\text{CV}=\frac{\sigma}{\mu}, \quad \mu \neq 0.
\end{align}

<!-- In practice, the population CV is unknown and needs to be estimated.  -->
<!-- The typical sample estimate of the population CV is given as: -->
<!-- \begin{align} -->
<!-- 	\text{CV}=\frac{S}{\bar{X}},  -->
<!-- \end{align} -->
<!-- where $S$ and $\bar{X}$ are the sample standard deviation and the sample mean, respectively. -->

We explore a robust alternative to the CV as discussed in \cite{Ospina2019}, involving the ratio of the mean absolute deviation from the median (MnAD) to the median, two well-known robust measures of scale and location, respectively. The MnAD is defined as $n^{-1}\sum_{i=1}^n|x_i-\widehat{Q}_2|$, where $\widehat{Q}_2$ represents an estimate of the population median, which can be approximated by the sample median.

# Hypothesis Testing  {#sec:test}

We aim at testing the following hypotheses:

$$
 \begin{cases}\mathcal{H}_0: \text{ The data comes from the } \Gamma_{\text{SAR}}\\ 
  \mathcal{H}_1:\text{ The data comes from the } G_I^0.\end{cases}
$$
In other words, we verify the hypothesis that the data are fully-developed speckle.
The proposed tests are described below.

## The Proposed Test Based on Non-parametric Entropy

The first test statistic assess the behavior of the data under the null hypothesis, through the empirical distribution.
\begin{equation}
\label{Eq:test_e}
S_H= \widetilde{H}-\left[H_{\Gamma_{\text{SAR}}}(L)+\ln \widebar{Z}\right].
\end{equation}
The values should be around zero under the null hypothesis and far otherwise. 
<!-- This suggests significant differences and implies the presence of heterogeneous clutter. -->

Figure&nbsp;\ref{fig:Plot_density} depicts empirical test statistic densities for various sample sizes, and Table&nbsp;\ref{tab:table_stat_combined} summarizes key statistics, including mean, SD, variance (Var), skewness (SK), excess kurtosis (EK), and Anderson–Darling $p$-values for normality. 
<!-- ACF Only show L=2 and L=8 -->
Results with $p$-values larger than $0.05$  indicate no evidence against normality. 
Small variance and near-zero skewness and excess kurtosis suggest limited spread, symmetry, and light tails.  
Normal QQ-plots, confirm that there is no evidence to reject the normal distribution.

<!-- In Fig.&nbsp;\ref{fig:Plot_normality_qq}, normal QQ-plots visually affirm distributional similarity with the expected normal line. -->

```{r Simulated_density, echo=FALSE, message=FALSE}
#, cache = TRUE, autodep = TRUE
set.seed(1234567890, kind = "Mersenne-Twister")

R <- 500
mu <- 1
B <- 1

sample.size <- c(25,49, 81, 121)
L_values <- c(3,  5, 8, 11)

all_summary_stats <- list()
all_TestStatistics <- list()

# For each L
for (L in L_values) {
  TestStatistics1 <- list()  
  summary_stats <- data.frame( LValue = character(),
                              SampleSize = numeric(),
                              Mean = numeric(),
                              SD = numeric(),
                              Variance = numeric(),
                              #CV= numeric(),
                              Skewness = numeric(),
                              Kurtosis = numeric(),
                              adpvalue = numeric()
                              )

  for (s in sample.size) {
    TestStat1 <- numeric(R)

    for (r in 1:R) {
      z <- gamma_sar_sample(L, mu, s)
      #z <- gi0_sample(mu, alpha1, L, s)
      TestStat1[r] <- bootstrap_correa_estimator_log_mean(z, B) + (-L + log(L) - lgamma(L) - (1 - L) * digamma(L))
    }

    TestStatistics1[[as.character(s)]] <- data.frame("SampleSize" = rep(s, R), "Test_Statistics" = TestStat1)

    mean_val <- mean(TestStat1)
    sd_val <- sd(TestStat1)
    var_val <- var(TestStat1)
    skewness_val <- skewness(TestStat1)
    kurtosis_val <- kurtosis(TestStat1)
   # lillie_p_value <- lillie.test(TestStatistics1[[as.character(s)]]$Test_Statistics)$p.value
    ad_p_value <- ad.test(TestStatistics1[[as.character(s)]]$Test_Statistics)$p.value
    #cv_val <- abs(sd_val / mean_val)

    summary_stats <- rbind(summary_stats, data.frame(LValue = as.character(L),
                                                     SampleSize = s,
                                                     Mean = mean_val,
                                                     SD = sd_val,
                                                     Variance = var_val,
                                                      #CV = cv_val,
                                                     Skewness = skewness_val,
                                                     Kurtosis = kurtosis_val,
                                                     adpvalue = ad_p_value
                                                     ))

  }

  all_TestStatistics[[as.character(L)]] <- TestStatistics1
  all_summary_stats[[as.character(L)]] <- summary_stats

  save(all_TestStatistics, all_summary_stats, file = paste0("./Data/results1_", L, ".Rdata"))
}

```

```{r Plot_density, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.pos="hbt",  fig.cap="Empirical density  under the null hypothesis.", fig.width=6, fig.height=5.0}

theme_set(theme_minimal() +
            theme(text = element_text(family = "serif"),
                  legend.position = "bottom",
                  legend.text = element_text(angle = 0, vjust = 0.5)))

all_plots <- list()

for (L in L_values) {
  load(paste0("./Data/results1_", L, ".Rdata"))

  combined_data <- do.call(rbind, all_TestStatistics[[as.character(L)]])
  combined_data$L <- L

  p <- ggplot(combined_data, aes(x = Test_Statistics, col = factor(SampleSize), linetype = factor(SampleSize))) +
    geom_line(stat = "density", linewidth = 1.0) +
    scale_color_viridis(discrete = TRUE, option = "C", direction = -1, begin = 0.1, end = 0.8, name = "Sample Size") +
    scale_linetype_manual(values = rep("solid", length(sample.size)), name = "Sample Size") +
    labs(x = "Test Statistics", y = "Density") +
    ggtitle(bquote(italic(L) == .(L))) +
    theme(plot.title = element_text(hjust = 0.5))#titulo centrado

  if (!any(L == L_values)) {
    p = p
  }

  all_plots[[as.character(L)]] <- p
}

combined_plot <- wrap_plots(all_plots, ncol = 2, nrow = 2) +
  plot_layout(guides = "collect")

print(combined_plot)

```

```{r Table_statistic, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#cat("\\setlength{\\tabcolsep}{1.0pt}\n")

load("./Data/results1_3.Rdata")
summary_stats_2 <- all_summary_stats[["3"]]

load("./Data/results1_5.Rdata")
summary_stats_5 <- all_summary_stats[["5"]]

load("./Data/results1_8.Rdata")
summary_stats_8 <- all_summary_stats[["8"]]

load("./Data/results1_11.Rdata")
summary_stats_11 <- all_summary_stats[["11"]]

combined_summary_stats <- rbind(
  transform(summary_stats_2),
  transform(summary_stats_5),
  transform(summary_stats_8),
  transform(summary_stats_11)

)
colnames(combined_summary_stats) <- c("$L$", "$n$", "Mean", "SD", "Var", "SK", "EK", "$p$-value")

# LaTeX notation 
combined_summary_stats[] <- lapply(combined_summary_stats, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.4f$", x), sprintf("$\\phantom{-}%.4f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

table_result_combined <- knitr::kable(
  combined_summary_stats,
  caption = "Descriptive analysis of $S_H$ with $L\\in\\left\\{3,5, 8,11\\right\\}$ and $\\mu=1$.",
  format = "latex",
  booktabs = TRUE,
  align = "lrrrrrrr",
  escape = FALSE,
  digits = 4,
  label = "table_stat_combined",
  centering = FALSE,
  table.envir = "table", position = "H") %>%
  #table.env = 'table*' # to span multiple columns
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")

print(table_result_combined)

```

```{r Plot_normality_qq, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.pos="hbt",  fig.cap="Normal QQ-plots for  $n=121$.", fig.width=6, fig.height=5}

theme_set(theme_minimal() +
            theme(text = element_text(family = "serif"),
                  legend.position = "bottom",
                  legend.text = element_text(angle = 0, vjust = 0.5)))

# Load saved results
load("./Data/results1_3.Rdata")
load("./Data/results1_5.Rdata")
load("./Data/results1_8.Rdata")
load("./Data/results1_11.Rdata")

# Specific sample size
selected_sample_size <- 121

# 
ggplot_list <- list()

#  L
for (L in c(3, 5, 8, 11)) {
  TestStatistics1 <- all_TestStatistics[[as.character(L)]]

  # 
  if (as.character(selected_sample_size) %in% names(TestStatistics1)) {
    p <- ggplot(TestStatistics1[[as.character(selected_sample_size)]], 
                aes(sample = Test_Statistics)) +
      geom_qq(col = "blue", shape = 20) +
      geom_qq_line(col = "darkgreen", linetype = 2) +
      labs(title = bquote(italic(L) == .(L)),
           x = "Theoretical Quantiles",
           y = "Sample Quantiles") +
      theme(plot.title = element_text(hjust = 0.5))#titulo centrado

    ggplot_list[[as.character(L)]] <- p
  } else {
    warning(paste("No size sample", selected_sample_size, " L =", L))
  }
}
combined_plot1 <- wrap_plots(ggplot_list, ncol = 2, nrow = 2) +
  plot_layout(guides = "collect")

print(combined_plot1)

# }

```

After verifying the normality of the data, we proceeded to explore the capabilities of the proposed test in terms of size and power.
Under $\mathcal{H}_0$, the distribution of test statistic is asymptotically normal. Therefore, the $p$-value  is calculated as $2\Phi(-\mid \varepsilon\mid)$, where  $\Phi$ is the standard Gaussian cumulative distribution and $\varepsilon$ is the standardized test statistic given by:
$$
\varepsilon=\frac{\tilde{H}-H_{\Gamma_{\text{SAR}}}}{\hat\sigma}.
$$

We considered nominal levels of \SI{1}{\percent}, \SI{5}{\percent}, and \SI{10}{\percent}. 
In reference to the size, $1000$ simulations have been used for various sample sizes with  satisfactory results. In all cases, the nominal level has been achieved. 
The power, in general, improves with the increase in the sample size and the number of looks. 
The results are presented in Table&nbsp;\ref{tab:table_size_power}.

```{r Simulated_error_type_I, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

# 
calculate_p_value <- function(test_statistic, mu_W, sigma_W, alpha_nominal) {
  epsilon <- test_statistic / sigma_W
  p_value <- 2 * (1 - pnorm(abs(epsilon)))

  return(p_value < alpha_nominal)  
}

R <- 100
mu <- 1
B <- 1
L_values <- c( 3, 5, 8, 11)
sample_sizes <- c( 25, 49, 81, 121)
alpha_nominals <- c(0.01, 0.05, 0.1)

results <- data.frame()

for (L in L_values) {
  for (alpha_nominal in alpha_nominals) {
    TestStatistics <- NULL
    mean_entropy <- numeric(length(sample_sizes))
    sd_entropy <- numeric(length(sample_sizes))

    for (s in sample_sizes) {
      TestStat <- numeric(R)

      for (r in 1:R) {
        z <- gamma_sar_sample(L, mu, s)
        TestStat[r] <- bootstrap_correa_estimator_log_mean(z, B) - (L - log(L) + lgamma(L) + (1 - L) * digamma(L))
      }
      mean_entropy[sample_sizes == s] <- mean(TestStat)
      sd_entropy[sample_sizes == s] <- sd(TestStat)

      TestStatistics <- rbind(TestStatistics, data.frame("L" = rep(L, R), "Sample_Size" = rep(s, R), "Test_Statistics" = TestStat))
    }

    mu_W <- mean_entropy
    sigma_W <- sqrt(sd_entropy^2)

    p_values <- apply(TestStatistics, 1, function(row) {
      calculate_p_value(row["Test_Statistics"], mu_W[sample_sizes == row["Sample_Size"]], sigma_W[sample_sizes == row["Sample_Size"]], alpha_nominal)
    })

    result <- data.frame("L" = TestStatistics$L, "Sample_Size" = TestStatistics$Sample_Size, "Alpha_Nominal" = alpha_nominal, "P_Value" = p_values)
    results <- rbind(results, result)
  }
}

save(results, file = "./Data/type_I_results1.Rdata")

```

```{r Simulated_power, echo=FALSE, message=FALSE}

set.seed(1234567890, kind = "Mersenne-Twister")

calculate_p_value <- function(test_statistic, mu_W, sigma_W) {
  epsilon <- test_statistic  / sigma_W
  p_value <- 2 * (1 - pnorm(abs(epsilon)))

  return(p_value)
}

calculate_type_II_error_rate <- function(p_values, alpha_nominal) {
  type_II_error_rate <- sum(p_values >= alpha_nominal) / length(p_values)
  return(type_II_error_rate)
}

# Function to calculate power for different L values and alpha_nominals
calculate_power <- function(R, mu, L_values, B, sample_sizes, alpha_nominals) {
  results <- data.frame()

  for (L in L_values) {
    for (alpha_nominal in alpha_nominals) {
      TestStatistics <- list()
      mean_entropy <- numeric(length(sample_sizes))
      sd_entropy <- numeric(length(sample_sizes))

      for (s in sample_sizes) {
        TestStat <- numeric(R)

        for (r in 1:R) {
          z <- gi0_sample(mu, -5, L, s)
          TestStat[r] <- bootstrap_correa_estimator_log_mean(z, B) - (L - log(L) + lgamma(L) + (1 - L) * digamma(L))
        }

        mean_entropy[sample_sizes == s] <- mean(TestStat)
        sd_entropy[sample_sizes == s] <- sd(TestStat)

        TestStatistics[[as.character(s)]] <- TestStat
      }

      mu_W <-  mean_entropy  
      sigma_W <- sqrt(sd_entropy^2)

      p_values <- lapply(TestStatistics, function(TestStat) {
        apply(data.frame("Test_Statistics" = TestStat), 1, function(row) {
          calculate_p_value(row["Test_Statistics"], mu_W, sigma_W[as.numeric(names(TestStatistics)) == as.numeric(s)])
        })
      })

      type_II_error_rates <- sapply(p_values, function(p_values_for_size) {
        calculate_type_II_error_rate(p_values_for_size, alpha_nominal)
      })

      power <- 1 - type_II_error_rates

      result_row <- data.frame(
        L = L,
        alpha_nominal = alpha_nominal,
        Sample_Size = sample_sizes,
        power = power
      )

      results <- rbind(results, result_row)
    }
  }

  return(results)
}

R <- 100
mu <- 1
L_values <- c(3, 5,  8, 11)
B <- 1
sample_sizes <- c(25, 49, 81, 121)
alpha_nominals <- c(0.01, 0.05, 0.1)

results_power <- calculate_power(R, mu, L_values, B, sample_sizes, alpha_nominals)
save(results_power, file = "./Data/results_power1.Rdata")

```

## The Proposed Test Based on Coefficient of Variation and a Robust Alternative

In addition to the entropy-based test, we also propose a test statistic based on the classical CV. This test statistic is defined as follows:
\begin{align}
	T_{\text{CV}}=\frac{S}{\widebar{Z}},
\end{align}
where $S$ and $\widebar{Z}$ are the sample standard deviation and the sample mean, respectively.

Similarly, we use another test statistic based on the ratio of the MnAD to the median. This statistic is given by:
\begin{align}
	T_{\text{CV}_{\text{MnAD}}}=\frac{\text{MnAD}}{\text{Median}}.
\end{align}
These test statistics allow assessing the variability of the data.

The situations where the use of CV and $\text{CV}_{\text{MnAD}}$ may be appropriate, i.e. where the observations are positive, the log-normal (LN) and inverse Gaussian (IG) distribution are often more justified compared to Gamma and Weibull distributions \cite{Chaubey2017,takagi1997application}.

The inverse Gaussian distribution is shown to be well approximated by the log-normal distribution \cite{whitmore1978} which implies that the IG distribution also does not share the problem of non-existence of fixed width confidence interval with the Gaussian case.

The LN distribution in two parameters has the following density function.
\begin{align}
	f_Z(z;\mu, \sigma )=\frac{1}{\sigma z\sqrt{2\pi}}\exp\left\{-\frac{(\ln z- \mu)^2}{2\sigma^2}\right\}, \quad z>0, \sigma>0, -\infty < \mu < \infty,
\end{align}
here, $\mu$ is the log-scale parameter and $\sigma$ is the shape parameter.

## Model Selection Criterion

Akaike information criterion (AIC) and Bayesian information criterion (BIC) are used to select the best fitting distribution.

The AIC is a fine technique that deals with the trade-off between the goodness-of-fit of the model and the simplicity of the model in terms of the number of model parameters&nbsp;\cite{Burnham2004}. 
The model or distribution with the lowest value of AIC is chosen to be the best.
The BIC assesses goodness-of-fit of a distribution or model, but avoids overfitting by penalising an additional degree of freedom&nbsp;\cite{Dziak2019}. 
The model with the lowest BIC value is chosen as the best.

The AIC and BIC results in Tables&nbsp;\ref{tab:table_aic_gamma}-\ref{tab:table_aic_gio_MnADmedian} indicate that the CV and $\text{CV}_{\text{MnAD}}$ data from different synthetic sample sizes, extracted from the Gamma SAR and $G_I^0$ distributions, respectively, adhere to the characteristics of an LN distribution.
The AIC and BIC values of the IG distribution and the LN distribution are not significantly different because both distributions are based on the right-skewed characteristic.
In this situation, the model for the LN distribution was considered to be the best due to it providing the lowest AIC and BIC values.

Figures&nbsp;\ref{fig:Plot_cv}-\ref{fig:Plot_madmed_gi0_MnADmedian} present density plots, Q-Q plots, P-P plots, and Cumulative Distribution Function plots, which offer an additional measure of goodness-of-fit and confirm that the most suitable distribution to be considered is the LN distribution.

```{r Simulated_AIC_BIC_gamma, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")
R <- 15000
mu <- 1
L <- 5
#B <- 5
#alpha1 <- -3
sample.sizes <- c(25, 49, 81, 121)

results_AIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

results_BIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

for (s in sample.sizes) {
  TestStatistics <- numeric(0)

  for (r in 1:R) {
    z <- gamma_sar_sample(L, mu, s)
    #z <- gi0_sample(mu, alpha1, L, s)
    TestStat <- sd(z) / mean(z)
    TestStatistics <- c(TestStatistics, TestStat)
  }

  fit_weibull <- fitdistr(TestStatistics, "weibull")
  fit_lognormal <- fitdistr(TestStatistics, "lognormal")
  fit_gamma <- fitdistr(TestStatistics, "gamma")
  fit_normal <- fitdistr(TestStatistics, "normal")

  AIC_values <- c(AIC(fit_weibull), AIC(fit_lognormal), AIC(fit_gamma), AIC(fit_normal))
  BIC_values <- c(BIC(fit_weibull), BIC(fit_lognormal), BIC(fit_gamma), BIC(fit_normal))

  dinvgauss <- function(x, mu, lambda) {
    sqrt(lambda / (2 * pi * x^3)) * exp(-lambda * (x - mu)^2 / (2 * mu^2 * x))
  }

  negLL_invgauss <- function(params, data) {
    mu <- params[1]
    lambda <- params[2]
    CV <- data

    -sum(log(dinvgauss(CV, mu, lambda)))
  }

  fit_invGauss <- fitdist(data = TestStatistics, distr = "invgauss", method = "mle",
                          start = list(mu = mean(TestStatistics), lambda = var(TestStatistics)))

  log_likelihood <- logLik(fit_invGauss)
  num_params <- length(coef(fit_invGauss))
  n <- length(TestStatistics)
  AIC_manual <- -2 * log_likelihood + 2 * num_params
  BIC_manual <- -2 * log_likelihood + num_params * log(n)

  results_AIC <- rbind(results_AIC, data.frame(Test = "AIC",
                                               SampleSize = s,
                                               Normal = AIC_values[4],
                                               Lognormal = AIC_values[2],
                                               Gamma = AIC_values[3],
                                               Weibull = AIC_values[1],
                                               InverseGaussian = AIC_manual))

  results_BIC <- rbind(results_BIC, data.frame(Test = "BIC",
                                               SampleSize = s,
                                               Normal = BIC_values[4],
                                               Lognormal = BIC_values[2],
                                               Gamma = BIC_values[3],
                                               Weibull = BIC_values[1],
                                               InverseGaussian = BIC_manual))
}

save(results_AIC, results_BIC, file = "./Data/results_AIC_BIC_gamma.Rdata")

```

```{r Table_AIC_BIC_gamma, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

load("./Data/results_AIC_BIC_gamma.Rdata")

combined <- rbind(results_AIC, results_BIC)
rownames(combined) <- NULL

colnames(combined) <- c("\\textbf{Criterion}", "$\\bm{n}$", "\\textbf{Normal}", "\\textbf{Lognormal}", "\\textbf{Gamma}", "\\textbf{Weibull}", "\\textbf{Inverse Gaussian}")

#combined$`Inverse Gaussian` <- c("\\multirow{2}{*}{Inverse}", "Gaussian")

# LaTeX notation 
combined[] <- lapply(combined, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.2f$", x), sprintf("$\\phantom{-}%.2f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

table_result_combined <- knitr::kable(
  combined,
  caption = "AIC and BIC values for evaluating the best distribution with CV data from Gamma SAR.",
  format = "latex",
  booktabs = TRUE,
  align = "lccccccc",
  escape = FALSE,
  digits = 2,
  label = "table_aic_gamma",
  centering = FALSE,
  table.envir = "table", position = "H") %>%
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")

print(table_result_combined)
```

```{r Simulated_AIC_BIC_GI0, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")
R <- 15000
mu <- 1
L <- 5
B <- 5
alpha1 <- -3
sample.sizes <- c(25, 49, 81, 121)

# Simulación y cálculo de resultados
results_AIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

results_BIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

for (s in sample.sizes) {
  TestStatistics <- numeric(0)

  for (r in 1:R) {
    z <- gi0_sample(mu, alpha1, L, s)
    TestStat <- sd(z) / mean(z)
    TestStatistics <- c(TestStatistics, TestStat)
  }

  fit_weibull <- fitdistr(TestStatistics, "weibull")
  fit_lognormal <- fitdistr(TestStatistics, "lognormal")
  fit_gamma <- fitdistr(TestStatistics, "gamma")
  fit_normal <- fitdistr(TestStatistics, "normal")

  AIC_values <- c(AIC(fit_weibull), AIC(fit_lognormal), AIC(fit_gamma), AIC(fit_normal))
  BIC_values <- c(BIC(fit_weibull), BIC(fit_lognormal), BIC(fit_gamma), BIC(fit_normal))

  dinvgauss <- function(x, mu, lambda) {
    sqrt(lambda / (2 * pi * x^3)) * exp(-lambda * (x - mu)^2 / (2 * mu^2 * x))
  }

  negLL_invgauss <- function(params, data) {
    mu <- params[1]
    lambda <- params[2]
    CV <- data

    -sum(log(dinvgauss(CV, mu, lambda)))
  }

  fit_invGauss <- fitdist(data = TestStatistics, distr = "invgauss", method = "mle",
                          start = list(mu = mean(TestStatistics), lambda = var(TestStatistics)))

  log_likelihood <- logLik(fit_invGauss)
  num_params <- length(coef(fit_invGauss))
  n <- length(TestStatistics)
  AIC_manual <- -2 * log_likelihood + 2 * num_params
  BIC_manual <- -2 * log_likelihood + num_params * log(n)

  results_AIC <- rbind(results_AIC, data.frame(Test = "AIC",
                                               SampleSize = s,
                                               Normal = AIC_values[4],
                                               Lognormal = AIC_values[2],
                                               Gamma = AIC_values[3],
                                               Weibull = AIC_values[1],
                                               InverseGaussian = AIC_manual))

  results_BIC <- rbind(results_BIC, data.frame(Test = "BIC",
                                               SampleSize = s,
                                               Normal = BIC_values[4],
                                               Lognormal = BIC_values[2],
                                               Gamma = BIC_values[3],
                                               Weibull = BIC_values[1],
                                               InverseGaussian = BIC_manual))
}

# Guardar los resultados en un archivo .Rdata
save(results_AIC, results_BIC, file = "./Data/results_AIC_BIC.Rdata")

```

```{r Table_AIC_BIC, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

load("./Data/results_AIC_BIC.Rdata")

combined <- rbind(results_AIC, results_BIC)
rownames(combined) <- NULL

colnames(combined) <- c("\\textbf{Criterion}", "$\\bm{n}$", "\\textbf{Normal}", "\\textbf{Lognormal}", "\\textbf{Gamma}", "\\textbf{Weibull}", "\\textbf{Inverse Gaussian}")

# Modificar el nombre de la última columna para que esté en dos líneas
#combined$`Inverse Gaussian` <- c("\\multirow{2}{*}{Inverse}", "Gaussian")

# LaTeX notation 
combined[] <- lapply(combined, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.2f$", x), sprintf("$\\phantom{-}%.2f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

table_result_combined <- knitr::kable(
  combined,
  caption = "AIC and BIC values for evaluating the best distribution with CV data from $G_I^0$.",
  format = "latex",
  booktabs = TRUE,
  align = "lccccccc",
  escape = FALSE,
  digits = 2,
  label = "table_aic",
  centering = FALSE,
  table.envir = "table", position = "H") %>%
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")

print(table_result_combined)
```

```{r Simulated_AIC_BIC_gamma_MnADmedian, echo=FALSE, message=FALSE, warning=FALSE}

set.seed(1234567890, kind = "Mersenne-Twister")
R <- 15000
mu <- 1
L <- 5
sample.sizes <- c(25, 49, 81, 121)

# madmedian <- function(x){
#   mdn <- median(x)
#   MAD <- 1.4826 * median(abs(x - mdn))
#   MAD/mdn
# }

MnADmedian <- function(x){
  mdn <- median(x)
  MnAD <- mean(abs(x - mdn))
  MnAD/mdn
}

results_AIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

results_BIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

for (s in sample.sizes) {
  TestStatistics <- numeric(0)

  for (r in 1:R) {
    z <- gamma_sar_sample(L, mu, s)
    # z <- gi0_sample(mu, alpha1, L, s)
    madmed <- MnADmedian(z)  
    TestStatistics <- c(TestStatistics, madmed)
  }

  fit_weibull <- fitdistr(TestStatistics, "weibull")
  fit_lognormal <- fitdistr(TestStatistics, "lognormal")
  fit_gamma <- fitdistr(TestStatistics, "gamma")
  fit_normal <- fitdistr(TestStatistics, "normal")

  AIC_values <- c(AIC(fit_weibull), AIC(fit_lognormal), AIC(fit_gamma), AIC(fit_normal))
  BIC_values <- c(BIC(fit_weibull), BIC(fit_lognormal), BIC(fit_gamma), BIC(fit_normal))

  dinvgauss <- function(x, mu, lambda) {
    sqrt(lambda / (2 * pi * x^3)) * exp(-lambda * (x - mu)^2 / (2 * mu^2 * x))
  }

  negLL_invgauss <- function(params, data) {
    mu <- params[1]
    lambda <- params[2]
    CV <- data

    -sum(log(dinvgauss(CV, mu, lambda)))
  }

  fit_invGauss <- fitdist(data = TestStatistics, distr = "invgauss", method = "mle",
                          start = list(mu = mean(TestStatistics), lambda = var(TestStatistics)))

  log_likelihood <- logLik(fit_invGauss)
  num_params <- length(coef(fit_invGauss))
  n <- length(TestStatistics)
  AIC_manual <- -2 * log_likelihood + 2 * num_params
  BIC_manual <- -2 * log_likelihood + num_params * log(n)

  results_AIC <- rbind(results_AIC, data.frame(Test = "AIC",
                                               SampleSize = s,
                                               Normal = AIC_values[4],
                                               Lognormal = AIC_values[2],
                                               Gamma = AIC_values[3],
                                               Weibull = AIC_values[1],
                                               InverseGaussian = AIC_manual))

  results_BIC <- rbind(results_BIC, data.frame(Test = "BIC",
                                               SampleSize = s,
                                               Normal = BIC_values[4],
                                               Lognormal = BIC_values[2],
                                               Gamma = BIC_values[3],
                                               Weibull = BIC_values[1],
                                               InverseGaussian = BIC_manual))
}

save(results_AIC, results_BIC, file = "./Data/results_AIC_BIC_gamma_MnADmedian.Rdata")

```

```{r Table_AIC_BIC_gamma_MnADmedian, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

load("./Data/results_AIC_BIC_gamma_MnADmedian.Rdata")

combined <- rbind(results_AIC, results_BIC)
rownames(combined) <- NULL

colnames(combined) <- c("\\textbf{Criterion}", "$\\bm{n}$", "\\textbf{Normal}", "\\textbf{Lognormal}", "\\textbf{Gamma}", "\\textbf{Weibull}", "\\textbf{Inverse Gaussian}")

#combined$`Inverse Gaussian` <- c("\\multirow{2}{*}{Inverse}", "Gaussian")

# LaTeX notation 
combined[] <- lapply(combined, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.2f$", x), sprintf("$\\phantom{-}%.2f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

table_result_combined <- knitr::kable(
  combined,
  caption = "AIC and BIC values for evaluating the best distribution with $\\text{CV}_{\\text{MnAD}}$ data from Gamma SAR.",
  format = "latex",
  booktabs = TRUE,
  align = "lccccccc",
  escape = FALSE,
  digits = 2,
  label = "table_aic_gamma_madmed",
  centering = FALSE,
  table.envir = "table", position = "H") %>%
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")

print(table_result_combined)
```

```{r Simulated_AIC_BIC_GI0_MnADmedian, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")
R <- 15000
mu <- 1
L <- 5
B <- 5
alpha1 <- -5
sample.sizes <- c(25, 49, 81, 121)

# madmedian <- function(x){
#   mdn <- median(x)
#   MAD <- 1.4826 * median(abs(x - mdn))
#   return(MAD/mdn)
# }

MnADmedian <- function(x){
  mdn <- median(x)
  MnAD <- mean(abs(x - mdn))
  MnAD/mdn
}
results_AIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

results_BIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

for (s in sample.sizes) {
  TestStatistics <- numeric(0)

  for (r in 1:R) {
    z <- gi0_sample(mu, alpha1, L, s)
    madmed <- MnADmedian(z) 
    TestStatistics <- c(TestStatistics, madmed)
  }

  fit_weibull <- fitdistr(TestStatistics, "weibull")
  fit_lognormal <- fitdistr(TestStatistics, "lognormal")
  fit_gamma <- fitdistr(TestStatistics, "gamma")
  fit_normal <- fitdistr(TestStatistics, "normal")

  AIC_values <- c(AIC(fit_weibull), AIC(fit_lognormal), AIC(fit_gamma), AIC(fit_normal))
  BIC_values <- c(BIC(fit_weibull), BIC(fit_lognormal), BIC(fit_gamma), BIC(fit_normal))

  dinvgauss <- function(x, mu, lambda) {
    sqrt(lambda / (2 * pi * x^3)) * exp(-lambda * (x - mu)^2 / (2 * mu^2 * x))
  }

  negLL_invgauss <- function(params, data) {
    mu <- params[1]
    lambda <- params[2]
    CV <- data

    -sum(log(dinvgauss(CV, mu, lambda)))
  }

  fit_invGauss <- fitdist(data = TestStatistics, distr = "invgauss", method = "mle",
                          start = list(mu = mean(TestStatistics), lambda = var(TestStatistics)))

  log_likelihood <- logLik(fit_invGauss)
  num_params <- length(coef(fit_invGauss))
  n <- length(TestStatistics)
  AIC_manual <- -2 * log_likelihood + 2 * num_params
  BIC_manual <- -2 * log_likelihood + num_params * log(n)

  results_AIC <- rbind(results_AIC, data.frame(Test = "AIC",
                                               SampleSize = s,
                                               Normal = AIC_values[4],
                                               Lognormal = AIC_values[2],
                                               Gamma = AIC_values[3],
                                               Weibull = AIC_values[1],
                                               InverseGaussian = AIC_manual))

  results_BIC <- rbind(results_BIC, data.frame(Test = "BIC",
                                               SampleSize = s,
                                               Normal = BIC_values[4],
                                               Lognormal = BIC_values[2],
                                               Gamma = BIC_values[3],
                                               Weibull = BIC_values[1],
                                               InverseGaussian = BIC_manual))
}

save(results_AIC, results_BIC, file = "./Data/results_AIC_BIC_GI0_MnADmedian.Rdata")

```

```{r Table_AIC_BIC_GI0_MnADmedian, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

load("./Data/results_AIC_BIC_GI0_MnADmedian.Rdata")

combined <- rbind(results_AIC, results_BIC)
rownames(combined) <- NULL

colnames(combined) <- c("\\textbf{Criterion}", "$\\bm{n}$", "\\textbf{Normal}", "\\textbf{Lognormal}", "\\textbf{Gamma}", "\\textbf{Weibull}", "\\textbf{Inverse Gaussian}")

#combined$`Inverse Gaussian` <- c("\\multirow{2}{*}{Inverse}", "Gaussian")

# LaTeX notation 
combined[] <- lapply(combined, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.2f$", x), sprintf("$\\phantom{-}%.2f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

table_result_combined <- knitr::kable(
  combined,
  caption = "AIC and BIC values for evaluating the best distribution with $\\text{CV}_{\\text{MnAD}}$ data from $G_I^0$.",
  format = "latex",
  booktabs = TRUE,
  align = "lccccccc",
  escape = FALSE,
  digits = 2,
  label = "table_aic_gio_MnADmedian",
  centering = FALSE,
  table.envir = "table", position = "H") %>%
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")

print(table_result_combined)
```

```{r Simulated_cv_gamma, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

sample.size <- c(49)
R <- 15000
mu <- 1
L <- 5
B <- 5
alpha1 <- -2

TestStatistics <- numeric(0)

for (s in sample.size) {
  for (r in 1:R) {
    z <- gamma_sar_sample(L, mu, s)
    #z <- gi0_sample(mu, alpha1, L, s)
    TestStat <- sd(z) / mean(z)
    TestStatistics <- c(TestStatistics, TestStat)  
  }
}

save(TestStatistics, file = "./Data/rTestStatistics_gamma_49.Rdata")
```

```{r Plot_cv_gamma, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H',  fig.cap="Goodness of fit plots for evaluating the best distribution with CV data from Gamma SAR, with  $n=49$, $L=5$ and $\\mu=1$.", fig.width=8, fig.height=6}

load("./Data/rTestStatistics_gamma_49.Rdata")

# Ajustar las distribuciones a los datos
fln <- fitdist(TestStatistics, "lnorm")
fg <- fitdist(TestStatistics, "gamma")
fn <- fitdist(TestStatistics, "norm")  # Ajuste para la distribución normal
fw <- fitdist(TestStatistics, "weibull")

# Graficar las comparaciones
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
plot.legend <- c( "Lognormal", "Gamma", "Normal", "Weibull") 
denscomp(list( fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlab = "CV")
qqcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)
cdfcomp(list(fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlab = "CV")
ppcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)

```

```{r Simulated_cv_gi0, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

sample.size <- c(49)
R <- 15000
mu <- 1
L <- 5
B <- 5
alpha1 <- -2

TestStatistics <- numeric(0)

for (s in sample.size) {
  for (r in 1:R) {
    z <- gi0_sample(mu, alpha1, L, s)
    TestStat <- sd(z) / mean(z)
    TestStatistics <- c(TestStatistics, TestStat)  
  }
}

save(TestStatistics, file = "./Data/rTestStatistics_gi0_49.Rdata")
```

```{r Plot_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H',  fig.cap="Goodness of fit plots for evaluating the best distribution with $\\text{CV}_{\\text{MnAD}}$ data from $G_I^0$, with  $n=49$, $L=5$, $\\mu=1$ and $\\alpha=-2$.", fig.width=8, fig.height=6}

load("./Data/rTestStatistics_gi0_49.Rdata")

#
fln <- fitdist(TestStatistics, "lnorm")
fg <- fitdist(TestStatistics, "gamma")
fn <- fitdist(TestStatistics, "norm")  
fw <- fitdist(TestStatistics, "weibull")

# 
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
plot.legend <- c( "Lognormal", "Gamma", "Normal", "Weibull") 
denscomp(list( fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlab = "CV")
qqcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)
cdfcomp(list(fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlab = "CV")
ppcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)

```

```{r Simulated_MnADmedian_gamma, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

sample.size <- c(49)
R <- 15000
mu <- 1
L <- 5
B <- 5
alpha1 <- -2

MnADmedian <- function(x){
  mdn <- median(x)
  MnAD <- mean(abs(x - mdn))
  MnAD/mdn
}

TestStatistics <- numeric(0)

for (s in sample.size) {
  for (r in 1:R) {
    z <- gamma_sar_sample(L, mu, s)
    #z <- gi0_sample(mu, alpha1, L, s)
    TestStat <- MnADmedian(z)
    TestStatistics <- c(TestStatistics, TestStat)  
  }
}

save(TestStatistics, file = "./Data/rTestStatistics_gamma_49_MnADmedian.Rdata")
```

```{r Plot_MnADmedian_gamma, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H',  fig.cap="Goodness of fit plots for evaluating the best distribution with $\\text{CV}_{\\text{MnAD}}$ data from Gamma SAR, with  $n=49$, $L=5$ and $\\mu=1$.", fig.width=8, fig.height=6}

load("./Data/rTestStatistics_gamma_49_MnADmedian.Rdata")

# Ajustar las distribuciones a los datos
fln <- fitdist(TestStatistics, "lnorm")
fg <- fitdist(TestStatistics, "gamma")
fn <- fitdist(TestStatistics, "norm")  # Ajuste para la distribución normal
fw <- fitdist(TestStatistics, "weibull")

# Graficar las comparaciones
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
plot.legend <- c( "Lognormal", "Gamma", "Normal", "Weibull") 
denscomp(list( fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlab = "CV")
qqcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)
cdfcomp(list(fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlab = "CV")
ppcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)

```

```{r Simulated_MnADmedian_gi0, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

sample.size <- c(49)
R <- 12000
mu <- 1
L <- 5
#B <- 5
alpha1 <- -5

MnADmedian <- function(x){
  mdn <- median(x)
  MnAD <- mean(abs(x - mdn))
  MnAD/mdn
}

TestStatistics_mad <- numeric(0)

for (s in sample.size) {
  for (r in 1:R) {
    z <- gi0_sample(mu, alpha1, L, s)
    madmed <- MnADmedian(z)
    TestStatistics_mad <- c(TestStatistics_mad, madmed)  
  }
}

save(TestStatistics_mad, file = "./Data/rTestStatistics_MnADmedian_gi0_49.Rdata")
```

```{r Plot_madmed_gi0_MnADmedian, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.pos="hbt",  fig.cap="Goodness of fit plots for evaluating the best distribution with $CV_{\\text{MnAD}}$ data from $G_I^0$, with  $n=49$, $L=5$, $\\mu=1$ and $\\alpha=-5$.", fig.width=8, fig.height=6}

load("./Data/rTestStatistics_MnADmedian_gi0_49.Rdata")

#
fln <- fitdist(TestStatistics_mad, "lnorm")
fg <- fitdist(TestStatistics_mad, "gamma")
fn <- fitdist(TestStatistics_mad, "norm")  
fw <- fitdist(TestStatistics_mad, "weibull")

# 
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
plot.legend <- c( "Lognormal", "Gamma", "Normal", "Weibull") 
denscomp(list( fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlab = "CV")
qqcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)
cdfcomp(list(fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlab = "CV")
ppcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","#FDA75F"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)

```

# Simulation Results {#sec:Results}

This section presents simulations performed to obtain an assessment of the performances of the proposed test statistics.

## Simulated Data

Figure 1a shows the phantom used for a simulated image with dimensions of $250\times 250$ pixels to illustrate the differences between heterogeneous (dark, $\alpha=-2$) and homogeneous (light, $\alpha=-10$) regions, with $L = 5$. 
The areas are formed by independent draws from the $G^0_I$ distribution.

The three test statistics are applied to the simulated image using local sliding windows of size $7\times 7$.

The resulting $p$-values for the tests $S_H$, $T_{\text{CV}}$, and $T_{\text{CV}_{\text{MnAD}}}$ are shown in Figures&nbsp;\ref{fig:sim_SAR_Images}(b)-(d) and Figures&nbsp;\ref{fig:sim_SAR_Images_p05}(a)-(c), respectively. 
The new images reveal that a darker background indicates homogeneity, while lighter backgrounds represent heterogeneous areas.

```{r sim_SAR_Images, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results of applying the test statistics. (a) Simulated image; $P$-values from the entropy-based test; (c) $P$-values from the CV test; (d) $P$-values from the   $\\text{CV}_{\\text{MnAD}}$ test; ',  fig.subcap=c('', '', '', ''), out.width = c("40mm", "40mm","40mm", "40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/simulated7x7.png")
knitr::include_graphics("../Figures/PNG/sim_pvalue7x7_total_H.png")
knitr::include_graphics("../Figures/PNG/sim_pvalue7x7_total_cv.png")
knitr::include_graphics("../Figures/PNG/sim_pvalue7x7_total_mnad.png")
```
```{r sim_SAR_Images_p05, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results of tests on simulated data with a threshold of $0.1$ of the $p$-value. (a) Entropy-based test; (b) CV test; (c) $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm","40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/sim_pvalue7x7_H.png")
knitr::include_graphics("../Figures/PNG/sim_pvalue7x7_cv.png")
knitr::include_graphics("../Figures/PNG/sim_pvalue7x7_mnad.png")

```

## Real Data

We applied the proposed test statistics in actual SAR images acquired over cities, including Flevoland, Netherlands and Ottawa, Canada.   
These images were captured by the Sentinel-1A satellite operating in the C band,  with  VV polarization, intensity format, and $L=5$ nominal looks.
Both images have dimensions of $512 \times 512$ pixels and contain urban areas, water forest regions, as shown in  Figures&nbsp;\ref{fig:real_SAR_Images_coe}(a)-(b).

The $p$-values obtained for the $S_H$, $T_{\text{CV}}$, and $T_{\text{CV}_{\text{MnAD}}}$ tests are presented in Figures&nbsp;\ref{fig:sim_SAR_Images_Frankfurt}-\ref{fig:sim_SAR_Images_01_Ottawa}, respectively.

```{r real_SAR_Images_coe, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='SAR images. (a) Frankfurt; (b) Ottawa. ',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm"),  fig.ncol=2, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/Intensity_Frankfurt_512_n.png")
knitr::include_graphics("../Figures/PNG/Intensity_Ottawa_512_n.png")
#knitr::include_graphics("../Figures/PNG/sim_pvalue7x7_mnad.png")

```
```{r sim_SAR_Images_Frankfurt, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results of applying the test statistics, Frankfurt. (a) $P$-values from the entropy-based test; (b) $P$-values from the CV test; (c) $P$-values from the   $\\text{CV}_{\\text{MnAD}}$ test; ',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm","40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/Frankfurt_512_pvalue7x7_total_H.png")
knitr::include_graphics("../Figures/PNG/Frankfurt_512_pvalue7x7_total_CV.png")
knitr::include_graphics("../Figures/PNG/Frankfurt_512_mnad_7_total.png")

```
```{r sim_SAR_Images_01Frankfurt, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results of tests on simulated data with a threshold of $0.1$ of the $p$-value, Frankfurt. (a) Entropy-based test; (b) CV test; (c) $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm","40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/Frankfurt_512_pvalue7x7_H_01.png")
knitr::include_graphics("../Figures/PNG/Frankfurt_512_pvalue7x7_CV.png")
knitr::include_graphics("../Figures/PNG/Frankfurt_512_mnad_7_0.1.png")

```

```{r sim_SAR_Images_Ottawa, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results of applying the test statistics, Ottawa. (a) $P$-values from the entropy-based test; (b) $P$-values from the CV test; (c) $P$-values from the   $\\text{CV}_{\\text{MnAD}}$ test; ',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm","40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/Ottawa_512_pvalue7x7_total_H.png")
knitr::include_graphics("../Figures/PNG/Ottawa_512_CV_pvalue7x7_total_CV.png")
knitr::include_graphics("../Figures/PNG/Ottawa_512_mnad_7_total.png")

```
```{r sim_SAR_Images_01_Ottawa, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results of tests on simulated data with a threshold of $0.1$ of the $p$-value, Ottawa. (a) Entropy-based test; (b) CV test; (c) $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm","40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/Ottawa_512_pvalue7x7_H_01.png")
knitr::include_graphics("../Figures/PNG/Ottawa_512_CV_pvalue7x7_CV.png")
knitr::include_graphics("../Figures/PNG/Ottawa_512_mnad_7_0.1.png")

```

# Conclusion {#sec:conclusion}

In this article,...