### 2024-04-04 Text/Identifying-Heterogeneity-in-SAR-Data-with-New-Test-Statistics.tex

%  LaTeX support: latex@mdpi.com
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
% pandoc conditionals added to preserve backwards compatibility with previous versions of rticles

\documentclass[remotesensing,article,submit,moreauthors,pdftex]{Definitions/mdpi}

%% Some pieces required from the pandoc template
\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}

%--------------------
% Class Options:
%--------------------

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. Remove "pdftex" for (1) compiling with LaTeX & dvi2pdf (if eps figures are used) or for (2) compiling with XeLaTeX.

%=================================================================
% MDPI internal commands - do not modify
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2023}
\copyrightyear{2023}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{ }
\daterevised{ } % Comment out if no revised date
\dateaccepted{ }
\datepublished{ }
%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%\pdfoutput=1 % Uncommented for upload to arXiv.org

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, float, amsmath, amssymb, lineno, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, colortbl, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, array, tabularx, pbox, ragged2e, tocloft, marginnote, marginfix, enotez, amsthm, natbib, hyperref, cleveref, scrextend, url, geometry, newfloat, caption, draftwatermark, seqsplit
% cleveref: load \crefname definitions after \begin{document}

%=================================================================
% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Identifying Heterogeneity in SAR Data with New Test Statistics}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Identifying Heterogeneity in SAR Data with New Test
Statistics}

% Author Orchid ID: enter ID or remove command
%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Alejandro C.
Frery$^{1,*}$\href{https://orcid.org/0000-0002-8002-5341}
{\orcidicon}, Janeth
Alpala$^{2}$\href{https://orcid.org/0000-0002-0265-6236}
{\orcidicon}, Abraão D. C.
Nascimento$^{2}$\href{https://orcid.org/0000-0003-2673-219X}
{\orcidicon}}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Alejandro C. Frery, Janeth Alpala, Abraão D. C. Nascimento}

% MDPI internal command: Authors, for citation in the left column
%\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
\AuthorCitation{Frery, A.C.; Alpala, J.; Nascimento,A.D.C.}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad School of Mathematics and Statistics, Victoria University
of Wellington, Wellington 6140, New Zealand; \\
$^{2}$ \quad Departamento de Estatística, Universidade Federal de
Pernambuco, Recife 50670-901, PE, Brazil; \\
}

% Contact information of the corresponding author
\corres{Correspondence: \href{mailto:alejandro.frery@vuw.ac.nz}{\nolinkurl{alejandro.frery@vuw.ac.nz}}}

% Current address and/or shared authorship

% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{This paper presents a statistical approach to identify the
underlying roughness characteristics in synthetic aperture radar (SAR)
intensity data. The physical modeling of this kind of data allows to use
the Gamma distribution in the presence of fully-developed speckle, i.e.,
when there are infinitely many independent backscatterers per resolution
cell. This situation is often called ``homogeneous'' or ``textureless''
areas. The \(\mathcal{G}_I^0\) distribution is also a widely accepted
law for heterogeneous and extremely heterogeneous areas. We propose
three test statistics for discriminating between homogeneous and
non-homogeneous areas, i.e., between Gamma and GI0 distributed data,
both with a known number of looks. The first test statistic utilizes a
bootstrapped non-parametric estimator of the Shannon entropy, offering a
robust evaluation in the face of uncertain distributional assumptions.
The second test uses the classic coefficient of variation. The third
test employs an alternative form of estimating the CV based on the ratio
of the mean absolute deviation from the median to the median. We apply
our test statistic to produce maps of \(p\)-values for the hypothesis of
homogeneity. These maps identify different types of targets. We show
that our proposal outperforms its competitors using simulated and SAR
data.}

% Keywords
\keyword{SAR; heterogeneity; entropy; coefficient of variation;
hypothesis tests}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Advances in Respiratory Medicine
%\addhighlights{yes}
%\renewcommand{\addhighlights}{%

%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars. Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it. Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
%\textbf{What are the main findings?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}\vspace{3pt}
%\textbf{What is the implication of the main finding?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage[english]{babel}
\usepackage{bm,bbm}
\usepackage{mathrsfs}
\usepackage{siunitx}
\usepackage{graphicx}
\usepackage{url}
\usepackage[T1]{fontenc}
\usepackage{polski}
\usepackage{booktabs}
\usepackage{color}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{subcaption}
\captionsetup[sub]{labelfont=bf,textfont=bf}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\setlength{\tabcolsep}{6pt} 
\newcommand{\bias}{\operatorname{Bias}}
\newcommand{\widebar}[1]{\overline{#1}}

\hypertarget{sec:Introduction}{%
\section{Introduction}\label{sec:Introduction}}

Synthetic Aperture Radar (SAR) technology has become an important tool
for environmental monitoring and disaster management, providing valuable
imagery in different conditions, including day or night and diverse
weather situations~\cite{Moreira2013,Mu2019}. However, the effective
utilization of SAR data depends on a thorough understanding of its
statistical properties, especially the impact of speckle effects, which
are inherent in SAR data due to the coherent nature of the imaging
process~\cite{Argenti2013}.

In the presence of non-Gaussian speckle interference, SAR data requires
a reliable statistical model for accurate processing. The
\(\mathcal{G}^0\) distribution, suitable for SAR data, includes the
Gamma law as a limit case for fully developed
speckle~\cite{Ferreira2020}, providing flexibility with fewer parameters
for analysis.

Our work is motivated to improve the identification of potential
roughness features in SAR intensity data. Physical modeling of SAR data
allows to take advantage of the gamma distribution in the presence of
fully developed speckle, creating a scenario with an infinite number of
independent backscatterers per unit of resolution, often referred to as
homogeneous areas.

In this context, we present a set of three novel test statistics
designed to discriminate between homogeneous and non-homogeneous areas,
specifically aimed at differentiating between Gamma and
\(\mathcal{G}^0\) distributed data, each with a known number of
appearances. Hence, to understand the core ideas behind our approach, we
explore into the use of properties such as entropy and coefficient of
variation.

The entropy is a fundamental concept in information theory with broad
applications to pattern recognition, statistical physics, stochastic
dynamics, and statistics. Shannon introduced it for a random variable in
1948~\cite{Shannon1948} as a measure of information and uncertainty.\\
In statistics, Shannon entropy is a crucial descriptive parameter,
particularly for assessing data dispersion and conducting tests for
normality, exponentiality, and
uniformity~\cite{Wieczorkowski1999,Zamanzade2012}. Entropy estimation
involves several practical difficulties, especially when the model is
unknown; in these cases, non-parametric methods are used.

Spacing methods have been discussed as a non-parametric approach
in~\cite{AlizadehNoughabi2010,Subhash2021}. This strategy offers
flexibility to address a wide range of models without imposing specific
parametric restrictions.

In our study, entropy serves as the fundamental parameter, providing a
robust assessment by means of a bootstrapped non-parametric estimator of
Shannon entropy.

Simultaneously, we use the classic coefficient of variation (CV) and an
alternative form based on the ratio between the median absolute
deviation and the median. The CV is known as an efficient and robust
index of textural information, measures the homogeneity of the
image~\cite{Lopes1990}.

The CV, introduced by Pearson~\cite{Pearson1896} in 1896, is a relative
measure of dispersion widely used in various areas of applied statistics
including sampling, biostatistics, medical and biological research,
climatology, and other fields
\cite{hendricks1936sampling, Tian2005,SubrahmanyaNairy2003,Chankham2024}.
It facilitates the comparison of variability between different
populations and is especially valuable for comparing variables with
different units. This is because when the main purpose is to compare
variations of several variables, the standard deviation cannot serve as
an adequate measure of variation unless all variables are expressed in
the same unit of measurement and have identical mean values. When these
requirements are not hold for the variables under study, CV is the
relative measure typically used in real applications. The variable with
the highest CV value is the one with the highest relative dispersion
around the mean~\cite{Banik2011}.

Applying these test statistics to generate \(p\)-value maps facilitates
testing of the homogeneity hypothesis, revealing diverse types of
targets within the SAR data. Through extensive experimentation involving
real and simulated SAR data, we demonstrate the superiority of our
proposed methodology over existing approaches.

The article is structured as follows: Section~\ref{sec:Background}
covers statistical modeling and entropy estimation for Intensity SAR
data. Section~\ref{sec:test} outlines hypothesis testing based on
non-parametric entropy estimators and and coefficient of variation. In
Section~\ref{sec:Results}, we present experimental results. Finally, in
Section~\ref{sec:conclusion} conclusions are exhibited.

\hypertarget{sec:Background}{%
\section{Background}\label{sec:Background}}

\hypertarget{statistical-modeling-of-intensity-sar-data}{%
\subsection{Statistical Modeling of Intensity SAR
data}\label{statistical-modeling-of-intensity-sar-data}}

The primary models used for intensity SAR data include the Gamma and
\(\mathcal{G}_I^0\) distributions~\citep{Frery1997}. The first is
suitable for fully developed speckle and is a limiting case of the
second, which is appealing due to its versatility in accurately
representing regions with various roughness
characteristics~\citep{Cassetti2022}. We denote
\(Z \sim \Gamma_{\text{SAR}}(L, \mu)\) and
\(Z \sim G_I^0(\alpha, \gamma, L)\) to indicate that \(Z\) follows the
distributions characterized by the respective probability density
functions: \begin{align}
    f_Z(z;L, \mu)&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gamma1}\\
    f_Z(z; \alpha, \gamma, L)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\cdot\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gi01}
\end{align} where, in~\eqref{E:gamma1} \(\mu > 0\) is the mean;
in~\eqref{E:gi01} \(\gamma > 0\) is the scale, \(\alpha < -1\) measures
the roughness, \(L \geq 1\) is the number of looks, \(\Gamma(\cdot)\) is
the gamma function, and \(\mathbbm 1_{A}(z)\) is the indicator function
of the set \(A\).

From \eqref{E:gi01}, the \(r\)th moment of \(Z\) is expressed as:
\begin{align}
    E_{G_I^0}\left(Z^r\right)=\left(\frac{\gamma}{L}\right)^r\frac{\Gamma(-\alpha-r)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+r)}{\gamma(L)}, \quad \alpha <-r. 
    \label{E:rmom}
\end{align}

Even though the \(\mathcal{G}_I^0\) distribution is defined by the
parameters \(\alpha\) and \(\gamma\), SAR literature commonly utilizes
the texture \(\alpha\) and the mean \(\mu\)~\cite{Nascimento2010}. In
this way, we compute the expected value \(\mu\) using the expression
in~\eqref{E:rmom}, and we reparametrize~\eqref{E:gi01} using \(\mu\),
\(\alpha\), and \(L\). Then \begin{align*}
    \mu=\left(\frac{\gamma}{L}\right)\frac{\Gamma(-\alpha-1)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+1)}{\gamma(L)}=-\frac{\gamma}{\alpha+1}.
\end{align*} Thus, the probability density function that characterize
the \(G_I^0(\mu, \alpha, L)\) law is \begin{equation}
        f_Z(z; \mu, \alpha, L)=\frac{L^L\Gamma(L-\alpha)}{\big(-\mu(\alpha+1)\big)^{\alpha}\Gamma(-\alpha)\Gamma(L)} \frac{z^{L-1}}{\big(-\mu(\alpha+1)+Lz\big)^{L-\alpha}}.\label{E:gi02}
\end{equation}

\hypertarget{the-shannon-entropy}{%
\subsection{The Shannon Entropy}\label{the-shannon-entropy}}

The parametric representation of Shannon entropy for a system described
by a continuous random variable is: \begin{equation}
  \label{E:entropy2}
  H(Z)=-\int_{-\infty }^\infty \ f(z)\ln f(z)\, \mathrm{d}z,
\end{equation} here, \(f(\cdot)\) is the probability density function
that characterizes the distribution of the real-valued random variable
\(Z\).

Using~\eqref{E:entropy2}, we can express the Shannon entropy of
\(\Gamma_{\text{SAR}}\)in~\eqref{E:gamma1} and \(G_I^0\)
in~\eqref{E:gi02} based on~\cite{Cassetti2022} and~\cite{Ferreira2020}:
\begin{equation}
\label{E:E-gamma}
H_{\Gamma_{\text{SAR}}}(L, \mu) =   L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) + \ln \mu, 
\end{equation} \begin{multline}
\label{E:E-GIO}
H_{G_I^0}(\mu, \alpha, L) =L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) +\ln \mu -\ln\Gamma(L-\alpha)\\
+ (L-\alpha) \psi^{(0)}(L-\alpha)-(1-\alpha)\psi^{(0)}(-\alpha)+\ln (-1-\alpha)+\ln\Gamma(-\alpha)-L,
\end{multline} where \(\psi^{(0)}(\cdot)\) is the digamma function.
Figure~\ref{fig:Plot_GI0_to_gamma1}, shows the behaviour of the entropy
of \(G_I^0\) when when
\(\alpha \in \left\{-\infty, -20, -8, -3\right\}\), observing its
convergence towards the entropy of \(\Gamma_{\text{SAR}}\) as \(\alpha\)
takes large negative values.

\begin{figure}[hbt]
\includegraphics[width=0.7\linewidth]{Identifying-Heterogeneity-in-SAR-Data-with-New-Test-Statistics_files/figure-latex/Plot_GI0_to_gamma1-1} \caption{$H_{ G_I^0}$ converges to the $H_{\Gamma_{\text{SAR}}}$ with $L=8$.}\label{fig:Plot_GI0_to_gamma1}
\end{figure}

\hypertarget{estimation-of-the-shannon-entropy}{%
\subsection{Estimation of the Shannon
Entropy}\label{estimation-of-the-shannon-entropy}}

The problem of non-parametric estimating of \(H(Z)\) has been considered
by many authors
including~\cite{vasicek1976test, Wieczorkowski1999,correa1995new,AlOmari2019},
who proposed estimators based on spacings.

One of the earliest non-parametric estimators relying on spacings was
introduced by Vasicek~\cite{vasicek1976test}. Assuming that
\(\bm{Z}=(Z_1, Z_2,\ldots,Z_n)\) is a random sample from the
distribution \(F(z)\), the estimator is defined as: \begin{equation*}
\label{E:Vas}
    \widehat{H}_{\text{V}}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\ln\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
    \end{equation*} where \(m<n/2\) is a positive integer,
\(Z_{(i+m)}-Z_{(i-m)}\) is the \(m\)-spacing and
\(Z_{(1)}\leq Z_{(2)}\leq\ldots\leq Z_{(n)}\) are the order statistics
and \(Z_{(i)}= Z_{(1)}\) if \(i<1\), \(Z_{(i)}= Z_{(n)}\) if \(i>n\).

Several authors have explored adaptations to Vasicek's estimator. We
consider three estimators known for their superior
performance~\cite{Cassetti2022}:

\begin{itemize}
\item
  Correa~\cite{correa1995new}: \(\widehat{H}_{\text{C}}\).
\item
  Ebrahimi et al~\cite{Ebrahimi1994}: \(\widehat{H}_{\mathrm{E}}\).
\item
  Al Omary~\cite{IbrahimAlOmari2014}: \(\widehat{H}_{\mathrm{AO}}\).
\end{itemize}

\hypertarget{enhanced-bootstrap-technique}{%
\subsection{Enhanced Bootstrap
Technique}\label{enhanced-bootstrap-technique}}

We employ the bootstrap technique to refine the precision of existing
non-parametric entropy estimators. This approach involves generating new
datasets through resampling with repetition from an existing
one~\cite{Michelucci2021}.

Let's assume that non-parametric entropy estimators
\(\widehat{H}=\widehat{\theta}(\bm{Z})\) are inherently biased, that is:
\begin{equation}
\label{Eq:bias1}
\operatorname{Bias}\big(\widehat{\theta}(\bm{Z})\big) = E\big[\widehat{\theta}(\bm{Z})\big] - \theta \neq 0.
\end{equation} Our bootstrap-improved estimators are of the form:
\begin{align*}
\widetilde{H} &= 2\widehat{\theta}(\bm{Z}) - \frac{1}{B}\sum_{b=1}^B \widehat{\theta}_b(\bm{Z}^{(b)}),
\end{align*} where \(B\) is the number of replications in the bootstrap
technique. Applying this methodology, the original estimators by Correa,
Ebrahimi, and Al-Omari are now denoted as the proposed
bootstrap-enhanced versions: \(\widetilde{H}_{\text{C}}\),
\(\widetilde{H}_{\text{E}}\), and \(\widetilde{H}_{\text{AO}}\),
respectively.

We analyzed the performance of these estimators with a Monte Carlo
study: \(500\) samples of \(G_I^0\) of size
\(n\in\left\{9, 25, 49, 81, 121\right\}\) and several values of the
mean, number of looks and roughness. We present the results with
\(mu\in\left\{1, 10\right\}\), \(\alpha=-10\), and \(L=5\). The results
are consistent for the other situations. In the case of the bootstrap
technique, each sample is replicated \(100\) times with replacement. We
choose to use the following heuristic formula for spacing,
\(m=\left[\sqrt{n}+0.5\right]\).

In Figure~\ref{fig:Plot_bias_mse_gi0} we depict comparisons of bias and
mean squared error (MSE) between the original non-parametric entropy
estimators and their respective bootstrap-enhanced versions. The use of
the bootstrap technique exhibits more precision, reduced bias and MSE,
and improved convergence. The results of simulation are exhibited in
Table~\ref{tab:table2}.

The precision of estimators, as evidenced by bias and MSE comparisons,
benefits significantly from the bootstrap technique, particularly for
sample sizes below 81.

\begin{figure}[hbt]
\includegraphics[width=0.8\linewidth]{Identifying-Heterogeneity-in-SAR-Data-with-New-Test-Statistics_files/figure-latex/Plot_bias_mse_gi0-1} \caption{Bias and MSE, $L=5$ and  $\alpha=-10$.}\label{fig:Plot_bias_mse_gi0}
\end{figure}

\begin{table}[H]
\centering
\caption{\label{tab:table2}Bias and MSE of non-parametric entropy estimators and bootstrap versions.}
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}[t]{lrrrrrrrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{6}{c}{Bias} & \multicolumn{6}{c}{MSE} \\
\cmidrule(l{3pt}r{3pt}){3-8} \cmidrule(l{3pt}r{3pt}){9-14}
\multicolumn{1}{c}{$\mu$} & \multicolumn{1}{c}{$n$} & \multicolumn{1}{c}{$\widehat{H}_{\text{AO}}$} & \multicolumn{1}{c}{$\widehat{H}_{\text{C}}$} & \multicolumn{1}{c}{$\widehat{H}_{\text{E}}$} & \multicolumn{1}{c}{$\widetilde{H}_{\text{AO}}$} & \multicolumn{1}{c}{$\widetilde{H}_{\text{C}}$} & \multicolumn{1}{c}{$\widetilde{H}_{\text{E}}$} & \multicolumn{1}{c}{$\widehat{H}_{\text{AO}}$} & \multicolumn{1}{c}{$\widehat{H}_{\text{C}}$} & \multicolumn{1}{c}{$\widehat{H}_{\text{E}}$} & \multicolumn{1}{c}{$\widetilde{H}_{\text{AO}}$} & \multicolumn{1}{c}{$\widetilde{H}_{\text{C}}$} & \multicolumn{1}{c}{$\widetilde{H}_{\text{E}}$}\\
\midrule
 & $9$ & $-0.188$ & $-0.261$ & $-0.358$ & $-0.083$ & $-0.030$ & $-0.114$ & $\phantom{-}0.131$ & $\phantom{-}0.157$ & $\phantom{-}0.217$ & $\phantom{-}0.140$ & $\phantom{-}0.124$ & $\phantom{-}0.140$\\

 & $25$ & $-0.081$ & $-0.120$ & $-0.157$ & $-0.042$ & $\phantom{-}0.001$ & $-0.049$ & $\phantom{-}0.040$ & $\phantom{-}0.046$ & $\phantom{-}0.057$ & $\phantom{-}0.042$ & $\phantom{-}0.044$ & $\phantom{-}0.044$\\

 & $49$ & $-0.032$ & $-0.059$ & $-0.080$ & $-0.007$ & $\phantom{-}0.024$ & $-0.005$ & $\phantom{-}0.018$ & $\phantom{-}0.021$ & $\phantom{-}0.023$ & $\phantom{-}0.022$ & $\phantom{-}0.021$ & $\phantom{-}0.022$\\

 & $81$ & $-0.007$ & $-0.024$ & $-0.038$ & $\phantom{-}0.004$ & $\phantom{-}0.033$ & $\phantom{-}0.025$ & $\phantom{-}0.010$ & $\phantom{-}0.010$ & $\phantom{-}0.011$ & $\phantom{-}0.011$ & $\phantom{-}0.014$ & $\phantom{-}0.013$\\

\multirow{-5}{*}[2\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash 1} & $121$ & $\phantom{-}0.000$ & $-0.010$ & $-0.020$ & $\phantom{-}0.007$ & $\phantom{-}0.037$ & $\phantom{-}0.025$ & $\phantom{-}0.006$ & $\phantom{-}0.006$ & $\phantom{-}0.006$ & $\phantom{-}0.008$ & $\phantom{-}0.009$ & $\phantom{-}0.008$\\
\cmidrule{1-14}
 & $9$ & $-0.239$ & $-0.322$ & $-0.418$ & $-0.116$ & $-0.084$ & $-0.180$ & $\phantom{-}0.169$ & $\phantom{-}0.218$ & $\phantom{-}0.289$ & $\phantom{-}0.164$ & $\phantom{-}0.172$ & $\phantom{-}0.176$\\

 & $25$ & $-0.062$ & $-0.108$ & $-0.145$ & $-0.007$ & $\phantom{-}0.006$ & $-0.026$ & $\phantom{-}0.040$ & $\phantom{-}0.048$ & $\phantom{-}0.057$ & $\phantom{-}0.051$ & $\phantom{-}0.048$ & $\phantom{-}0.047$\\

 & $49$ & $-0.018$ & $-0.044$ & $-0.064$ & $-0.001$ & $\phantom{-}0.033$ & $\phantom{-}0.017$ & $\phantom{-}0.016$ & $\phantom{-}0.017$ & $\phantom{-}0.019$ & $\phantom{-}0.018$ & $\phantom{-}0.020$ & $\phantom{-}0.018$\\

 & $81$ & $-0.015$ & $-0.032$ & $-0.046$ & $-0.006$ & $\phantom{-}0.025$ & $\phantom{-}0.015$ & $\phantom{-}0.010$ & $\phantom{-}0.010$ & $\phantom{-}0.011$ & $\phantom{-}0.011$ & $\phantom{-}0.013$ & $\phantom{-}0.011$\\

\multirow{-5}{*}[2\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash 10} & $121$ & $-0.009$ & $-0.022$ & $-0.031$ & $-0.005$ & $\phantom{-}0.025$ & $\phantom{-}0.016$ & $\phantom{-}0.007$ & $\phantom{-}0.008$ & $\phantom{-}0.008$ & $\phantom{-}0.009$ & $\phantom{-}0.010$ & $\phantom{-}0.009$\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{coefficient-of-variation-and-a-robust-alternative}{%
\subsection{Coefficient of variation and a robust
alternative}\label{coefficient-of-variation-and-a-robust-alternative}}

The population CV is defined as a ratio of the population standard
deviation \((\sigma)\) to the population mean \((\mu)\):

\begin{align}
    \text{CV}=\frac{\sigma}{\mu}, \quad \mu \neq 0.
\end{align}

We explore a robust alternative to the CV as discussed in
\cite{Ospina2019}, involving the ratio of the mean absolute deviation
from the median (MnAD) to the median, two well-known robust measures of
scale and location, respectively. The MnAD is defined as
\(n^{-1}\sum_{i=1}^n|x_i-\widehat{Q}_2|\), where \(\widehat{Q}_2\)
represents an estimate of the population median, which can be
approximated by the sample median.

\hypertarget{sec:test}{%
\section{Hypothesis Testing}\label{sec:test}}

We aim at testing the following hypotheses:

\[
 \begin{cases}\mathcal{H}_0: \text{ The data comes from the } \Gamma_{\text{SAR}}\\ 
  \mathcal{H}_1:\text{ The data comes from the } G_I^0.\end{cases}
\] In other words, we verify the hypothesis that the data are
fully-developed speckle. The proposed tests are described below.

\hypertarget{the-proposed-test-based-on-non-parametric-entropy}{%
\subsection{The Proposed Test Based on Non-parametric
Entropy}\label{the-proposed-test-based-on-non-parametric-entropy}}

The first test statistic assess the behavior of the data under the null
hypothesis, through the empirical distribution. \begin{equation}
\label{Eq:test_e}
S_H= \widetilde{H}-\left[H_{\Gamma_{\text{SAR}}}(L)+\ln \overline{Z}\right].
\end{equation} The values should be around zero under the null
hypothesis and far otherwise.

Figure~\ref{fig:Plot_density} depicts empirical test statistic densities
for various sample sizes, and Table~\ref{tab:table_stat_combined}
summarizes key statistics, including mean, SD, variance (Var), skewness
(SK), excess kurtosis (EK), and Anderson--Darling \(p\)-values for
normality. Results with \(p\)-values larger than \(0.05\) indicate no
evidence against normality. Small variance and near-zero skewness and
excess kurtosis suggest limited spread, symmetry, and light tails.\\
Normal QQ-plots, confirm that there is no evidence to reject the normal
distribution.

\begin{figure}[hbt]
\includegraphics[width=0.8\linewidth]{Identifying-Heterogeneity-in-SAR-Data-with-New-Test-Statistics_files/figure-latex/Plot_density-1} \caption{Empirical density  under the null hypothesis.}\label{fig:Plot_density}
\end{figure}

\begin{table}[H]

\caption{\label{tab:table_stat_combined}Descriptive analysis of $S_H$ with $L\in\left\{3,5, 8,11\right\}$ and $\mu=1$.}
\begin{tabular}[t]{lrrrrrrr}
\toprule
\multicolumn{1}{c}{$L$} & \multicolumn{1}{c}{$n$} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{Var} & \multicolumn{1}{c}{SK} & \multicolumn{1}{c}{EK} & \multicolumn{1}{c}{$p$-value}\\
\midrule
 & $25$ & $-0.0596$ & $\phantom{-}0.2491$ & $\phantom{-}0.0620$ & $\phantom{-}0.2206$ & $\phantom{-}0.8568$ & $\phantom{-}0.0051$\\

 & $49$ & $-0.0232$ & $\phantom{-}0.1767$ & $\phantom{-}0.0312$ & $\phantom{-}0.2322$ & $\phantom{-}0.5200$ & $\phantom{-}0.2622$\\

 & $81$ & $\phantom{-}0.0047$ & $\phantom{-}0.1297$ & $\phantom{-}0.0168$ & $\phantom{-}0.3550$ & $\phantom{-}0.5413$ & $\phantom{-}0.0725$\\

\multirow{-4}{*}[1.5\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash 3} & $121$ & $\phantom{-}0.0011$ & $\phantom{-}0.1018$ & $\phantom{-}0.0104$ & $-0.0065$ & $\phantom{-}0.5275$ & $\phantom{-}0.2092$\\
\cmidrule{1-8}
 & $25$ & $-0.0692$ & $\phantom{-}0.2563$ & $\phantom{-}0.0657$ & $\phantom{-}0.0527$ & $\phantom{-}0.5330$ & $\phantom{-}0.0336$\\

 & $49$ & $-0.0197$ & $\phantom{-}0.1786$ & $\phantom{-}0.0319$ & $\phantom{-}0.4864$ & $\phantom{-}0.5653$ & $\phantom{-}0.0008$\\

 & $81$ & $-0.0075$ & $\phantom{-}0.1389$ & $\phantom{-}0.0193$ & $\phantom{-}0.2414$ & $\phantom{-}0.2546$ & $\phantom{-}0.0183$\\

\multirow{-4}{*}[1.5\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash 5} & $121$ & $\phantom{-}0.0019$ & $\phantom{-}0.1083$ & $\phantom{-}0.0117$ & $\phantom{-}0.0857$ & $\phantom{-}0.4701$ & $\phantom{-}0.1208$\\
\cmidrule{1-8}
 & $25$ & $-0.0669$ & $\phantom{-}0.2745$ & $\phantom{-}0.0753$ & $\phantom{-}0.4127$ & $\phantom{-}0.7551$ & $\phantom{-}0.0033$\\

 & $49$ & $-0.0366$ & $\phantom{-}0.1808$ & $\phantom{-}0.0327$ & $\phantom{-}0.3316$ & $\phantom{-}0.7583$ & $\phantom{-}0.0802$\\

 & $81$ & $-0.0064$ & $\phantom{-}0.1221$ & $\phantom{-}0.0149$ & $\phantom{-}0.0012$ & $-0.0649$ & $\phantom{-}0.9260$\\

\multirow{-4}{*}[1.5\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash 8} & $121$ & $-0.0083$ & $\phantom{-}0.1173$ & $\phantom{-}0.0138$ & $-0.1351$ & $\phantom{-}0.0770$ & $\phantom{-}0.1648$\\
\cmidrule{1-8}
 & $25$ & $-0.0954$ & $\phantom{-}0.2552$ & $\phantom{-}0.0651$ & $\phantom{-}0.2723$ & $\phantom{-}0.6899$ & $\phantom{-}0.2190$\\

 & $49$ & $-0.0549$ & $\phantom{-}0.1716$ & $\phantom{-}0.0294$ & $\phantom{-}0.0352$ & $\phantom{-}0.2073$ & $\phantom{-}0.2847$\\

 & $81$ & $-0.0159$ & $\phantom{-}0.1309$ & $\phantom{-}0.0171$ & $\phantom{-}0.0199$ & $-0.2283$ & $\phantom{-}0.4678$\\

\multirow{-4}{*}[1.5\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash 11} & $121$ & $\phantom{-}0.0020$ & $\phantom{-}0.1182$ & $\phantom{-}0.0140$ & $-0.0046$ & $\phantom{-}0.1101$ & $\phantom{-}0.8906$\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}[hbt]
\includegraphics[width=0.8\linewidth]{Identifying-Heterogeneity-in-SAR-Data-with-New-Test-Statistics_files/figure-latex/Plot_normality_qq-1} \caption{Normal QQ-plots for  $n=121$.}\label{fig:Plot_normality_qq}
\end{figure}

After verifying the normality of the data, we proceeded to explore the
capabilities of the proposed test in terms of size and power. Under
\(\mathcal{H}_0\), the distribution of test statistic is asymptotically
normal. Therefore, the \(p\)-value is calculated as
\(2\Phi(-\mid \varepsilon\mid)\), where \(\Phi\) is the standard
Gaussian cumulative distribution and \(\varepsilon\) is the standardized
test statistic given by: \[
\varepsilon=\frac{\tilde{H}-H_{\Gamma_{\text{SAR}}}}{\hat\sigma}.
\]

We considered nominal levels of \SI{1}{\percent}, \SI{5}{\percent}, and
\SI{10}{\percent}. In reference to the size, \(1000\) simulations have
been used for various sample sizes with satisfactory results. In all
cases, the nominal level has been achieved. The power, in general,
improves with the increase in the sample size and the number of looks.
The results are presented in Table~\ref{tab:table_size_power}.

\hypertarget{the-proposed-test-based-on-coefficient-of-variation-and-a-robust-alternative}{%
\subsection{The Proposed Test Based on Coefficient of Variation and a
Robust
Alternative}\label{the-proposed-test-based-on-coefficient-of-variation-and-a-robust-alternative}}

In addition to the entropy-based test, we also propose a test statistic
based on the classical CV. This test statistic is defined as follows:
\begin{align}
    T_{\text{CV}}=\frac{S}{\overline{Z}},
\end{align} where \(S\) and \(\overline{Z}\) are the sample standard
deviation and the sample mean, respectively.

Similarly, we use another test statistic based on the ratio of the MnAD
to the median. This statistic is given by: \begin{align}
    T_{\text{CV}_{\text{MnAD}}}=\frac{\text{MnAD}}{\text{Median}}.
\end{align} These test statistics allow assessing the variability of the
data.

The situations where the use of CV and \(\text{CV}_{\text{MnAD}}\) may
be appropriate, i.e.~where the observations are positive, the log-normal
(LN) and inverse Gaussian (IG) distribution are often more justified
compared to Gamma and Weibull distributions
\cite{Chaubey2017,takagi1997application}.

The inverse Gaussian distribution is shown to be well approximated by
the log-normal distribution \cite{whitmore1978} which implies that the
IG distribution also does not share the problem of non-existence of
fixed width confidence interval with the Gaussian case.

The LN distribution in two parameters has the following density
function. \begin{align}
    f_Z(z;\mu, \sigma )=\frac{1}{\sigma z\sqrt{2\pi}}\exp\left\{-\frac{(\ln z- \mu)^2}{2\sigma^2}\right\}, \quad z>0, \sigma>0, -\infty < \mu < \infty,
\end{align} here, \(\mu\) is the log-scale parameter and \(\sigma\) is
the shape parameter.

\hypertarget{model-selection-criterion}{%
\subsection{Model Selection Criterion}\label{model-selection-criterion}}

Akaike information criterion (AIC) and Bayesian information criterion
(BIC) are used to select the best fitting distribution.

The AIC is a fine technique that deals with the trade-off between the
goodness-of-fit of the model and the simplicity of the model in terms of
the number of model parameters~\cite{Burnham2004}. The model or
distribution with the lowest value of AIC is chosen to be the best. The
BIC assesses goodness-of-fit of a distribution or model, but avoids
overfitting by penalising an additional degree of
freedom~\cite{Dziak2019}. The model with the lowest BIC value is chosen
as the best.

The AIC and BIC results in
Tables~\ref{tab:table_aic_gamma}-\ref{tab:table_aic_gio_MnADmedian}
indicate that the CV and \(\text{CV}_{\text{MnAD}}\) data from different
synthetic sample sizes, extracted from the Gamma SAR and \(G_I^0\)
distributions, respectively, adhere to the characteristics of an LN
distribution. The AIC and BIC values of the IG distribution and the LN
distribution are not significantly different because both distributions
are based on the right-skewed characteristic. In this situation, the
model for the LN distribution was considered to be the best due to it
providing the lowest AIC and BIC values.

Figures~\ref{fig:Plot_cv}-\ref{fig:Plot_madmed_gi0_MnADmedian} present
density plots, Q-Q plots, P-P plots, and Cumulative Distribution
Function plots, which offer an additional measure of goodness-of-fit and
confirm that the most suitable distribution to be considered is the LN
distribution.

\begin{table}[H]

\caption{\label{tab:table_aic_gamma}AIC and BIC values for evaluating the best distribution with CV data from Gamma SAR.}
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{\textbf{Criterion}} & \multicolumn{1}{c}{$\bm{n}$} & \multicolumn{1}{c}{\textbf{Normal}} & \multicolumn{1}{c}{\textbf{Lognormal}} & \multicolumn{1}{c}{\textbf{Gamma}} & \multicolumn{1}{c}{\textbf{Weibull}} & \multicolumn{1}{c}{\textbf{Inverse Gaussian}}\\
\midrule
 & $25$ & $-38031.95$ & $-38266.75$ & $-38311.84$ & $-36413.23$ & $-38261.60$\\

 & $49$ & $-47698.20$ & $-47913.72$ & $-47905.63$ & $-45553.97$ & $-47911.65$\\

 & $81$ & $-55382.11$ & $-55494.72$ & $-55494.88$ & $-53220.44$ & $-55493.85$\\

\multirow{-4}{*}[1.5\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash AIC} & $121$ & $-61344.92$ & $-61470.83$ & $-61453.83$ & $-58876.04$ & $-61470.48$\\
\cmidrule{1-7}
 & $25$ & $-38016.72$ & $-38251.52$ & $-38296.61$ & $-36397.99$ & $-38246.37$\\

 & $49$ & $-47682.96$ & $-47898.49$ & $-47890.40$ & $-45538.73$ & $-47896.42$\\

 & $81$ & $-55366.88$ & $-55479.49$ & $-55479.65$ & $-53205.21$ & $-55478.62$\\

\multirow{-4}{*}[1.5\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash BIC} & $121$ & $-61329.68$ & $-61455.60$ & $-61438.60$ & $-58860.81$ & $-61455.25$\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]

\caption{\label{tab:table_aic}AIC and BIC values for evaluating the best distribution with CV data from $G_I^0$.}
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{\textbf{Criterion}} & \multicolumn{1}{c}{$\bm{n}$} & \multicolumn{1}{c}{\textbf{Normal}} & \multicolumn{1}{c}{\textbf{Lognormal}} & \multicolumn{1}{c}{\textbf{Gamma}} & \multicolumn{1}{c}{\textbf{Weibull}} & \multicolumn{1}{c}{\textbf{Inverse Gaussian}}\\
\midrule
 & $25$ & $\phantom{-}8254.04$ & $\phantom{-}2186.31$ & $\phantom{-}3628.40$ & $\phantom{-}8383.58$ & $\phantom{-}2257.63$\\

 & $49$ & $\phantom{-}8821.79$ & $\phantom{-}1689.03$ & $\phantom{-}3483.10$ & $\phantom{-}9533.53$ & $\phantom{-}1835.29$\\

 & $81$ & $\phantom{-}8525.81$ & $\phantom{-}866.29$ & $\phantom{-}2853.31$ & $\phantom{-}9822.48$ & $\phantom{-}1057.91$\\

\multirow{-4}{*}[1.5\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash AIC} & $121$ & $\phantom{-}8708.81$ & $\phantom{-}131.86$ & $\phantom{-}2341.06$ & $\phantom{-}10506.49$ & $\phantom{-}398.53$\\
\cmidrule{1-7}
 & $25$ & $\phantom{-}8269.27$ & $\phantom{-}2201.54$ & $\phantom{-}3643.63$ & $\phantom{-}8398.81$ & $\phantom{-}2272.86$\\

 & $49$ & $\phantom{-}8837.02$ & $\phantom{-}1704.26$ & $\phantom{-}3498.33$ & $\phantom{-}9548.76$ & $\phantom{-}1850.52$\\

 & $81$ & $\phantom{-}8541.04$ & $\phantom{-}881.52$ & $\phantom{-}2868.55$ & $\phantom{-}9837.72$ & $\phantom{-}1073.14$\\

\multirow{-4}{*}[1.5\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash BIC} & $121$ & $\phantom{-}8724.04$ & $\phantom{-}147.09$ & $\phantom{-}2356.29$ & $\phantom{-}10521.72$ & $\phantom{-}413.76$\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]

\caption{\label{tab:table_aic_gamma_madmed}AIC and BIC values for evaluating the best distribution with $\text{CV}_{\text{MnAD}}$ data from Gamma SAR.}
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{\textbf{Criterion}} & \multicolumn{1}{c}{$\bm{n}$} & \multicolumn{1}{c}{\textbf{Normal}} & \multicolumn{1}{c}{\textbf{Lognormal}} & \multicolumn{1}{c}{\textbf{Gamma}} & \multicolumn{1}{c}{\textbf{Weibull}} & \multicolumn{1}{c}{\textbf{Inverse Gaussian}}\\
\midrule
 & $25$ & $-38375.56$ & $-39147.85$ & $-39066.29$ & $-36652.49$ & $-39143.61$\\

 & $49$ & $-48386.11$ & $-48795.32$ & $-48745.48$ & $-46240.91$ & $-48793.83$\\

 & $81$ & $-56072.87$ & $-56322.32$ & $-56290.02$ & $-53836.04$ & $-56322.12$\\

\multirow{-4}{*}[1.5\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash AIC} & $121$ & $-62217.14$ & $-62394.77$ & $-62369.32$ & $-59861.80$ & $-62394.57$\\
\cmidrule{1-7}
 & $25$ & $-38360.32$ & $-39132.62$ & $-39051.05$ & $-36637.26$ & $-39128.38$\\

 & $49$ & $-48370.87$ & $-48780.09$ & $-48730.25$ & $-46225.67$ & $-48778.60$\\

 & $81$ & $-56057.64$ & $-56307.09$ & $-56274.79$ & $-53820.81$ & $-56306.89$\\

\multirow{-4}{*}[1.5\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash BIC} & $121$ & $-62201.91$ & $-62379.54$ & $-62354.09$ & $-59846.56$ & $-62379.33$\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]

\caption{\label{tab:table_aic_gio_MnADmedian}AIC and BIC values for evaluating the best distribution with $\text{CV}_{\text{MnAD}}$ data from $G_I^0$.}
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{\textbf{Criterion}} & \multicolumn{1}{c}{$\bm{n}$} & \multicolumn{1}{c}{\textbf{Normal}} & \multicolumn{1}{c}{\textbf{Lognormal}} & \multicolumn{1}{c}{\textbf{Gamma}} & \multicolumn{1}{c}{\textbf{Weibull}} & \multicolumn{1}{c}{\textbf{Inverse Gaussian}}\\
\midrule
 & $25$ & $-13302.39$ & $-15575.23$ & $-15158.42$ & $-11933.35$ & $-15565.03$\\

 & $49$ & $-23265.27$ & $-24529.72$ & $-24284.47$ & $-20986.94$ & $-24522.28$\\

 & $81$ & $-29908.19$ & $-30960.38$ & $-30747.90$ & $-25233.41$ & $-30946.54$\\

\multirow{-4}{*}[1.5\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash AIC} & $121$ & $-36496.78$ & $-37128.07$ & $-36991.41$ & $-32366.52$ & $-37123.20$\\
\cmidrule{1-7}
 & $25$ & $-13287.16$ & $-15559.99$ & $-15143.19$ & $-11918.12$ & $-15549.80$\\

 & $49$ & $-23250.04$ & $-24514.48$ & $-24269.24$ & $-20971.71$ & $-24507.05$\\

 & $81$ & $-29892.96$ & $-30945.15$ & $-30732.67$ & $-25218.17$ & $-30931.31$\\

\multirow{-4}{*}[1.5\dimexpr\aboverulesep+\belowrulesep+\cmidrulewidth]{\raggedright\arraybackslash BIC} & $121$ & $-36481.55$ & $-37112.84$ & $-36976.18$ & $-32351.28$ & $-37107.97$\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{Identifying-Heterogeneity-in-SAR-Data-with-New-Test-Statistics_files/figure-latex/Plot_cv_gamma-1}

}

\caption{Goodness of fit plots for evaluating the best distribution with CV data from Gamma SAR, with  $n=49$, $L=5$ and $\mu=1$.}\label{fig:Plot_cv_gamma}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{Identifying-Heterogeneity-in-SAR-Data-with-New-Test-Statistics_files/figure-latex/Plot_cv-1}

}

\caption{Goodness of fit plots for evaluating the best distribution with $\text{CV}_{\text{MnAD}}$ data from $G_I^0$, with  $n=49$, $L=5$, $\mu=1$ and $\alpha=-2$.}\label{fig:Plot_cv}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{Identifying-Heterogeneity-in-SAR-Data-with-New-Test-Statistics_files/figure-latex/Plot_MnADmedian_gamma-1}

}

\caption{Goodness of fit plots for evaluating the best distribution with $\text{CV}_{\text{MnAD}}$ data from Gamma SAR, with  $n=49$, $L=5$ and $\mu=1$.}\label{fig:Plot_MnADmedian_gamma}
\end{figure}

\begin{figure}[hbt]
\includegraphics[width=1\linewidth]{Identifying-Heterogeneity-in-SAR-Data-with-New-Test-Statistics_files/figure-latex/Plot_madmed_gi0_MnADmedian-1} \caption{Goodness of fit plots for evaluating the best distribution with $CV_{\text{MnAD}}$ data from $G_I^0$, with  $n=49$, $L=5$, $\mu=1$ and $\alpha=-5$.}\label{fig:Plot_madmed_gi0_MnADmedian}
\end{figure}

\hypertarget{sec:Results}{%
\section{Simulation Results}\label{sec:Results}}

This section presents simulations performed to obtain an assessment of
the performances of the proposed test statistics.

\hypertarget{simulated-data}{%
\subsection{Simulated Data}\label{simulated-data}}

Figure 1a shows the phantom used for a simulated image with dimensions
of \(250\times 250\) pixels to illustrate the differences between
heterogeneous (dark, \(\alpha=-2\)) and homogeneous (light,
\(\alpha=-10\)) regions, with \(L = 5\). The areas are formed by
independent draws from the \(G^0_I\) distribution.

The three test statistics are applied to the simulated image using local
sliding windows of size \(7\times 7\).

The resulting \(p\)-values for the tests \(S_H\), \(T_{\text{CV}}\), and
\(T_{\text{CV}_{\text{MnAD}}}\) are shown in
Figures~\ref{fig:sim_SAR_Images}(b)-(d) and
Figures~\ref{fig:sim_SAR_Images_p05}(a)-(c), respectively. The new
images reveal that a darker background indicates homogeneity, while
lighter backgrounds represent heterogeneous areas.

\begin{figure}[H]

{\centering \subfloat[\label{fig:sim_SAR_Images-1}]{\includegraphics[width=40mm]{../Figures/PNG/simulated7x7} }\subfloat[\label{fig:sim_SAR_Images-2}]{\includegraphics[width=40mm]{../Figures/PNG/sim_pvalue7x7_total_H} }\subfloat[\label{fig:sim_SAR_Images-3}]{\includegraphics[width=40mm]{../Figures/PNG/sim_pvalue7x7_total_cv} }\newline\subfloat[\label{fig:sim_SAR_Images-4}]{\includegraphics[width=40mm]{../Figures/PNG/sim_pvalue7x7_total_mnad} }

}

\caption{Results of applying the test statistics. (a) Simulated image; $P$-values from the entropy-based test; (c) $P$-values from the CV test; (d) $P$-values from the   $\text{CV}_{\text{MnAD}}$ test; }\label{fig:sim_SAR_Images}
\end{figure}
\begin{figure}[H]

{\centering \subfloat[ \label{fig:sim_SAR_Images_p05-1}]{\includegraphics[width=40mm]{../Figures/PNG/sim_pvalue7x7_H} }\subfloat[ \label{fig:sim_SAR_Images_p05-2}]{\includegraphics[width=40mm]{../Figures/PNG/sim_pvalue7x7_cv} }\subfloat[ \label{fig:sim_SAR_Images_p05-3}]{\includegraphics[width=40mm]{../Figures/PNG/sim_pvalue7x7_mnad} }

}

\caption{Results of tests on simulated data with a threshold of $0.1$ of the $p$-value. (a) Entropy-based test; (b) CV test; (c) $\text{CV}_{\text{MnAD}}$ test.}\label{fig:sim_SAR_Images_p05}
\end{figure}

\hypertarget{real-data}{%
\subsection{Real Data}\label{real-data}}

We applied the proposed test statistics in actual SAR images acquired
over cities, including Flevoland, Netherlands and Ottawa, Canada.\\
These images were captured by the Sentinel-1A satellite operating in the
C band, with VV polarization, intensity format, and \(L=5\) nominal
looks. Both images have dimensions of \(512 \times 512\) pixels and
contain urban areas, water forest regions, as shown in
Figures~\ref{fig:real_SAR_Images_coe}(a)-(b).

The \(p\)-values obtained for the \(S_H\), \(T_{\text{CV}}\), and
\(T_{\text{CV}_{\text{MnAD}}}\) tests are presented in
Figures~\ref{fig:sim_SAR_Images_Frankfurt}-\ref{fig:sim_SAR_Images_01_Ottawa},
respectively.

\begin{figure}[H]

{\centering \subfloat[ \label{fig:real_SAR_Images_coe-1}]{\includegraphics[width=40mm]{../Figures/PNG/Intensity_Frankfurt_512_n} }\subfloat[ \label{fig:real_SAR_Images_coe-2}]{\includegraphics[width=40mm]{../Figures/PNG/Intensity_Ottawa_512_n} }

}

\caption{SAR images. (a) Frankfurt; (b) Ottawa. }\label{fig:real_SAR_Images_coe}
\end{figure}
\begin{figure}[H]

{\centering \subfloat[ \label{fig:sim_SAR_Images_Frankfurt-1}]{\includegraphics[width=40mm]{../Figures/PNG/Frankfurt_512_pvalue7x7_total_H} }\subfloat[ \label{fig:sim_SAR_Images_Frankfurt-2}]{\includegraphics[width=40mm]{../Figures/PNG/Frankfurt_512_pvalue7x7_total_CV} }\subfloat[ \label{fig:sim_SAR_Images_Frankfurt-3}]{\includegraphics[width=40mm]{../Figures/PNG/Frankfurt_512_mnad_7_total} }

}

\caption{Results of applying the test statistics, Frankfurt. (a) $P$-values from the entropy-based test; (b) $P$-values from the CV test; (c) $P$-values from the   $\text{CV}_{\text{MnAD}}$ test; }\label{fig:sim_SAR_Images_Frankfurt}
\end{figure}
\begin{figure}[H]

{\centering \subfloat[ \label{fig:sim_SAR_Images_01Frankfurt-1}]{\includegraphics[width=40mm]{../Figures/PNG/Frankfurt_512_pvalue7x7_H_01} }\subfloat[ \label{fig:sim_SAR_Images_01Frankfurt-2}]{\includegraphics[width=40mm]{../Figures/PNG/Frankfurt_512_pvalue7x7_CV} }\subfloat[ \label{fig:sim_SAR_Images_01Frankfurt-3}]{\includegraphics[width=40mm]{../Figures/PNG/Frankfurt_512_mnad_7_0.1} }

}

\caption{Results of tests on simulated data with a threshold of $0.1$ of the $p$-value, Frankfurt. (a) Entropy-based test; (b) CV test; (c) $\text{CV}_{\text{MnAD}}$ test.}\label{fig:sim_SAR_Images_01Frankfurt}
\end{figure}

\begin{figure}[H]

{\centering \subfloat[ \label{fig:sim_SAR_Images_Ottawa-1}]{\includegraphics[width=40mm]{../Figures/PNG/Ottawa_512_pvalue7x7_total_H} }\subfloat[ \label{fig:sim_SAR_Images_Ottawa-2}]{\includegraphics[width=40mm]{../Figures/PNG/Ottawa_512_CV_pvalue7x7_total_CV} }\subfloat[ \label{fig:sim_SAR_Images_Ottawa-3}]{\includegraphics[width=40mm]{../Figures/PNG/Ottawa_512_mnad_7_total} }

}

\caption{Results of applying the test statistics, Ottawa. (a) $P$-values from the entropy-based test; (b) $P$-values from the CV test; (c) $P$-values from the   $\text{CV}_{\text{MnAD}}$ test; }\label{fig:sim_SAR_Images_Ottawa}
\end{figure}
\begin{figure}[H]

{\centering \subfloat[ \label{fig:sim_SAR_Images_01_Ottawa-1}]{\includegraphics[width=40mm]{../Figures/PNG/Ottawa_512_pvalue7x7_H_01} }\subfloat[ \label{fig:sim_SAR_Images_01_Ottawa-2}]{\includegraphics[width=40mm]{../Figures/PNG/Ottawa_512_CV_pvalue7x7_CV} }\subfloat[ \label{fig:sim_SAR_Images_01_Ottawa-3}]{\includegraphics[width=40mm]{../Figures/PNG/Ottawa_512_mnad_7_0.1} }

}

\caption{Results of tests on simulated data with a threshold of $0.1$ of the $p$-value, Ottawa. (a) Entropy-based test; (b) CV test; (c) $\text{CV}_{\text{MnAD}}$ test.}\label{fig:sim_SAR_Images_01_Ottawa}
\end{figure}

\hypertarget{sec:conclusion}{%
\section{Conclusion}\label{sec:conclusion}}

In this article,\ldots{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\supplementary{This article is fully reproducible using RMarkdown. All
code and data are accessible at
\url{https://github.com/rjaneth/identifying-heterogeneity-in-sar-data-with-new-test-statistics}
(accessed on 5 April 2024).}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short
paragraph specifying their individual contributions must be provided.
The following statements should be used ``X.X. and Y.Y. conceive and
designed the experiments; X.X. performed the experiments; X.X. and Y.Y.
analyzed the data; W.W. contributed reagents/materials/analysis tools;
Y.Y. wrote the paper.'\,' Authorship must be limited to those who have
contributed substantially to the work reported.}

\funding{Please add: ``This research received no external funding'' or
``This research was funded by NAME OF FUNDER grant number XXX.'' and
``The APC was funded by XXX'\,'. Check carefully that the details given
are accurate and use the standard spelling of funding agency names at
\url{https://search.crossref.org/funding}, any errors may affect your
future funding.}

\institutionalreview{Not applicable.}

\dataavailability{Not applicable.}

\acknowledgments{This study was financed in part by the Coordenação de
Aperfeiçoamento de Pessoal de Nível Superior - Brasil (CAPES) - Finance
Code 001.}

\conflictsofinterest{The authors declare no conflict of interest.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional

%% Only for journal Encyclopedia

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{adjustwidth}{-\extralength}{0cm}

%\printendnotes[custom] % Un-comment to print a list of endnotes

\reftitle{References}
\bibliography{references.bib}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\PublishersNote{}
\end{adjustwidth}

\end{document}