### 2024-04-24 Text/Identifying Heterogeneity in SAR Data with New Test Statistics.Rmd
Made corrections and added new images

---
title: "Identifying Heterogeneity in SAR Data with New Test Statistics"
author:
  - name: Alejandro C. Frery
    affil: 1,*
    orcid: 0000-0002-8002-5341
  - name: Janeth Alpala
    affil: 2 
    orcid: 0000-0002-0265-6236
  - name: Abraão D. C. Nascimento
    affil: 2
    orcid: 0000-0003-2673-219X
affiliation:
  - num: 1
    address: |
      School of Mathematics and Statistics, Victoria University of Wellington, Wellington 6140, New Zealand
#email: alejandro.frery@vuw.ac.nz
  - num: 2
    address: |
      Departamento de Estatística, Universidade Federal de Pernambuco, Recife 50670-901, PE, Brazil
#email: janeth.alpala@ufpe.br, abraao@de.ufpe.br

# author citation list in chicago format
authorcitation: |
  Frery, A.C.; Alpala, J.; Nascimento,A.D.C.
# firstnote to eighthnote
#firstnote: |
 # Current address: Updated affiliation
#secondnote: |
  #These authors contributed equally to this work.
correspondence: |
  alejandro.frery@vuw.ac.nz
# document options
journal: remotesensing
type: article
status: submit
# front matter
#simplesummary: |
  #A Simple summary goes here.
abstract: |
  This paper presents a statistical approach to identify the underlying roughness characteristics in synthetic aperture radar (SAR) intensity data. 
  The physical modeling of this kind of data allows to use the Gamma distribution in the presence of fully-developed speckle, i.e., when there are infinitely many independent backscatterers per resolution cell, and none dominates the return. This situation is often called "homogeneous" or  "textureless" regions. 
  The $\mathcal{G}_I^0$ distribution is also a widely accepted law for heterogeneous and extremely heterogeneous regions, i.e., areas in which the fully-developed speckle hypotheses do not hold. 
  We propose three test statistics to distinguish between homogeneous and inhomogeneous regions, i.e. between gamma and \(\mathcal{G}_I^0\) distributed data, both with a known number of looks. 
  The first test statistic uses a bootstrapped non-parametric estimator of Shannon entropy, which provides a robust assessment in the face of uncertain distributional assumptions.
  The second test uses the classical coefficient of variation (CV). 
  The third test uses an alternative form of estimating the CV based on the ratio of the mean absolute deviation from the median to the median. 
  We apply our test statistic to create maps of \(p\) values for the homogeneity hypothesis.
  Finally, we show that our proposal, the entropy-based test, outperforms existing methods such as the classical CV and its alternative variant, in identifying heterogeneity when applied to both simulated and actual data.
#These maps identify different types of targets.
# We show that our proposal outperforms its competitors using simulated and SAR data.
# ACF Which is "our proposal" and which are the "competitors"? # ok
keywords: |
  SAR; heterogeneity; entropy;  coefficient of variation; hypothesis tests
acknowledgement: |
  The author J.A. would like to thank the Coordenação de Aperfeiçoamento de Pessoal de Nível Superior (CAPES) and the Fundação de Amparo à Ciência e Tecnologia de Pernambuco (FACEPE) for their support. 
authorcontributions: |
  All authors (A.C.F., J.A, and A.D.C.N.) discussed the results and contributed to the research. All authors have read and agreed to the published version of the manuscript.

funding: |
  Please add: "This research received no external funding" or "This research 
  was funded by NAME OF FUNDER grant number XXX."  and ``The APC was funded 
  by XXX''. Check carefully that the details given are accurate and use the 
  standard spelling of funding agency names at 
  \url{https://search.crossref.org/funding}, any errors may affect your future 
  funding.
institutionalreview: |
 Not applicable.
#informedconsent: |
  # Any research article describing a study involving humans should contain this 
  # statement. Please add ``Informed consent was obtained from all subjects 
  # involved in the study.'' OR ``Patient consent was waived due to REASON 
  # (please provide a detailed justification).'' OR ``Not applicable'' for 
  # studies not involving humans. You might also choose to exclude this statement 
  # if the study did not involve humans.

  # Written informed consent for publication must be obtained from participating 
  # patients who can be identified (including by the patients themselves). Please 
  # state ``Written informed consent has been obtained from the patient(s) to 
  # publish this paper'' if applicable.
dataavailability: |
  The Sentinel-1  images are made publicly available by European Space Agency via Copernicus Data Space Ecosystem at \url{https://dataspace.copernicus.eu/} (accessed on 24 April 2024).
conflictsofinterest: |
  The authors declare no conflict of interest.
supplementary: |
 This article was written in Rmarkdown and is fully reproducible. The code and data are accessible at
 \url{https://github.com/rjaneth/identifying-heterogeneity-in-sar-data-with-new-test-statistics} (accessed on 24 April 2024).
# abbreviations:
#   - short: MDPI
#     long: Multidisciplinary Digital Publishing Institute
#   - short: DOAJ
#     long: Directory of open access journals
#   - short: TLA
#     long: Three letter acronym
#   - short: LD 
#     long: linear dichroism

header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   - \usepackage[T1]{fontenc}
   - \usepackage{polski}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{xcolor}
   - \usepackage{amsmath}
   - \usepackage{multirow}
   - \usepackage{subcaption}
   - \captionsetup[sub]{labelfont=bf,textfont=bf}

bibliography: references.bib
#appendix: appendix.tex
#endnotes: false
output:
  rticles::mdpi_article:
    extra_dependencies: longtable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE)
#knitr::opts_chunk$set(echo = FALSE)
library(fitdistrplus)
library(ggplot2)
library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(rmutil)
library(invgamma)
library(tidyverse)
library(RColorBrewer)
library(ggsci)
library(ggpubr)
library(patchwork)
library(dplyr)
#options(kableExtra.latex.load_packages = FALSE)
#library(devtools)
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)
library(latex2exp)
library(e1071)# kurtosis
library(viridis)
library(nortest)# AD
#library(MASS)#
# model selection AIC and BIC

theme_set(theme_minimal() +
            theme(text=element_text(family="serif"),
                  legend.position = "bottom")#  top , right , bottom , or left#, panel.grid = element_blank()
)

source("../Code/R/MainFunctions/gamma_sar_sample.R")
source("../Code/R/MainFunctions/entropy_gamma_sar.R")
source("../Code/R/MainFunctions/entropy_gI0.R")
source("../Code/R/MainFunctions/gi0_sample.R")

source("../Code/R/MainFunctions/van_es_estimator.R")
source("../Code/R/MainFunctions/correa_estimator.R")
source("../Code/R/MainFunctions/bootstrap_correa_estimator_log_mean.R")
source("../Code/R/MainFunctions/ebrahimi_estimator.R")
#source("../Code/R/MainFunctions/noughabi_arghami_estimator.R")
source("../Code/R/MainFunctions/vasicek_estimator.R")
source("../Code/R/MainFunctions/al_omari_1_estimator.R")
source("../Code/R/MainFunctions/al_omari_2_estimator.R")

source("../Code/R/MainFunctions/bootstrap_van_es_estimator.R")
source("../Code/R/MainFunctions/bootstrap_correa_estimator.R")
source("../Code/R/MainFunctions/bootstrap_ebrahimi_estimator.R")
#source("../Code/R/MainFunctions/bootstrap_noughabi_arghami_estimator.R")
source("../Code/R/MainFunctions/bootstrap_vasicek_estimator.R")
source("../Code/R/MainFunctions/bootstrap_al_omari_1_estimator.R")
source("../Code/R/MainFunctions/bootstrap_al_omari_2_estimator.R")
#The next function contains the functions: generate_samples, calculate_bias_mse, generate_plot
source("../Code/R/Programs/functions_sample_bias_mse.R")# read_ENVI_images
source("../Code/R/Programs/read_ENVI_images.R")

```

\newcommand{\bias}{\operatorname{Bias}}
\newcommand{\widebar}[1]{\overline{#1}}

# Introduction {#sec:Introduction}

Synthetic Aperture Radar (SAR) technology has become an important tool for environmental monitoring and disaster management. It provides valuable images under various conditions, including day or night and different weather situations&nbsp;\cite{Moreira2013,Mu2019}.
However, the effective use of SAR data depends on a thorough understanding of their statistical properties because they are corrupted by speckle.
This noise-like interference effect is inherent in SAR data due to the coherent nature of the imaging process&nbsp;\cite{Argenti2013}.

Speckle is non-Gaussian speckle, thus SAR data require reliable statistical models for accurate processing. 
The $\mathcal{G}^0$ distribution, which is suitable for SAR data, includes the
Gamma law as the limiting case for fully-developed speckle&nbsp;\cite{Ferreira2020} and provides flexibility with fewer parameters for analysis.

Our work aims to improve the identification of potential roughness features in SAR intensity data.
<!-- ACF The following sentence is misleading: the gamma law is a consequence of infinitely many backscatterers with none dominating the return, not the other way aroud. Rephrase. -->
<!-- JA Corrected -->
Physical modeling of SAR data allows the use of the Gamma distribution in the presence of fully-developed speckle, where an infinite number of independent backscatterers per resolution unit is assumed, commonly referred to as homogeneous regions

In this context, we present a set of three novel test statistics that aim to distinguish between homogeneous and non-homogeneous return, in particular between gamma and \(\mathcal{G}^0\) distributed data, assuming the number of looks is known. 
We use properties such as entropy and coefficient of variation.

Entropy is a fundamental concept in information theory with far-reaching applications in pattern recognition, statistical physics, image processing, edge detection and SAR image analysis&nbsp;\cite{Presse2013,MohammadDjafari2015,Avval2021, Nascimento2014,Nascimento2019}. 
Shannon introduced it in 1948&nbsp;\cite{Shannon1948} for a random variable as a measure of information and uncertainty.
In statistics, Shannon entropy is a crucial descriptive parameter, especially for evaluating data dispersion and performing tests for normality, exponentiality and uniformity&nbsp;\cite{Wieczorkowski1999,Zamanzade2012}. 
Entropy estimation is associated with some practical difficulties, especially when the model is unknown. In these cases, non-parametric methods are used.
Spacing methods have been discussed as a non-parametric approach in&nbsp;\cite{AlizadehNoughabi2010,Subhash2021}. This strategy provides the flexibility to consider a wide range of models without enforcing certain parametric constraints.

The CV, introduced in 1896 by Pearson&nbsp;\cite{Pearson1896}, is a relative dispersion measure widely used in various fields of applied statistics, including sampling, biostatistics, medical and biological research, climatology and other fields&nbsp;\cite{hendricks1936sampling,Tian2005,SubrahmanyaNairy2003,Chankham2024}.
It facilitates the comparison of variability between different populations and is particularly valuable for comparing variables with different units. 
This is because when the main purpose is to compare the variations of several variables, the standard deviation can only serve as an adequate measure of variation if all variables are expressed in the same unit of measurement and have identical means. 
If these conditions are not met for the variables under investigation, the CV is the relative measure that is usually used in real applications. The variable with the highest CV value is the one with the largest relative dispersion around the mean value&nbsp;\cite{Banik2011}.
The coefficient of variation is the basic measure of heterogeneity in SAR data&nbsp;\cite{oliverquegan98}.
We study two ways of estimating the coefficient of variation.

The other parameter we study is the Shannon entropy.
Different models for SAR data have different entropy values, but this fundamental quantity can be also estimated in a model-agnostic way.
We exploit this property and design a bootstrap-improved non-parametric estimator for the Shannon entropy.

We devise test statistics based on these three estimators: the classical coefficient of variation, a robust version, and the Shannon entropy estimator.
We apply these test statistics to generate maps of evidence of homogeneity
that reveal different types of targets in the SAR data. 
Through extensive experiments with simulated and SAR data, we demonstrate the superiority of our proposed method over existing approaches.

The article is structured as follows: Section&nbsp;\ref{sec:Background}
deals with statistical modeling and entropy estimation for intensity SAR
data. 
Section&nbsp;\ref{sec:test} outlines hypothesis tests based on non-parametric entropy and coefficients of variation estimators. 
In Section&nbsp;\ref{sec:Results} we present experimental results. 
Finally, we draw our conclusions in Section&nbsp;\ref{sec:conclusion}.

# Background {#sec:Background}

## Statistical Modeling of Intensity SAR data

The primary models used for intensity SAR data include the Gamma and $\mathcal{G}_I^0$  distributions&nbsp;[@Frery1997]. 
The first is suitable for fully-developed speckle and is a limiting case of the second model, which is interesting due to its versatility in accurately representing regions with different roughness properties&nbsp;[@Cassetti2022].
We denote $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ and $Z \sim \mathcal{G}_I^0(\alpha, \gamma, L)$ to indicate that $Z$ follows the distributions characterized by the respective probability density
functions (pdfs): 
\begin{align}
    f_Z(z;L, \mu)&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\} \mathbbm 1_{\mathbbm R_+}(z)\label{E:gamma1}\\
    \intertext{ and }
    f_Z(z; \alpha, \gamma, L)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\cdot\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gi01}
\end{align} 
where \(\mu > 0\) is the mean, \(\gamma > 0\) is the scale, \(\alpha < -1\) measures
the roughness, \(L \geq 1\) is the number of looks, \(\Gamma(\cdot)\) is
the gamma function, and \(\mathbbm 1_{A}(z)\) is the indicator function
of the set \(A\).

If we take the law with the density&nbsp;\eqref{E:gi01}, the \(r\)th non-central moment of \(Z\) is expressed as follows:
\begin{align}
    E_{\mathcal{G}_I^0}\left(Z^r\right)=\left(\frac{\gamma}{L}\right)^r\frac{\Gamma(-\alpha-r)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+r)}{\Gamma(L)}, \quad \alpha <-r. 
    \label{E:rmom}
\end{align}

Although the \(\mathcal{G}_I^0\) distribution is defined by the parameters \(\alpha\) and \(\gamma\), in the SAR literature&nbsp;\cite{Nascimento2010} the texture \(\alpha\) and the mean \(\mu\) are usually used. 
The expected value \(\mu\) using the expression in&nbsp;\eqref{E:rmom} is
\begin{equation}
    \mu=\left(\frac{\gamma}{L}\right)\frac{\Gamma(-\alpha-1)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+1)}{\gamma(L)}=-\frac{\gamma}{\alpha+1},
\end{equation} 
and we reparametrize&nbsp;\eqref{E:gi01} with \(\mu\) obtaining:
\begin{equation}
        f_Z(z; \mu, \alpha, L)=\frac{L^L\Gamma(L-\alpha)}{\big[-\mu(\alpha+1)\big]^{\alpha}\Gamma(-\alpha)\Gamma(L)} \frac{z^{L-1}}{\big[-\mu(\alpha+1)+Lz\big]^{L-\alpha}}.\label{E:gi02}
\end{equation}

## The Shannon Entropy
The parametric representation of Shannon entropy for a system described by a continuous random variable is:
\begin{equation}
  \label{E:entropy2}
  H(Z)=-\int_{-\infty }^\infty \ f(z)\ln f(z)\, \mathrm{d}z,
\end{equation}
here, \(f(\cdot)\) is the pdf that characterizes the distribution of the real-valued random variable \(Z\).

Using&nbsp;\eqref{E:entropy2}, we obtain the Shannon entropy of $\Gamma_{\text{SAR}}$ in&nbsp;\eqref{E:gamma1} and $G_I^0$ in&nbsp;\eqref{E:gi02}:
\begin{equation}
\label{E:E-gamma}
H_{\Gamma_{\text{SAR}}}(L, \mu) =   L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) + \ln \mu, 
\end{equation}
\begin{multline}
\label{E:E-GIO}
H_{G_I^0}(\mu, \alpha, L) =L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) +\ln \mu -\ln\Gamma(L-\alpha)\\
+ (L-\alpha) \psi^{(0)}(L-\alpha)-(1-\alpha)\psi^{(0)}(-\alpha)+\ln (-1-\alpha)+\ln\Gamma(-\alpha)-L,
\end{multline}
where $\psi^{(0)}(\cdot)$ is the digamma function. Figure&nbsp;\ref{fig:Plot_GI0_to_gamma1}, shows the entropy of \(G_I^0\) as a function of $\mu$ when \(\alpha \in \left\{-\infty, -20, -8, -3\right\}\).
Notice that it converges to the entropy of \(\Gamma_{\text{SAR}}\) when \(\alpha\to-\infty\), as expected.
The more heterogeneous (large $\alpha$ values) the SAR region is, the larger the entropy (or degree of disorder) is.

```{r Plot_GI0_to_gamma1, echo=FALSE, message=FALSE, warning=FALSE, out.width="70%", fig.show="hold", fig.pos="hbt", fig.cap="$H_{ G_I^0}$ converges to the $H_{\\Gamma_{\\text{SAR}}}$ with $L=8$."}

L <- c(8)
alphas <- c(  -3, -8, -20, -1000)
alpha_labels <- c( expression(italic(alpha) == -3), expression(italic(alpha) == -8), expression(italic(alpha) == -20), expression(italic(alpha) == -1000))

mu <- seq(0.1, 10, length.out = 500)

# Entropy GI0

muEntropy <- data.frame()

for (alpha in alphas) {
  entropies_GI0 <- sapply(L, function(L) entropy_gI0(mu, alpha, L))
  muEntropy <- rbind(muEntropy, data.frame(mu = mu, Entropy = entropies_GI0, alpha = as.factor(alpha)))
}

muEntropy.molten <- melt(muEntropy, id.vars = c("mu", "alpha"),  value.name = "Entropy")

# Entropy Gamma SAR

entropies_gamma <- sapply(L, function(L) entropy_gamma_sar(L, mu))

Entropy_gamma <- data.frame(mu, entropies_gamma)

Entropy_gamma.molten <- melt(Entropy_gamma, id.vars = "mu", value.name = "Entropy Gamma")

#plot

ggplot(muEntropy.molten, aes(x = mu, y = Entropy, col = alpha)) +
  geom_line(data = Entropy_gamma.molten, aes(x = mu, y = `Entropy Gamma`), color = "blue", linetype = "solid",linewidth = 1.5) + 
  geom_line(linetype = "longdash",  linewidth = 2, alpha=.7) +
  annotate("text", x = max(mu)+0.2, y = max(Entropy_gamma.molten$`Entropy Gamma`), 
           label = TeX("$\\Gamma_{\\tiny{SAR}}$"), vjust = 0.9, hjust = 0.1, color = "blue")+
  theme_minimal() +
  scale_color_manual(values = brewer.pal(7, "Dark2")[1:5], labels = alpha_labels) +
  #scale_color_manual(values = pal_jama()(7)[2:5], labels = alpha_labels) +
  #scale_color_manual(values = 	pal_cosmic()(7)[1:5], labels = alpha_labels) +
  labs(col = "Roughness", linetype = NULL) +
  xlab(expression(paste(mu))) +
  ylab("Entropy") +  
  theme(text = element_text(family = "serif"),
        legend.position = "bottom")

```

## Estimation of the Shannon Entropy

Several authors&nbsp;\cite{vasicek1976test, Wieczorkowski1999,correa1995new,AlOmari2019} have studied the problem of the non-parametric estimation of $H(Z)$.
Their proposals use estimators based on differences between order statistics: spacings.

One of the first non-parametric estimators based on spacings was introduced by Vasicek&nbsp;\cite{vasicek1976test}. 
Under the assumption that \(\bm{Z}=(Z_1, Z_2,\ldots,Z_n)\) is a random sample from the distribution \(F(z)\), the estimator is defined as:
\begin{equation*}
\label{E:Vas}
	\widehat{H}_{\text{V}}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\ln\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, $Z_{(i+m)}-Z_{(i-m)}$ is the $m$-spacing and $Z_{(1)}\leq Z_{(2)}\leq\ldots\leq Z_{(n)}$ are the order statistics, $Z_{(i)}= Z_{(1)}$ if $i<1$, $Z_{(i)}= Z_{(n)}$ if $i>n$.

Several authors have explored adaptations to Vasicek's estimator. 
We consider three estimators known for their superior performance&nbsp;\cite{Cassetti2022}.

* Correa&nbsp;\cite{correa1995new}:
\begin{equation}
\widehat{H}_{\text{C}}(\bm{Z})=-\frac{1}{n} \sum_{i=1}^n \log \frac{\sum_{j=i-m}^{i+m}(j-i)\left(Z_{(j)}-\widebar{Z}_{(i)}\right)}{n\sum_{j=i-m}^{i+m}\left(Z_{(j)}-\widebar{Z}_{(i)}\right)^2},
\end{equation}
where  $\widebar{Z}_{(i)}=(2 m+1)^{-1} \sum_{j=i-m}^{i+m} Z_{(j)}$, $m< \frac{n}{2}$, $Z_{(i)}=Z_{(1)}$ for $i<1$ and $Z_{(i)}=Z_{(n)}$ for $i>n$. Based on simulation, he showed that his estimator has a smaller mean square error in comparison with Vasicek's approach.

* Ebrahimi et al.&nbsp;\cite{Ebrahimi1994}:
\begin{equation}
\widehat{H}_{\text{E}}(\bm{Z})=\frac{1}{n} \sum_{i=1}^n \log \left[\frac{n}{c_i m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
\end{equation}
where
$$
c_i=\begin{cases}1+(i-1) / m & \text { if } \quad 1 \leq i \leq m, \\ 2 & \text { if }\quad m+1 \leq i \leq n-m,\\ 1+(n-i) / m & \text { if }\quad n-m+1 \leq i \leq n.\end{cases}
$$
The authors proved that $\widehat{H}_{\text{E}}(\bm{Z})\xrightarrow{p}H[Z]$, when $m,n\rightarrow\infty$ and  $m/n\rightarrow0$.

* Al-Omari&nbsp;\cite{IbrahimAlOmari2014}:
$$
\widehat{H}_{\text{AO}}(\bm{Z})=\frac{1}{n} \sum_{i=1}^n \log \left[\frac{n}{\omega_i m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right]
$$
where
$$
\omega_i= \begin{cases}3/2 & \text { if }\quad 1 \leq i \leq m, \\ 2 & \text { if }\quad m+1 \leq i \leq n-m, \\ 3/2 & \text { if } \quad n-m+1 \leq i \leq n,\end{cases}
$$
in which $Z_{(i-m)}=Z_{(1)}$ for $i \leq m$, and $Z_{(i+m)}=Z_{(n)}$ for $i \geq n-m$.

These estimators are asymptotically consistent, 
but we will use improved bootstrap-improved versions because we need them to perform well with small samples.

## Enhanced estimators with Bootstrap

We use the bootstrap technique to refine the accuracy of non-parametric entropy estimators. 
In this approach, new data sets are generated by replicate sampling from an existing data set&nbsp;\cite{Michelucci2021}.

Let us assume that non-parametric entropy estimators
\(\widehat{H}=\widehat{\theta}(\bm{Z})\) are inherently biased, i.e:\begin{equation}
\label{Eq:bias1}
\operatorname{Bias}\big(\widehat{\theta}(\bm{Z})\big) = E\big[\widehat{\theta}(\bm{Z})\big] - \theta \neq 0.
\end{equation} Our bootstrap-improved estimators are of the form:
\begin{align*}
\widetilde{H} &= 2\widehat{\theta}(\bm{Z}) - \frac{1}{B}\sum_{b=1}^B \widehat{\theta}_b(\bm{Z}^{(b)}),
\end{align*} 
where \(B\) is the number of observations obtained by resampling from $\bm Z$ with replacement. 
Applying this methodology, the original estimators of Correa,
Ebrahimi and Al-Omari are now referred to as the proposed bootstrap-enhanced versions:
\(\widetilde{H}_{\text{C}}\),
\(\widetilde{H}_{\text{E}}\), and \(\widetilde{H}_{\text{AO}}\),
respectively.

We analyzed the performance of these estimators with a Monte Carlo study: \(1000\) samples from the \(\Gamma_{\text{SAR}}\) distribution of size
\(n\in\left\{9, 25, 49, 81, 121\right\}\), with
\(\mu\in\left\{1, 10\right\}\) and \(L=5\). 
The results are consistent with the other situations. 
We used $B=200$ bootstrap samples and the heuristic spacing
\(m=\left[\sqrt{n}+0.5\right]\), as recommended in the literature.

In Figure&nbsp;\ref{fig:Plot_bias_mse_gi0} we show the bias and mean squared error (MSE) of the original non-parametric entropy estimators and their respective bootstrap-enhanced versions. 
Bootstrap-enhanced estimators have smaller bias and MSE, especially for sample sizes below $81$.
The results of the simulation can be found in
Table&nbsp;\ref{tab:table2}.

```{r Simulated_data_gi0, echo=FALSE, message=FALSE}

set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c(9, 25, 49, 81, 121 )

# Number of replications
R <-500

# Number of bootstrap replications
B1 <- 30
mu_values <- c(1, 10)
alpha <- -50
L <- 5

estimators <- list(
  "Correa" = correa_estimator,
  "Ebrahimi" = ebrahimi_estimator,
  "Al Omari" = al_omari_1_estimator,
  "Correa Bootstrap" = bootstrap_correa_estimator,
  "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
  "Al Omari Bootstrap" = bootstrap_al_omari_1_estimator
)

calculate_results_gi0 <- function(sample_sizes, R, B1, mu_values, alpha, L, estimators) {
  results_list <- list()

  for (mu_val in mu_values) {

    results <- calculate_bias_mse_gi0(sample_sizes, R, B1, mu_val, alpha, L, estimators)
    df <- as.data.frame(results)

    results_list[[as.character(mu_val)]] <- df
  }

  return(results_list)
}

results_gi0 <- calculate_results_gi0(sample_sizes, R, B1, mu_values, alpha, L, estimators)

save(results_gi0, file = "./Data/resultsgi0_1.Rdata")

```

```{r Plot_bias_mse_gi0, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.pos="hbt",  fig.cap="Bias and MSE, $L=5$ and  $\\alpha=-10$.", fig.width=8, fig.height=5.5}

load("./Data/resultsgi0_1.Rdata")

estimators_to_plot <- c("Correa", "Ebrahimi", "Al Omari",  "Correa Bootstrap", "Ebrahimi Bootstrap", "Al Omari Bootstrap" )
  latex_estimator_names <- c("Correa" = expression("$\\widehat{italic(H)}_{C}$"),
                             "Correa Bootstrap" = expression("$\\widetilde{italic(H)}_{C}$"),
                            "Ebrahimi" = expression("$\\widehat{italic(H)}_{E}$"),
                            "Al Omari" = expression("$\\widehat{italic(H)}_{AO}$"),
                            "Ebrahimi Bootstrap" = expression("$\\widetilde{italic(H)}_{E}$"),
                            "Al Omari Bootstrap" = expression("$\\widetilde{italic(H)}_{AO}$"))
selected_estimators_latex <- latex_estimator_names[estimators_to_plot]

combined_plot_gi0 <- generate_plot_gi0_esp(results_gi0, mu_values, selected_estimators_latex, ncol = 1, nrow = 2)

print(combined_plot_gi0)

```

```{r Table_gi0, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#cat("\\setlength{\\tabcolsep}{0.5pt}\n")
load("./Data/resultsgi0_1.Rdata")

# Define estimators
estimators_to_table <- c( "Correa", "Ebrahimi", "Al Omari", "Correa Bootstrap","Ebrahimi Bootstrap","Al Omari Bootstrap")

# Filter and reshape data
filtered_results <- purrr::map_dfr(results_gi0, ~ .x %>% filter(Estimator %in% estimators_to_table), .id = "mu")

reshaped_results <- filtered_results %>%
  pivot_wider(names_from = Estimator, values_from = c(Bias, MSE))

colnames(reshaped_results) <- c("$\\mu$", "$n$",  "$\\widehat{H}_{\\text{C}}$", "$\\widehat{H}_{\\text{E}}$","$\\widehat{H}_{\\text{AO}}$",  "$\\widetilde{H}_{\\text{C}}$", "$\\widetilde{H}_{\\text{E}}$","$\\widetilde{H}_{\\text{AO}}$",   "$\\widehat{H}_{\\text{C}}$", "$\\widehat{H}_{\\text{E}}$","$\\widehat{H}_{\\text{AO}}$", "$\\widetilde{H}_{\\text{C}}$", "$\\widetilde{H}_{\\text{E}}$","$\\widetilde{H}_{\\text{AO}}$")

reshaped_results[] <- lapply(reshaped_results, function(x, col_name) {
  if (is.numeric(x)) {
    if (col_name == "\\mu") {
      formatted_numbers <- sprintf("$\\scriptstyle %d$", x)
    } else {
      if (all(x %% 1 == 0)) {
        formatted_numbers <- sprintf("$%d$", x)
      } else {
        formatted_numbers <- ifelse(x < 0, sprintf("$%.3f$", x), sprintf("$\\phantom{-}%.3f$", x))
      }
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
}, col_name = names(reshaped_results)[1])

#print(
 # kbl(
knitr::kable(
    reshaped_results,
    caption = "Bias and MSE of non-parametric entropy estimators and bootstrap versions.",
    format = "latex",
    booktabs = TRUE,
    align = "lrrrrrrrrrrrrrrr",
    escape = FALSE, 
    digits = 2,  
    label = "table2",
    centering = FALSE,
    table.envir = "table", position = "H") %>%
    add_header_above(c(" ", " ", "Bias" = 6,  "MSE" = 6)) %>%
    collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle") %>%
    row_spec(0,  align = "c")%>%
    kable_styling(latex_options = "scale_down")
  #)
  #kable_styling(latex_options = c("scale_down"))

```

## Coefficient of variation and a robust alternative

The population CV is defined as a ratio of the population standard deviation $(\sigma)$ to the population mean $(\mu)$:
\begin{align}
	\text{CV}=\frac{\sigma}{\mu}, \quad \mu \neq 0.
\end{align}

We explore a robust alternative to CV, as describedby&nbsp;\cite{Ospina2019}, which incorporates the ratio between the mean absolute deviation from the median (MnAD) and the median, two well-known robust measures of
scale and location, respectively. 
The sample version for the MnAD is defined as
\(n^{-1}\sum_{i=1}^n|x_i-\widehat{Q}_2|\), where \(\widehat{Q}_2\)
is an estimate for the median of the population, which can be approximated by the median of the sample.

# Hypothesis Testing  {#sec:test}

We aim at testing the following hypotheses: 
$$
\begin{cases}
  \mathcal{H}_0: \text{ The data come from the } \Gamma_{\text{SAR}}\text{ law},\\ 
  \mathcal{H}_1:\text{ The data come from the } G_I^0 \text{ distribution}.
\end{cases}
$$
We are testing the hypothesis that the data are fully-developed speckle versus the alternative of data with roughness.
As for the parametric problem, once it is not possible to define the hypothesis $\mathcal{H}_0=\alpha=-\infty$, it is impossible to solve this problem with parametric  inference alternatives (such as likelihood ratio, score, gradient and Wald hypothesis test).
The proposed tests to solve this physical problem in SAR systems are described below.

## The Proposed Test Based on Non-parametric Entropy

For a random sample \(\bm{Z}=(Z_1, Z_2,\ldots,Z_n)\) from a distribution $\mathcal{D}$, a test statistic is proposed. It is based on an  empirical distribution that arises  from the difference between non-parametrically estimated entropies $\widetilde{H}(\bm{Z})$ and the analytical entropy of Gamma SAR&nbsp;\eqref{E:E-gamma} evaluated at the logarithm of the sample mean, where $L_0\geq 1$ is known.

Hence, the entropy-based test statistic is defined as:
\begin{equation}
\label{Eq:test_e}
S(\bm{Z};L_0)= \widetilde{H}(\bm{Z})-\left[H_{\Gamma_{\text{SAR}}}(L_0)+\ln \widebar{\bm{Z}}\right].
\end{equation}

This test statistic  aims to assess the behavior of the data under the null hypothesis using the empirical distribution.
If the data represent fully-developed speckle, the density should center around zero, i.e., $S(\bm{Z};L_0)\approx 0$. Otherwise, under the alternative hypothesis, the empirical distribution would shift from zero, suggesting significant differences and implying the presence of heterogeneous clutter.

<!-- ACF This is an important contribution. Highlight the rationale: under H0, \widetilde{H} (estimated) and H_{\Gamma_{\text{SAR}}}(L) (fixed) should be close. Therefore, their difference should be small. Explain why you need \widebar{Z} -->
<!-- JA Corrected -->

Figure&nbsp;\ref{fig:Plot_density} shows the empirical test statistic densities for different sample sizes using \(\widetilde{H}_{\text{AO}}\) estimator, and Table&nbsp;\ref{tab:table_stat_combined} summarizes the main descriptive statistics, including mean, SD, variance (Var), skewness (SK), excessive kurtosis (EK) and Anderson--Darling \(p\) values for normality. 
Results with \(p\) values greater than \(0.05\) do not indicate a violation of the normality assumption.
A low variance as well as a skewness and an excessive kurtosis of almost zero indicate limited dispersion, symmetry and a light tail.
Normal QQ plots confirm that there is no evidence against a normal distribution, as shown in Figure&nbsp;\ref{fig:Plot_normality_qq}.

```{r Simulated_density, echo=FALSE, message=FALSE}
#, cache = TRUE, autodep = TRUE
set.seed(1234567890, kind = "Mersenne-Twister")

R <- 2000
mu <- 1
B <- 50

sample.size <- c(25,49, 81, 121)
L_values <- c(3,  5, 8, 11)

all_summary_stats <- list()
all_TestStatistics <- list()

# For each L
for (L in L_values) {
  TestStatistics1 <- list()  
  summary_stats <- data.frame( LValue = character(),
                              SampleSize = numeric(),
                              Mean = numeric(),
                              SD = numeric(),
                              Variance = numeric(),
                              #CV= numeric(),
                              Skewness = numeric(),
                              Kurtosis = numeric(),
                              adpvalue = numeric()
                              )

  for (s in sample.size) {
    TestStat1 <- numeric(R)

    for (r in 1:R) {
      z <- gamma_sar_sample(L, mu, s)
      #z <- gi0_sample(mu, alpha1, L, s)
      TestStat1[r] <-bootstrap_al_omari_1_estimator(z,B) - (log(mean(z)) + (L - log(L) + lgamma(L) + (1 - L) * digamma(L)))
      #TestStat1[r] <- bootstrap_correa_estimator_log_mean(z, B) + (-L + log(L) - lgamma(L) - (1 - L) * digamma(L))
    }

    TestStatistics1[[as.character(s)]] <- data.frame("SampleSize" = rep(s, R), "Test_Statistics" = TestStat1)

    mean_val <- mean(TestStat1)
    sd_val <- sd(TestStat1)
    var_val <- var(TestStat1)
    skewness_val <- skewness(TestStat1)
    kurtosis_val <- kurtosis(TestStat1)
   # lillie_p_value <- lillie.test(TestStatistics1[[as.character(s)]]$Test_Statistics)$p.value
    ad_p_value <- ad.test(TestStatistics1[[as.character(s)]]$Test_Statistics)$p.value
    #cv_val <- abs(sd_val / mean_val)

    summary_stats <- rbind(summary_stats, data.frame(LValue = as.character(L),
                                                     SampleSize = s,
                                                     Mean = mean_val,
                                                     SD = sd_val,
                                                     Variance = var_val,
                                                      #CV = cv_val,
                                                     Skewness = skewness_val,
                                                     Kurtosis = kurtosis_val,
                                                     adpvalue = ad_p_value
                                                     ))

  }

  all_TestStatistics[[as.character(L)]] <- TestStatistics1
  all_summary_stats[[as.character(L)]] <- summary_stats

  save(all_TestStatistics, all_summary_stats, file = paste0("./Data/results1_", L, ".Rdata"))
}

```

```{r Plot_density, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.pos="hbt",  fig.cap="Empirical density  under the null hypothesis.", fig.width=6, fig.height=5.0}

theme_set(theme_minimal() +
            theme(text = element_text(family = "serif"),
                  legend.position = "bottom",
                  legend.text = element_text(angle = 0, vjust = 0.5)))

all_plots <- list()

for (L in L_values) {
  load(paste0("./Data/results1_", L, ".Rdata"))

  combined_data <- do.call(rbind, all_TestStatistics[[as.character(L)]])
  combined_data$L <- L

  p <- ggplot(combined_data, aes(x = Test_Statistics, col = factor(SampleSize), linetype = factor(SampleSize))) +
    geom_line(stat = "density", linewidth = 1.0) +
    scale_color_viridis(discrete = TRUE, option = "C", direction = -1, begin = 0.1, end = 0.8, name = "Sample Size") +
    scale_linetype_manual(values = rep("solid", length(sample.size)), name = "Sample Size") +
    labs(x = "Test Statistics", y = "Density") +
    ggtitle(bquote(italic(L) == .(L))) +
    theme(plot.title = element_text(hjust = 0.5))#titulo centrado

  if (!any(L == L_values)) {
    p = p
  }

  all_plots[[as.character(L)]] <- p
}

combined_plot <- wrap_plots(all_plots, ncol = 2, nrow = 2) +
  plot_layout(guides = "collect")

print(combined_plot)

```

```{r Table_statistic, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#cat("\\setlength{\\tabcolsep}{1.0pt}\n")

load("./Data/results1_3.Rdata")
summary_stats_2 <- all_summary_stats[["3"]]

load("./Data/results1_5.Rdata")
summary_stats_5 <- all_summary_stats[["5"]]

load("./Data/results1_8.Rdata")
summary_stats_8 <- all_summary_stats[["8"]]

load("./Data/results1_11.Rdata")
summary_stats_11 <- all_summary_stats[["11"]]

combined_summary_stats <- rbind(
  transform(summary_stats_2),
  transform(summary_stats_5),
  transform(summary_stats_8),
  transform(summary_stats_11)

)
colnames(combined_summary_stats) <- c("$\\bm{L}$", "$\\bm{n}$", "\\textbf{Mean}", "\\textbf{SD}", "\\textbf{Var}", "\\textbf{SK}", "\\textbf{EK}", "$p$-\\textbf{value}")

# LaTeX notation 
combined_summary_stats[] <- lapply(combined_summary_stats, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.4f$", x), sprintf("$\\phantom{-}%.4f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

table_result_combined <- knitr::kable(
  combined_summary_stats,
  caption = "Descriptive analysis of $S(\\bm Z; L)$ with $L\\in\\left\\{3,5, 8,11\\right\\}$ and $\\mu=1$.",
  format = "latex",
  booktabs = TRUE,
  align = "lrrrrrrr",
  escape = FALSE,
  digits = 4,
  label = "table_stat_combined",
  centering = FALSE,
  table.envir = "table", position = "H") %>%
  #table.env = 'table*' # to span multiple columns
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")

print(table_result_combined)

```

```{r Plot_normality_qq, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.pos="hbt",  fig.cap="Normal QQ-plots for  $n=121$.", fig.width=6, fig.height=5}

theme_set(theme_minimal() +
            theme(text = element_text(family = "serif"),
                  legend.position = "bottom",
                  legend.text = element_text(angle = 0, vjust = 0.5)))

# Load saved results
load("./Data/results1_3.Rdata")
load("./Data/results1_5.Rdata")
load("./Data/results1_8.Rdata")
load("./Data/results1_11.Rdata")

# Specific sample size
selected_sample_size <- 121

# 
ggplot_list <- list()

#  L
for (L in c(3, 5, 8, 11)) {
  TestStatistics1 <- all_TestStatistics[[as.character(L)]]

  # 
  if (as.character(selected_sample_size) %in% names(TestStatistics1)) {
    p <- ggplot(TestStatistics1[[as.character(selected_sample_size)]], 
                aes(sample = Test_Statistics)) +
      geom_qq(col = "blue", shape = 20) +
      geom_qq_line(col = "darkgreen", linetype = 2) +
      labs(title = bquote(italic(L) == .(L)),
           x = "Theoretical Quantiles",
           y = "Sample Quantiles") +
      theme(plot.title = element_text(hjust = 0.5))#titulo centrado

    ggplot_list[[as.character(L)]] <- p
  } else {
    warning(paste("No size sample", selected_sample_size, " L =", L))
  }
}
combined_plot1 <- wrap_plots(ggplot_list, ncol = 2, nrow = 2) +
  plot_layout(guides = "collect")

print(combined_plot1)

# }

```

After checking the normality of the test statistics, we examined the possibilities of the proposed test in terms of size and power.
Under \(\mathcal{H}_0\), the distribution of the test statistic is asymptotically normal.
Therefore, the \(p\) value is calculated as
\(2\Phi(-|\varepsilon|)\), where \(\Phi\) is the standard
Gaussian cumulative distribution function, and \(\varepsilon\) is the standardized test statistic given by :
\[
\varepsilon=\frac{\widetilde{H}-H_{\Gamma_{\text{SAR}}}}{\hat\sigma}.
\]

We have nominal levels of \SI{1}{\percent}, \SI{5}{\percent}, and
\SI{10}{\percent}. 
In terms of size, \(1000\) simulations were used for different sample sizes from the $\Gamma_{\text{SAR}}$ distribution, with different values of $L$,  $\mu=1$ and \(\widetilde{H}_{\text{AO}}\) estimator. 
In all cases, the nominal level was achieved. 
We assessed the test power using \(1000\) simulations for different sample sizes from the $G_I^0$ distribution, with  $\mu=1$, $\alpha=-2$, and \(\widetilde{H}_{\text{AO}}\) estimator.
The power generally improves with increasing sample size and number of looks.
The results are shown in Table&nbsp;\ref{tab:table_size_power}.
<!-- ACF Please clarify which alternatives you used for the power -->
<!-- JA Corrected -->

```{r Simulated_error_type_I, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")
# 
calculate_p_value <- function(test_statistic, mu_W, sigma_W, alpha_nominal) {
  epsilon <- test_statistic / sigma_W
  p_value <- 2 * (1 - pnorm(abs(epsilon)))

  return(p_value < alpha_nominal)  
}
R <- 1000
mu <- 1
B <- 100
L_values <- c( 3, 5, 8, 11)
sample_sizes <- c( 25, 49, 81, 121)
alpha_nominals <- c(0.01, 0.05, 0.1)
results <- data.frame()
for (L in L_values) {
  for (alpha_nominal in alpha_nominals) {
    TestStatistics <- NULL
    mean_entropy <- numeric(length(sample_sizes))
    sd_entropy <- numeric(length(sample_sizes))

    for (s in sample_sizes) {
      TestStat <- numeric(R)

      for (r in 1:R) {
        z <- gamma_sar_sample(L, mu, s)
        TestStat[r] <-bootstrap_al_omari_1_estimator(z,B) - (log(mean(z)) + (L - log(L) + lgamma(L) + (1 - L) * digamma(L)))
        #TestStat[r] <- bootstrap_correa_estimator(z, B) - (L - log(L) + lgamma(L) + (1 - L) * digamma(L))
      }
      mean_entropy[sample_sizes == s] <- mean(TestStat)
      sd_entropy[sample_sizes == s] <- sd(TestStat)

      TestStatistics <- rbind(TestStatistics, data.frame("L" = rep(L, R), "Sample_Size" = rep(s, R), "Test_Statistics" = TestStat))
    }

    mu_W <- mean_entropy
    sigma_W <- sqrt(sd_entropy^2)

    p_values <- apply(TestStatistics, 1, function(row) {
      calculate_p_value(row["Test_Statistics"], mu_W[sample_sizes == row["Sample_Size"]], sigma_W[sample_sizes == row["Sample_Size"]], alpha_nominal)
    })

    result <- data.frame("L" = TestStatistics$L, "Sample_Size" = TestStatistics$Sample_Size, "Alpha_Nominal" = alpha_nominal, "P_Value" = p_values)
    results <- rbind(results, result)
  }
}
save(results, file = "./Data/type_I_results.Rdata")
```

```{r Simulated_power, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")
calculate_p_value <- function(test_statistic, mu_W, sigma_W) {
  epsilon <- test_statistic  / sigma_W
  p_value <- 2 * (1 - pnorm(abs(epsilon)))

  return(p_value)
}
calculate_type_II_error_rate <- function(p_values, alpha_nominal) {
  type_II_error_rate <- sum(p_values >= alpha_nominal) / length(p_values)
  return(type_II_error_rate)
}
# Function to calculate power for different L values and alpha_nominals
calculate_power <- function(R, mu, L_values, B, sample_sizes, alpha_nominals) {
  results <- data.frame()

  for (L in L_values) {
    for (alpha_nominal in alpha_nominals) {
      TestStatistics <- list()
      mean_entropy <- numeric(length(sample_sizes))
      sd_entropy <- numeric(length(sample_sizes))

      for (s in sample_sizes) {
        TestStat <- numeric(R)

        for (r in 1:R) {
          z <- gi0_sample(mu, -2, L, s)
          TestStat[r] <-bootstrap_al_omari_1_estimator(z,B) - (log(mean(z)) + (L - log(L) + lgamma(L) + (1 - L) * digamma(L)))
          #TestStat[r] <- bootstrap_correa_estimator_log_mean(z, B) - (L - log(L) + lgamma(L) + (1 - L) * digamma(L))
        }

        mean_entropy[sample_sizes == s] <- mean(TestStat)
        sd_entropy[sample_sizes == s] <- sd(TestStat)

        TestStatistics[[as.character(s)]] <- TestStat
      }

      mu_W <-  mean_entropy  
      sigma_W <- sqrt(sd_entropy^2)

      p_values <- lapply(TestStatistics, function(TestStat) {
        apply(data.frame("Test_Statistics" = TestStat), 1, function(row) {
          calculate_p_value(row["Test_Statistics"], mu_W, sigma_W[as.numeric(names(TestStatistics)) == as.numeric(s)])
        })
      })

      type_II_error_rates <- sapply(p_values, function(p_values_for_size) {
        calculate_type_II_error_rate(p_values_for_size, alpha_nominal)
      })

      power <- 1 - type_II_error_rates

      result_row <- data.frame(
        L = L,
        alpha_nominal = alpha_nominal,
        Sample_Size = sample_sizes,
        power = power
      )

      results <- rbind(results, result_row)
    }
  }

  return(results)
}
R <- 1000
mu <- 1
L_values <- c(3, 5,  8, 11)
B <- 100
sample_sizes <- c(25, 49, 81, 121)
alpha_nominals <- c(0.01, 0.05, 0.1)
results_power <- calculate_power(R, mu, L_values, B, sample_sizes, alpha_nominals)
save(results_power, file = "./Data/results_power.Rdata")
```

```{r Table_size_and_power, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#cat("\\setlength{\\tabcolsep}{4pt}\n")
load("./Data/type_I_results.Rdata")
type_I_error_rates <- tapply(results$P_Value, INDEX = list(results$L, results$Sample_Size, results$Alpha_Nominal), FUN = function(p_values) {
  sum(p_values) / length(p_values)  
})
type_I_error_results <- as.data.frame(as.table(type_I_error_rates))
colnames(type_I_error_results) <- c("L", "Sample_Size", "Alpha_Nominal", "Type_I_Error_Rate")
spread_results <- spread(type_I_error_results, key = Alpha_Nominal, value = Type_I_Error_Rate)
spread_results <- spread_results %>% 
  select(L, Sample_Size, `0.01`, `0.05`, `0.1`)
load("./Data/results_power.Rdata")
summary_stats <- results_power %>%
  spread(key = alpha_nominal, value = power) %>%
  select(L, Sample_Size, `0.01`, `0.05`, `0.1`)
combined_results <- merge(spread_results, summary_stats, by = c("L", "Sample_Size"))
combined_results <- combined_results %>% 
  arrange(L, Sample_Size)
combined_results <- combined_results %>% 
  select(L, Sample_Size, `0.01.x`, `0.05.x`, `0.1.x`, `0.01.y`, `0.05.y`, `0.1.y`)
colnames(combined_results) <- c("$L$", "$n$", "$1\\%$", "$5\\%$", "$10\\%$", "$1\\%$", "$5\\%$", "$10\\%$")
# LaTeX notation 
combined_results[] <- lapply(combined_results, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.4f$", x), sprintf("$\\phantom{-}%.4f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})
table_combined_result <- knitr::kable(
  combined_results,
  caption = "Size and Power of the entropy-based test.",
  format = "latex",
  booktabs = TRUE,
  align = "lcccccccc",
  escape = FALSE,
  digits = 4,
  label = "table_size_power",
  centering = FALSE,
  #linesep = "",
  table.envir = "table", position = "H")%>%
  add_header_above(c(" ", " ", "Size" = 3,  "Power" = 3)) %>%
  collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")
print(table_combined_result)
```

## The Proposed Test Based on Coefficient of Variation and a Robust Alternative

In addition to the entropy-based test, we also propose a test statistic based on the classical CV. This test statistic is defined as follows:
\begin{align}
	T_{\text{CV}}=\frac{S}{\widebar{Z}},
\end{align}
where $S$ and $\widebar{Z}$ are the sample standard deviation and the sample mean, respectively.

Similarly, we use another test statistic based on the ratio of the MnAD to the median. This statistic is given by:
\begin{align}
	T_{\text{CV}_{\text{MnAD}}}=\frac{\text{MnAD}}{\text{Median}}.
\end{align}
These test statistics allow assessing the variability of the data.

The situations in which the use of CV and \(\text{CV}_{\text{MnAD}}\) may be appropriate, i.e. when the observations are positive, the log-normal
(LN) and the inverse Gaussian distribution (IG) are often more appropriate than the Gamma and Weibull distributions&nbsp;\cite{Chaubey2017,takagi1997application}.

It is shown that the IG distribution is well approximated by the log-normal distribution, which means that the
IG distribution also does not share the problem of the non-existence of a fixed-width confidence interval with the Gaussian case&nbsp;\cite{whitmore1978}.

The LN distribution in two parameters has the following density function.

\begin{align}
    f_Z(z;\mu_{\text{LN}}, \sigma_{\text{LN}} )=\frac{1}{\sigma_{\text{LN}} z\sqrt{2\pi}}\exp\left\{-\frac{(\ln z- \mu_{\text{LN}})^2}{2\sigma_{\text{LN}}^2}\right\}\mathbbm 1_{\mathbbm R_+}(z),
\end{align}  
here, $\mu_{\text{LN}}$ and $\sigma_{\text{LN}}$ are the mean and standard deviation of the logarithm.

## Model Selection Criterion

Akaike information criterion (AIC) and Bayesian information criterion (BIC) are used to select the best fitting distribution.

The AIC is a fine technique that deals with the trade-off between the goodness-of-fit of the model and the simplicity of the model in terms of the number of model parameters&nbsp;\cite{Burnham2004}. 
The model or distribution with the lowest value of AIC is chosen to be the best.
The BIC assesses goodness-of-fit of a distribution or model, but avoids overfitting by penalising an additional degree of freedom&nbsp;\cite{Dziak2019}. 
The model with the lowest BIC value is chosen as the best.

The AIC and BIC results in Tables&nbsp;\ref{tab:table_aic_gamma}--\ref{tab:table_aic_gio_MnADmedian} indicate that the CV and \(\text{CV}_{\text{MnAD}}\) data from different distributed Gamma and \(G_I^0\) synthetic sample sizes match the properties of an LN distribution.
It is important to note that this conclusion was drawn empirically based on a dictionary of distributions that are analytically tractable and well-defined under biparametric, unimodal, asymmetric, and positive distributions.

<!-- \textcolor{red}{% -->
<!-- ADCN: Caros, esse é o ponto mais delicado da excelente ideia desse paper em minha opinião. Não me oponho a colocarmos a lognormal como uma possibilidade, mas é preciso deixar claro que é uma possibilidade empírica. Uma alternativa empírica para uma solução do momento usando distribuições mais desenvolvidas na literatura e que sejam analiticamente tratáveis. Estou pensando corretamente? Se sim, por que não log-t-student, log-Laplace, power exponential, Log-hyperbolic,.... (e a lista segue). Sugiro talvez indicar (i) a dedução da distribuição assintótica das estatísticas em função de CV e MNAD dando com entrada G0I ($\mathcal{H}_1$) e a Gamma ($\mathcal{H}_0$) como trabalho futuro. Do que eu entendi, a suposição da normal assintótica para entropia vem de uma coleção de trabalhos teóricos. Já a LN, do que percebi, não. Coloquei um comentário em verde. Vejam se estão de acordo. -->
<!-- } -->

Figures&nbsp;\ref{fig:Plot_cv}--\ref{fig:Plot_madmed_gi0_MnADmedian} show empirical and fitted density plots, Q-Q plots, P-P plots as well as empirical and fitted cumulative distribution functions.
They provide qualitative sources that confirm that the LN distribution is the most appropriate distribution

```{r Simulated_AIC_BIC_gamma, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")
R <- 15000
mu <- 1
L <- 5
#B <- 5
#alpha1 <- -3
sample.sizes <- c(25, 49, 81, 121)

results_AIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

results_BIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

for (s in sample.sizes) {
  TestStatistics <- numeric(0)

  for (r in 1:R) {
    z <- gamma_sar_sample(L, mu, s)
    #z <- gi0_sample(mu, alpha1, L, s)
    TestStat <- sd(z) / mean(z)
    TestStatistics <- c(TestStatistics, TestStat)
  }

  fit_weibull <- fitdistr(TestStatistics, "weibull")
  fit_lognormal <- fitdistr(TestStatistics, "lognormal")
  fit_gamma <- fitdistr(TestStatistics, "gamma")
  fit_normal <- fitdistr(TestStatistics, "normal")

  AIC_values <- c(AIC(fit_weibull), AIC(fit_lognormal), AIC(fit_gamma), AIC(fit_normal))
  BIC_values <- c(BIC(fit_weibull), BIC(fit_lognormal), BIC(fit_gamma), BIC(fit_normal))

  dinvgauss <- function(x, mu, lambda) {
    sqrt(lambda / (2 * pi * x^3)) * exp(-lambda * (x - mu)^2 / (2 * mu^2 * x))
  }

  negLL_invgauss <- function(params, data) {
    mu <- params[1]
    lambda <- params[2]
    CV <- data

    -sum(log(dinvgauss(CV, mu, lambda)))
  }

  fit_invGauss <- fitdist(data = TestStatistics, distr = "invgauss", method = "mle",
                          start = list(mu = mean(TestStatistics), lambda = var(TestStatistics)))

  log_likelihood <- logLik(fit_invGauss)
  num_params <- length(coef(fit_invGauss))
  n <- length(TestStatistics)
  AIC_manual <- -2 * log_likelihood + 2 * num_params
  BIC_manual <- -2 * log_likelihood + num_params * log(n)

  results_AIC <- rbind(results_AIC, data.frame(Test = "AIC",
                                               SampleSize = s,
                                               Normal = AIC_values[4],
                                               Lognormal = AIC_values[2],
                                               Gamma = AIC_values[3],
                                               Weibull = AIC_values[1],
                                               InverseGaussian = AIC_manual))

  results_BIC <- rbind(results_BIC, data.frame(Test = "BIC",
                                               SampleSize = s,
                                               Normal = BIC_values[4],
                                               Lognormal = BIC_values[2],
                                               Gamma = BIC_values[3],
                                               Weibull = BIC_values[1],
                                               InverseGaussian = BIC_manual))
}

save(results_AIC, results_BIC, file = "./Data/results_AIC_BIC_gamma.Rdata")

```

```{r Table_AIC_BIC_gamma, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

load("./Data/results_AIC_BIC_gamma.Rdata")

combined <- rbind(results_AIC, results_BIC)
rownames(combined) <- NULL

colnames(combined) <- c("\\textbf{Criterion}", "$\\bm{n}$", "\\textbf{Normal}", "\\textbf{Lognormal}", "\\textbf{Gamma}", "\\textbf{Weibull}", "\\textbf{Inverse Gaussian}")

#combined$`Inverse Gaussian` <- c("\\multirow{2}{*}{Inverse}", "Gaussian")

# LaTeX notation 
combined[] <- lapply(combined, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.2f$", x), sprintf("$\\phantom{-}%.2f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

table_result_combined <- knitr::kable(
  combined,
  caption = "AIC and BIC values for evaluating the best distribution with CV data from Gamma SAR.",
  format = "latex",
  booktabs = TRUE,
  align = "lccccccc",
  escape = FALSE,
  digits = 2,
  label = "table_aic_gamma",
  centering = FALSE,
  table.envir = "table", position = "H") %>%
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")

print(table_result_combined)
```

```{r Simulated_AIC_BIC_GI0, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")
R <- 15000
mu <- 1
L <- 5
B <- 5
alpha1 <- -3
sample.sizes <- c(25, 49, 81, 121)

# Simulación y cálculo de resultados
results_AIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

results_BIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

for (s in sample.sizes) {
  TestStatistics <- numeric(0)

  for (r in 1:R) {
    z <- gi0_sample(mu, alpha1, L, s)
    TestStat <- sd(z) / mean(z)
    TestStatistics <- c(TestStatistics, TestStat)
  }

  fit_weibull <- fitdistr(TestStatistics, "weibull")
  fit_lognormal <- fitdistr(TestStatistics, "lognormal")
  fit_gamma <- fitdistr(TestStatistics, "gamma")
  fit_normal <- fitdistr(TestStatistics, "normal")

  AIC_values <- c(AIC(fit_weibull), AIC(fit_lognormal), AIC(fit_gamma), AIC(fit_normal))
  BIC_values <- c(BIC(fit_weibull), BIC(fit_lognormal), BIC(fit_gamma), BIC(fit_normal))

  dinvgauss <- function(x, mu, lambda) {
    sqrt(lambda / (2 * pi * x^3)) * exp(-lambda * (x - mu)^2 / (2 * mu^2 * x))
  }

  negLL_invgauss <- function(params, data) {
    mu <- params[1]
    lambda <- params[2]
    CV <- data

    -sum(log(dinvgauss(CV, mu, lambda)))
  }

  fit_invGauss <- fitdist(data = TestStatistics, distr = "invgauss", method = "mle",
                          start = list(mu = mean(TestStatistics), lambda = var(TestStatistics)))

  log_likelihood <- logLik(fit_invGauss)
  num_params <- length(coef(fit_invGauss))
  n <- length(TestStatistics)
  AIC_manual <- -2 * log_likelihood + 2 * num_params
  BIC_manual <- -2 * log_likelihood + num_params * log(n)

  results_AIC <- rbind(results_AIC, data.frame(Test = "AIC",
                                               SampleSize = s,
                                               Normal = AIC_values[4],
                                               Lognormal = AIC_values[2],
                                               Gamma = AIC_values[3],
                                               Weibull = AIC_values[1],
                                               InverseGaussian = AIC_manual))

  results_BIC <- rbind(results_BIC, data.frame(Test = "BIC",
                                               SampleSize = s,
                                               Normal = BIC_values[4],
                                               Lognormal = BIC_values[2],
                                               Gamma = BIC_values[3],
                                               Weibull = BIC_values[1],
                                               InverseGaussian = BIC_manual))
}

# Guardar los resultados en un archivo .Rdata
save(results_AIC, results_BIC, file = "./Data/results_AIC_BIC.Rdata")

```

```{r Table_AIC_BIC, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

load("./Data/results_AIC_BIC.Rdata")

combined <- rbind(results_AIC, results_BIC)
rownames(combined) <- NULL

colnames(combined) <- c("\\textbf{Criterion}", "$\\bm{n}$", "\\textbf{Normal}", "\\textbf{Lognormal}", "\\textbf{Gamma}", "\\textbf{Weibull}", "\\textbf{Inverse Gaussian}")

# Modificar el nombre de la última columna para que esté en dos líneas
#combined$`Inverse Gaussian` <- c("\\multirow{2}{*}{Inverse}", "Gaussian")

# LaTeX notation 
combined[] <- lapply(combined, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.2f$", x), sprintf("$\\phantom{-}%.2f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

table_result_combined <- knitr::kable(
  combined,
  caption = "AIC and BIC values for evaluating the best distribution with CV data from $G_I^0$.",
  format = "latex",
  booktabs = TRUE,
  align = "lccccccc",
  escape = FALSE,
  digits = 2,
  label = "table_aic",
  centering = FALSE,
  table.envir = "table", position = "H") %>%
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")

print(table_result_combined)
```

```{r Simulated_AIC_BIC_gamma_MnADmedian, echo=FALSE, message=FALSE, warning=FALSE}

set.seed(1234567890, kind = "Mersenne-Twister")
R <- 15000
mu <- 1
L <- 5
sample.sizes <- c(25, 49, 81, 121)

# madmedian <- function(x){
#   mdn <- median(x)
#   MAD <- 1.4826 * median(abs(x - mdn))
#   MAD/mdn
# }

MnADmedian <- function(x){
  mdn <- median(x)
  MnAD <- mean(abs(x - mdn))
  MnAD/mdn
}

results_AIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

results_BIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

for (s in sample.sizes) {
  TestStatistics <- numeric(0)

  for (r in 1:R) {
    z <- gamma_sar_sample(L, mu, s)
    # z <- gi0_sample(mu, alpha1, L, s)
    madmed <- MnADmedian(z)  
    TestStatistics <- c(TestStatistics, madmed)
  }

  fit_weibull <- fitdistr(TestStatistics, "weibull")
  fit_lognormal <- fitdistr(TestStatistics, "lognormal")
  fit_gamma <- fitdistr(TestStatistics, "gamma")
  fit_normal <- fitdistr(TestStatistics, "normal")

  AIC_values <- c(AIC(fit_weibull), AIC(fit_lognormal), AIC(fit_gamma), AIC(fit_normal))
  BIC_values <- c(BIC(fit_weibull), BIC(fit_lognormal), BIC(fit_gamma), BIC(fit_normal))

  dinvgauss <- function(x, mu, lambda) {
    sqrt(lambda / (2 * pi * x^3)) * exp(-lambda * (x - mu)^2 / (2 * mu^2 * x))
  }

  negLL_invgauss <- function(params, data) {
    mu <- params[1]
    lambda <- params[2]
    CV <- data

    -sum(log(dinvgauss(CV, mu, lambda)))
  }

  fit_invGauss <- fitdist(data = TestStatistics, distr = "invgauss", method = "mle",
                          start = list(mu = mean(TestStatistics), lambda = var(TestStatistics)))

  log_likelihood <- logLik(fit_invGauss)
  num_params <- length(coef(fit_invGauss))
  n <- length(TestStatistics)
  AIC_manual <- -2 * log_likelihood + 2 * num_params
  BIC_manual <- -2 * log_likelihood + num_params * log(n)

  results_AIC <- rbind(results_AIC, data.frame(Test = "AIC",
                                               SampleSize = s,
                                               Normal = AIC_values[4],
                                               Lognormal = AIC_values[2],
                                               Gamma = AIC_values[3],
                                               Weibull = AIC_values[1],
                                               InverseGaussian = AIC_manual))

  results_BIC <- rbind(results_BIC, data.frame(Test = "BIC",
                                               SampleSize = s,
                                               Normal = BIC_values[4],
                                               Lognormal = BIC_values[2],
                                               Gamma = BIC_values[3],
                                               Weibull = BIC_values[1],
                                               InverseGaussian = BIC_manual))
}

save(results_AIC, results_BIC, file = "./Data/results_AIC_BIC_gamma_MnADmedian.Rdata")

```

```{r Table_AIC_BIC_gamma_MnADmedian, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

load("./Data/results_AIC_BIC_gamma_MnADmedian.Rdata")

combined <- rbind(results_AIC, results_BIC)
rownames(combined) <- NULL

colnames(combined) <- c("\\textbf{Criterion}", "$\\bm{n}$", "\\textbf{Normal}", "\\textbf{Lognormal}", "\\textbf{Gamma}", "\\textbf{Weibull}", "\\textbf{Inverse Gaussian}")

#combined$`Inverse Gaussian` <- c("\\multirow{2}{*}{Inverse}", "Gaussian")

# LaTeX notation 
combined[] <- lapply(combined, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.2f$", x), sprintf("$\\phantom{-}%.2f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

table_result_combined <- knitr::kable(
  combined,
  caption = "AIC and BIC values for evaluating the best distribution with $\\text{CV}_{\\text{MnAD}}$ data from Gamma SAR.",
  format = "latex",
  booktabs = TRUE,
  align = "lccccccc",
  escape = FALSE,
  digits = 2,
  label = "table_aic_gamma_madmed",
  centering = FALSE,
  table.envir = "table", position = "H") %>%
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")

print(table_result_combined)
```

```{r Simulated_AIC_BIC_GI0_MnADmedian, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")
R <- 15000
mu <- 1
L <- 5
B <- 5
alpha1 <- -5
sample.sizes <- c(25, 49, 81, 121)

# madmedian <- function(x){
#   mdn <- median(x)
#   MAD <- 1.4826 * median(abs(x - mdn))
#   return(MAD/mdn)
# }

MnADmedian <- function(x){
  mdn <- median(x)
  MnAD <- mean(abs(x - mdn))
  MnAD/mdn
}
results_AIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

results_BIC <- data.frame(Test = character(),
                          SampleSize = integer(),
                          Normal = numeric(),
                          Lognormal = numeric(),
                          Gamma = numeric(),
                          Weibull = numeric(),
                          InverseGaussian = numeric())

for (s in sample.sizes) {
  TestStatistics <- numeric(0)

  for (r in 1:R) {
    z <- gi0_sample(mu, alpha1, L, s)
    madmed <- MnADmedian(z) 
    TestStatistics <- c(TestStatistics, madmed)
  }

  fit_weibull <- fitdistr(TestStatistics, "weibull")
  fit_lognormal <- fitdistr(TestStatistics, "lognormal")
  fit_gamma <- fitdistr(TestStatistics, "gamma")
  fit_normal <- fitdistr(TestStatistics, "normal")

  AIC_values <- c(AIC(fit_weibull), AIC(fit_lognormal), AIC(fit_gamma), AIC(fit_normal))
  BIC_values <- c(BIC(fit_weibull), BIC(fit_lognormal), BIC(fit_gamma), BIC(fit_normal))

  dinvgauss <- function(x, mu, lambda) {
    sqrt(lambda / (2 * pi * x^3)) * exp(-lambda * (x - mu)^2 / (2 * mu^2 * x))
  }

  negLL_invgauss <- function(params, data) {
    mu <- params[1]
    lambda <- params[2]
    CV <- data

    -sum(log(dinvgauss(CV, mu, lambda)))
  }

  fit_invGauss <- fitdist(data = TestStatistics, distr = "invgauss", method = "mle",
                          start = list(mu = mean(TestStatistics), lambda = var(TestStatistics)))

  log_likelihood <- logLik(fit_invGauss)
  num_params <- length(coef(fit_invGauss))
  n <- length(TestStatistics)
  AIC_manual <- -2 * log_likelihood + 2 * num_params
  BIC_manual <- -2 * log_likelihood + num_params * log(n)

  results_AIC <- rbind(results_AIC, data.frame(Test = "AIC",
                                               SampleSize = s,
                                               Normal = AIC_values[4],
                                               Lognormal = AIC_values[2],
                                               Gamma = AIC_values[3],
                                               Weibull = AIC_values[1],
                                               InverseGaussian = AIC_manual))

  results_BIC <- rbind(results_BIC, data.frame(Test = "BIC",
                                               SampleSize = s,
                                               Normal = BIC_values[4],
                                               Lognormal = BIC_values[2],
                                               Gamma = BIC_values[3],
                                               Weibull = BIC_values[1],
                                               InverseGaussian = BIC_manual))
}

save(results_AIC, results_BIC, file = "./Data/results_AIC_BIC_GI0_MnADmedian.Rdata")

```

```{r Table_AIC_BIC_GI0_MnADmedian, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

load("./Data/results_AIC_BIC_GI0_MnADmedian.Rdata")

combined <- rbind(results_AIC, results_BIC)
rownames(combined) <- NULL

colnames(combined) <- c("\\textbf{Criterion}", "$\\bm{n}$", "\\textbf{Normal}", "\\textbf{Lognormal}", "\\textbf{Gamma}", "\\textbf{Weibull}", "\\textbf{Inverse Gaussian}")

#combined$`Inverse Gaussian` <- c("\\multirow{2}{*}{Inverse}", "Gaussian")

# LaTeX notation 
combined[] <- lapply(combined, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.2f$", x), sprintf("$\\phantom{-}%.2f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

table_result_combined <- knitr::kable(
  combined,
  caption = "AIC and BIC values for evaluating the best distribution with $\\text{CV}_{\\text{MnAD}}$ data from $G_I^0$.",
  format = "latex",
  booktabs = TRUE,
  align = "lccccccc",
  escape = FALSE,
  digits = 2,
  label = "table_aic_gio_MnADmedian",
  centering = FALSE,
  table.envir = "table", position = "H") %>%
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")

print(table_result_combined)
```

```{r Simulated_cv_gamma, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

sample.size <- c(49)
R <- 15000
mu <- 1
L <- 5
B <- 5
alpha1 <- -2

TestStatistics <- numeric(0)

for (s in sample.size) {
  for (r in 1:R) {
    z <- gamma_sar_sample(L, mu, s)
    #z <- gi0_sample(mu, alpha1, L, s)
    TestStat <- sd(z) / mean(z)
    TestStatistics <- c(TestStatistics, TestStat)  
  }
}

save(TestStatistics, file = "./Data/rTestStatistics_gamma_49.Rdata")
```

```{r Plot_cv_gamma, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H',  fig.cap="Goodness of fit plots for evaluating the best distribution with CV data from Gamma SAR, with  $n=49$, $L=5$ and $\\mu=1$.", fig.width=8, fig.height=6}

load("./Data/rTestStatistics_gamma_49.Rdata")

# Ajustar las distribuciones a los datos
fln <- fitdist(TestStatistics, "lnorm")
fg <- fitdist(TestStatistics, "gamma")
fn <- fitdist(TestStatistics, "norm")  # Ajuste para la distribución normal
fw <- fitdist(TestStatistics, "weibull")

# Graficar las comparaciones
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
plot.legend <- c( "Lognormal", "Gamma", "Normal", "Weibull") 
denscomp(list( fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3), fitlty = 1,xlab = "CV")
qqcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)
cdfcomp(list(fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3), fitlty = 1,xlab = "CV")
ppcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)

```

```{r Simulated_cv_gi0, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

sample.size <- c(49)
R <- 15000
mu <- 1
L <- 5
B <- 5
alpha1 <- -2

TestStatistics <- numeric(0)

for (s in sample.size) {
  for (r in 1:R) {
    z <- gi0_sample(mu, alpha1, L, s)
    TestStat <- sd(z) / mean(z)
    TestStatistics <- c(TestStatistics, TestStat)  
  }
}

save(TestStatistics, file = "./Data/rTestStatistics_gi0_49.Rdata")
```

```{r Plot_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H',  fig.cap="Goodness of fit plots for evaluating the best distribution with $\\text{CV}_{\\text{MnAD}}$ data from $G_I^0$, with  $n=49$, $L=5$, $\\mu=1$ and $\\alpha=-2$.", fig.width=8, fig.height=6}

load("./Data/rTestStatistics_gi0_49.Rdata")

#
fln <- fitdist(TestStatistics, "lnorm")
fg <- fitdist(TestStatistics, "gamma")
fn <- fitdist(TestStatistics, "norm")  
fw <- fitdist(TestStatistics, "weibull")

# 
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
plot.legend <- c( "Lognormal", "Gamma", "Normal", "Weibull") 
denscomp(list( fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3), fitlty = 1,xlab = "CV")
qqcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)
cdfcomp(list(fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3), fitlty = 1,xlab = "CV")
ppcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)

```

```{r Simulated_MnADmedian_gamma, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

sample.size <- c(49)
R <- 15000
mu <- 1
L <- 5
B <- 5
alpha1 <- -2

MnADmedian <- function(x){
  mdn <- median(x)
  MnAD <- mean(abs(x - mdn))
  MnAD/mdn
}

TestStatistics <- numeric(0)

for (s in sample.size) {
  for (r in 1:R) {
    z <- gamma_sar_sample(L, mu, s)
    #z <- gi0_sample(mu, alpha1, L, s)
    TestStat <- MnADmedian(z)
    TestStatistics <- c(TestStatistics, TestStat)  
  }
}

save(TestStatistics, file = "./Data/rTestStatistics_gamma_49_MnADmedian.Rdata")
```

```{r Plot_MnADmedian_gamma, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H',  fig.cap="Goodness of fit plots for evaluating the best distribution with $\\text{CV}_{\\text{MnAD}}$ data from Gamma SAR, with  $n=49$, $L=5$ and $\\mu=1$.", fig.width=8, fig.height=6}

load("./Data/rTestStatistics_gamma_49_MnADmedian.Rdata")

# Ajustar las distribuciones a los datos
fln <- fitdist(TestStatistics, "lnorm")
fg <- fitdist(TestStatistics, "gamma")
fn <- fitdist(TestStatistics, "norm")  # Ajuste para la distribución normal
fw <- fitdist(TestStatistics, "weibull")

# Graficar las comparaciones
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
plot.legend <- c( "Lognormal", "Gamma", "Normal", "Weibull") 
denscomp(list( fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3), fitlty = 1,xlab = "CV")
qqcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)
cdfcomp(list(fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3), fitlty = 1,xlab = "CV")
ppcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)

```

```{r Simulated_MnADmedian_gi0, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

sample.size <- c(49)
R <- 12000
mu <- 1
L <- 5
#B <- 5
alpha1 <- -5

MnADmedian <- function(x){
  mdn <- median(x)
  MnAD <- mean(abs(x - mdn))
  MnAD/mdn
}

TestStatistics_mad <- numeric(0)

for (s in sample.size) {
  for (r in 1:R) {
    z <- gi0_sample(mu, alpha1, L, s)
    madmed <- MnADmedian(z)
    TestStatistics_mad <- c(TestStatistics_mad, madmed)  
  }
}

save(TestStatistics_mad, file = "./Data/rTestStatistics_MnADmedian_gi0_49.Rdata")
```

```{r Plot_madmed_gi0_MnADmedian, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.pos="hbt",  fig.cap="Goodness of fit plots for evaluating the best distribution with $CV_{\\text{MnAD}}$ data from $G_I^0$, with  $n=49$, $L=5$, $\\mu=1$ and $\\alpha=-5$.", fig.width=8, fig.height=6}

load("./Data/rTestStatistics_MnADmedian_gi0_49.Rdata")

#
fln <- fitdist(TestStatistics_mad, "lnorm")
fg <- fitdist(TestStatistics_mad, "gamma")
fn <- fitdist(TestStatistics_mad, "norm")  
fw <- fitdist(TestStatistics_mad, "weibull")

# 
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
plot.legend <- c( "Lognormal", "Gamma", "Normal", "Weibull") 
denscomp(list( fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3), fitlty = 1,xlab = "CV")
qqcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)
cdfcomp(list(fln, fg, fn,fw), legendtext = plot.legend, fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3), fitlty = 1,xlab = "CV")
ppcomp(list(fln, fg, fn,fw), legendtext = plot.legend,fitcol = c("darkblue", "#637029","indianred3","darkgrey"), fitlwd=c(3,3,3,3),xlegend = "topleft", ylegend = NULL)

```

# Simulation Results {#sec:Results}

This section presents simulations that were performed to obtain an evaluation of the performance of the proposed test statistics.

## Simulated Data

Figure&nbsp;\ref{fig:sim_Phantom} shows the phantom used for a simulated image with dimensions of $500\times 500$ pixels that consists of four smaller phantoms, each of which exhibits differences between homogeneous and heterogeneous regions.  
The regions are formed by independent draws from the $G^0_I$ distribution, with $L = 5$ and varying $\alpha$.

The $\alpha$ parameter of the $G_I^0$ distribution is essential for interpreting texture characteristics. 
Values close to zero greater than $-3$ suggest extremely textured targets, such as urban zones&nbsp;\cite{Frery2019a}. 
As the value decreases, it indicates regions with moderate texture (in $\left[-6,-3\right]$), related to forest zones, while values below $-6$ correspond to textureless regions, such as pasture, agricultural fields, and water bodies&nbsp;\cite{Neto2023}.

```{r sim_Phantom, echo=FALSE, message=FALSE, warning=FALSE, out.width="40%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Simulated image: Light regions in upper/lower left phantoms show homogeneity ($\\alpha=-10$), while dark ones suggest heterogeneity ($\\alpha=-2$). Conversely, in upper/lower right phantoms, dark regions mean homogeneity ($\\alpha=-9$), light ones mean heterogeneity ($\\alpha=-3$). ', fig.width = 4, fig.height = 4}
knitr::include_graphics("../Figures/PNG/Phantom.png")

```

The three test statistics (entropy-based, ${\text{CV}}$, and ${\text{CV}_{\text{MnAD}}}$) are applied to the simulated image using local sliding windows of size $7\times 7$, as shown in Figures&nbsp;\ref{fig:sim_results}(a)--(c). 
For the first test we considered the \(\widetilde{H}_{\text{AO}}\)  estimator due to its superior performance in simulations and shorter processing time compared to other entropy estimators.

```{r sim_results, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results of applying the test statistics: (a)  Entropy-based test; (b)  CV test; (c)   $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c('', '', ''), out.width = c( "40mm","40mm", "40mm"),  fig.ncol=3, fig.width = 40, fig.height = 30}

knitr::include_graphics("../Figures/PNG/test-entropy-Phantom_AO_7W_L5_100b.png")
knitr::include_graphics("../Figures/PNG/cv_values_Phantom.png")
knitr::include_graphics("../Figures/PNG/mnad_Phantom.png")
```

The resulting $p$-values for each test  are shown in Figures&nbsp;\ref{fig:sim_SAR_Images}(a)--(c) and Figures&nbsp;\ref{fig:sim_SAR_Images_p05}(a)--(c), respectively. 
The new images reveal that a lighter  background indicates homogeneity, while darker backgrounds represent heterogeneous areas.

```{r sim_SAR_Images, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Map of $p$-values: (a) Entropy-based test; (b)  CV test; (c)  $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c('', '', ''), out.width = c( "40mm","40mm", "40mm"),  fig.ncol=3, fig.width = 40, fig.height = 30}

knitr::include_graphics("../Figures/PNG/H_pvalue_Phantom_AO_7x7_L5_100b.png")
knitr::include_graphics("../Figures/PNG/cv_p_values_Phantom.png")
knitr::include_graphics("../Figures/PNG/mnad_p_values_Phantom.png")
```
```{r sim_SAR_Images_p05, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results for a threshold of $0.05$ of the $p$-value: (a) Entropy-based test; (b) CV test; (c) $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm","40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/H_005_pvalues_Phantom_AO_7x7.png")
knitr::include_graphics("../Figures/PNG/nominal_level_005_pvalues-Phantom.png")
knitr::include_graphics("../Figures/PNG/mnad_005_pvalues-Phantom.png")

```

## Real SAR Data

We evaluated the proposed test statistics using three real SAR images. 
For our analysis, we used images of the coast of Jalisco, Mexico, and of Illinois, USA. 
These images were acquired by the Sentinel-1B satellite operating in C-band, with VV polarization, intensity format, \(L=18\) and \(L=36\)  nominal looks.

The first two images have a size of \(512 \times 512\) pixels, while the third has \(1024 \times 1024\) pixels  and they contain mountainous areas, agricultural regions, water bodies and  urban areas, as shown in  Figures&nbsp;\ref{fig:real_SAR_Images_coe}(a)--(c).

The three statistical tests are applied to the SAR images using $7\times 7$ local sliding windows, as illustrated in Figures&nbsp;\ref{fig:real_images_test_Mexico}, \ref{fig:test_lake} and&nbsp;\ref{fig:real_images_test_Illinois}.

 The $p$-values obtained for each test are presented in Figures&nbsp;\ref{fig:Mexico_pvalue}, \ref{fig:lake_pvalue} and&nbsp;\ref{fig:Illinois_crops_pvalue}, respectively.

In Figures &nbsp;\ref{fig:Mexico_crops_0.05}, \ref{fig:lake_0.05} and&nbsp;\ref{fig:Illinois_crops_0.05}, the maps of $p$-values represent the decisions at a $5\%$ significance level. 
Values below  $0.05$  are represented by dark areas, indicating evidence to reject the null hypothesis and suggesting heterogeneity in these regions. 
In contrast, values above 0.05 are represented as white areas, indicating no evidence to reject the fully-developed speckle hypothesis.

```{r real_SAR_Images_coe, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Real SAR images: (a) Coast of Jalisco; (b) Illinois-Region 1; (c) Illinois-Region 2. ',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm", "40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/Mexico_512.png")
knitr::include_graphics("../Figures/PNG/lake_512.png")
knitr::include_graphics("../Figures/PNG/Illinois_1024_36L.png")

```

```{r real_images_test_Mexico, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results of applying the test statistics, Coast of Jalisco: (a)  Entropy-based test using $\\widetilde{H}_{\\text{AO}}$ estimator; (b)  CV test; (c)   $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c('', '', ''), out.width = c( "40mm","40mm", "40mm"),  fig.ncol=3, fig.width = 40, fig.height = 30}

knitr::include_graphics("../Figures/PNG/Entropy_Mexico_512_18L_AO_200b.png")
knitr::include_graphics("../Figures/PNG/cv_mexico_512.png")
knitr::include_graphics("../Figures/PNG/mnad_mexico_512.png")
```

```{r Mexico_pvalue, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Map of $p$-values, Coast of Jalisco: (a) Entropy-based test; (b)  CV test; (c)  $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm","40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/H_pvalue_Mexico_512_18L_AO_200b.png")
knitr::include_graphics("../Figures/PNG/cv_pvalues_mexico_512.png")
knitr::include_graphics("../Figures/PNG/mnad_p_values_mexico_512.png")

```

```{r Mexico_crops_0.05, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results for a threshold of $0.05$ of the $p$-value for the Coast of Jalisco. (a) Entropy-based test; (b) CV test; (c) $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm","40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/H_005__Mexico_512_18L_AO_200b.png")
knitr::include_graphics("../Figures/PNG/cv_005_pvalues_mexico_512.png")
knitr::include_graphics("../Figures/PNG/mnad_005_mexico_512.png")

```

```{r test_lake, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results of applying the test statistics, Illinois-Region 1: (a)  Entropy-based test using $\\widetilde{H}_{\\text{AO}}$ estimator; (b)  CV test; (c)   $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c('', '', ''), out.width = c( "40mm","40mm", "40mm"),  fig.ncol=3, fig.width = 40, fig.height = 30}

knitr::include_graphics("../Figures/PNG/Entropy_lake_512_36L_AO_100b.png")
knitr::include_graphics("../Figures/PNG/cv_lake_512.png")
knitr::include_graphics("../Figures/PNG/mnad_lake_512.png")
```

```{r lake_pvalue, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Map of $p$-values, Illinois-Region 1: (a) Entropy-based test; (b)  CV test; (c)  $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm","40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/H_pvalue_lake_512_36L_AO_100b.png")
knitr::include_graphics("../Figures/PNG/cv_pvalues_lake_512.png")
knitr::include_graphics("../Figures/PNG/mnad_p_values_lake_512.png")

```

```{r lake_0.05, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results for a threshold of $0.05$ of the $p$-value, Illinois-Region 1. (a) Entropy-based test; (b) CV test; (c) $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm","40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/H_005_lake_512_36L_AO_100b.png")
knitr::include_graphics("../Figures/PNG/cv_005_pvalues_lake_512.png")
knitr::include_graphics("../Figures/PNG/mnad_005_lake_512.png")

```

```{r real_images_test_Illinois, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results of applying the test statistics, Illinois-Region 2: (a)  Entropy-based test using $\\widetilde{H}_{\\text{AO}}$ estimator; (b)  CV test; (c)   $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c('', '', ''), out.width = c( "40mm","40mm", "40mm"),  fig.ncol=3, fig.width = 40, fig.height = 30}

knitr::include_graphics("../Figures/PNG/Entropy_Illinois_1024_36L_AO_200b.png")
knitr::include_graphics("../Figures/PNG/cv_Illinois_crops_1024.png")
knitr::include_graphics("../Figures/PNG/mnad_Illinois_crops_1024.png")
```

```{r Illinois_crops_pvalue, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Map of $p$-values, Illinois-Region 2: (a) Entropy-based test; (b)  CV test; (c)  $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm","40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/H_pvalue_Illinois_1024_36L_AO_200b.png")
knitr::include_graphics("../Figures/PNG/cv_pvalues_Illinois_crops_1024.png")
knitr::include_graphics("../Figures/PNG/mnad_p_values_Illinois_crops_1024.png")

```
```{r Illinois_crops_0.05, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results for a threshold of $0.05$ of the $p$-value, Illinois-Region 2. (a) Entropy-based test; (b) CV test; (c) $\\text{CV}_{\\text{MnAD}}$ test.',  fig.subcap=c(' ', ' ',' ' ), out.width = c("40mm", "40mm","40mm"),  fig.ncol=3, fig.width = 20, fig.height = 10}
knitr::include_graphics("../Figures/PNG/H_005_pvalues_Illinois_1024_36L_AO_200b.png")
knitr::include_graphics("../Figures/PNG/cv_005_pvalues_Illinois_crops_1024.png")
knitr::include_graphics("../Figures/PNG/mnad_005_Illinois_crops_1024.png")

```

In general, using Shannon entropy is more meaningful than using the original and robust CV  to capture heterogeneity.

It is justified that the dark areas of the maps based on the $\text{CV}$ and $\text{CV}_{\text{MnAD}}$ tests cover the areas reported for the Shannon entropy-based map.

It is noticeable that the entropy and CV-based tools predicted heterogeneity regions as well as boundaries where the statistical properties of texture vary. 
Considering a higher significance level may increase the sensitivity to edge detection, but it may also increase the risk of detecting false heterogeneous regions.

Moreover, assuming $5\%$ as the threshold for the $p$ values, the dark region produced by the $\text{CV}$ test was lower than produced by the $\text{CV}_{\text{MnAD}}$ and entropy test.

The latter observation is to be expected as the latter tests are robust alternatives to the former test, which is essentially based on the expected value operator that reacts to outliers.

# Conclusion {#sec:conclusion}

In this article, we have provided a practical and theoretical answer to the following physical question: How to detect heterogeneity in SAR images, assuming that the SAR intensity follows the $\Gamma_{\text{SAR}}$ model.
<!-- Solving this problem is not easy, considering that the texturelessness in the parametric space of $\mathcal{G}^0_I$ means the infinity value, which is neither analytically tractable nor practically provable. -->
To this end, we proposed three novel hypothesis tests, one from the Shannon entropy and two others from the variation coefficient variants.
The performance of our proposals was evaluated using a Monte Carlo study. 
The results showed that they were conservative in estimating the probability of a type I error (false alarm rate) and the test power (probability of detection), which increases with sample size.
An application to three recent SAR images was performed. 
The results showed that the Shannon entropy-based test was more robust than the CV-based tests. 
In addition, all tests were able to recognize images with different textures and identified edges where the texture type changes.